---
title: "Reburn analysis"
author: "CT"
date: "2022-11-30"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


This script analyzes data for reburn analyses.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(randomForest)
library(ranger) #a faster version of the RF algorithm
# library(pdp)# for partial dependence plots (PDPs)

library(tidyverse)
library(here)
library(sp)
library(sf)
library(terra)
library(raster)

library(mgcv)
library(tidymv)

library(gridExtra)
library(BiodiversityR)

library(patchwork)


data_dir = readLines(here("data_dir_D.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

Load reburn points with low/moderate initial CBI for plotting

```{r}
data_thin = read_csv(datadir("prelim-outputs/points/thinned_data_newsev1k.csv"))

#dfs for each ecoregion
data_lowmod_ca_notscaled = data_thin %>% filter(ecoregion == "California Coast")
data_lowmod_nm_notscaled = data_thin %>% filter(ecoregion == "Northern Mountains")
data_lowmod_sw_notscaled = data_thin %>% filter(ecoregion == "Southwest")
data_lowmod_wm_notscaled = data_thin %>% filter(ecoregion == "Western Mountains")

```


## Explore relationships and distrubitons of response of predictors variables 

Climate
```{r}
ggplot(data = data_thin, aes(x = ecoregion, y = AET)) +
  geom_boxplot() + theme_bw() + ylab("Actual evapotranspiration (AET)") + xlab("")

ggplot(data = data_thin, aes(x = ecoregion, y = PRISM_ppt_)) +
  geom_boxplot()

ggplot(data = data_thin, aes(x = ecoregion, y = AET)) +
  geom_boxplot()
```


```{r}
#ecoregion pallets 
pal1 = c("#264653","#2A9D8F","#dbd053","#6a0136")
pal2 = c("#264653","#2A9D8F","#883c0c", "#dbd053")


library(scales)
#climate space plot
clim_plot = slice(data_thin, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = ecoregion, x = PRISM_tmea, y = PRISM_ppt_)) +
  geom_point(alpha = .3, size = 1.8)+
  stat_ellipse(size = 1, type = "t")+ # 95% confidence level for a multivariate t-distribution
  theme_bw(13)+
  ylab("Precipitation (mm)") +
  xlab(expression("Temperature " ( degree*C)))+
  scale_color_manual(values = pal2) + 
  theme(legend.position="bottom", legend.title=element_blank())+
  guides(colour = guide_legend(nrow = 2))

#weather space plot
weather_plot = slice(data_thin, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = ecoregion, x = erc_perc, y = vpd_perc)) +
  geom_point(alpha = .3, size = 1.8)+
  # stat_ellipse(size = 1, type = "t")+ # 95% confidence level for a multivariate t-distribution
  theme_bw(13)+
  ylab("ERC percentile") +
  xlab("VPD percentile")+
  scale_color_manual(values = pal2) + 
  theme(legend.position = "none")

weather_plot  + clim_plot + plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect") & theme(legend.position = 'bottom') & theme(legend.title=element_blank())
# # ggsave('scatter_plots.png', height = 4.5, width = 7)
```

Histograms
```{r}
## Histograms 
library(ggridges)
#time between
time_hist = ggplot(data = data_thin, aes(fill = ecoregion, x = time_betwe*10)) +
  geom_histogram(bins = 20)+
  theme_bw(13)+
  xlab("time between initial fire and reburn") +
  scale_fill_manual(values = pal2)+
  facet_wrap(~ ecoregion)+
      theme(legend.position = "none")


#reburn cbi
reburn_hist = ggplot(data = data_thin, aes(fill = ecoregion, x = cbi_reburn)) +
  geom_histogram()+
  theme_bw(13)+
  xlab("Reburn CBI") +
    ylab("")+
  scale_fill_manual(values = pal2)+
  scale_x_continuous(breaks = c(0.51, 1, 1.5, 2, 2.5))+
  facet_wrap(~ecoregion)+
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


#initial cbi
initial_hist = ggplot(data = data_thin, aes(fill = ecoregion, x = cbi_initial)) +
  geom_histogram()+
  theme_bw(13)+
  xlab("Initial CBI") +
      ylab("")+
  scale_fill_manual(values = pal2)+
  scale_x_continuous(breaks = c(0.51, 1, 1.5, 2, 2.5))+
  facet_wrap(~ecoregion)+
        theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")


time_hist + initial_hist + reburn_hist

#raincloud plots
library(ggforce)
library(ggdist)
#time between but only one point per unique fire combo instead of by sample point
time_plot = data_thin %>%
    group_by(initial_fireID, reburn_fireID, ecoregion) %>%
    summarise(time_betwe = mean(time_betwe)) %>%
    ggplot(aes(x = time_betwe*10, y = ecoregion, fill = ecoregion, color = ecoregion)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA) + 
  geom_point(size = 1, alpha = .05,
    position = position_jitter(seed = 1, width = .2, height = .2))  +
    geom_boxplot(width = .15, outlier.shape = NA, fill = NA, color = "black", size = 0.5) +
  xlab("Years between initial fire and reburn") +
  ylab("")+
  scale_fill_manual(values = pal2)+
  scale_color_manual(values = pal2)+
      theme_bw(13)+
      theme(legend.position = "none")


```

Variable raincloud plots
```{r}
##Raincloud plots
rain_plot_func = function(var){
  p = ggplot(data_thin, aes(x = var, y = ecoregion, 
                                   fill = ecoregion, color =
                          ecoregion)) + 
    ggdist::stat_halfeye(
      adjust = .5, width = .6, .width = 0, justification = -.2,
      point_colour = NA) + 
    geom_point(size = 1, alpha = .05,
               position = position_jitter(seed = 1, height = .15))+
    geom_boxplot(width = .15, outlier.shape = NA, 
                 fill = NA, color = "black", size = 0.5) +
    ylab("")+
    scale_fill_manual(values = pal2)+
    scale_color_manual(values = pal2)+
    theme_bw(13)+
    ylab("")
  
  return(p)
}

#initial cbi
(icbi_plot = rain_plot_func(data_thin$cbi_initial) +
        xlab("Initial fire severity (CBI)") +
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none") +
        scale_x_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5)))

#reburn cbi
(rcbi_plot = rain_plot_func(data_thin$cbi_reburn) +
        xlab("Reburn fire severity (CBI)") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")+
        scale_x_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5,3)))

#reburn cbi
(tbf_plot = rain_plot_func(data_thin$time_betwe*10) +
        xlab("Time between fires (years)") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))

#wind speed
(vs_plot = rain_plot_func(data_thin$vs_perc) +
        xlab("Wind speed percentile") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))

(erc_plot = rain_plot_func(data_thin$erc_perc) +
        xlab("ERC percentile") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))

(vpd_plot = rain_plot_func(data_thin$vpd_perc) +
        xlab("VPD percentile") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))



rcbi_plot + tbf_plot + erc_plot +
  icbi_plot + vs_plot + vpd_plot  + plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect") & theme(legend.position = 'bottom') & theme(legend.title=element_blank())

ggsave(datadir("figures/predictor_plots_newsev1k.png"), width = 9, height = 6)
``` 

Explore vegetation
```{r}
data_thin %>% group_by(CT_lumped_codes4) %>% summarise(n = n())

pal3 = c("#FFD662", "#244630", "#cb8c04","#46895D", "#B4C692", "#600060", "#4F3511", "#4682b4", "#264653")

library(formattable)

veg1 = data_thin %>% group_by(CT_lumped_codes4) %>%
  summarise(n = n()) %>% 
  mutate(freq = formattable::percent(n / sum(n)))

veg = data_thin %>% group_by(ecoregion, CT_lumped_codes4) %>%
  summarise(n = n()) %>% 
  mutate(freq = formattable::percent(n / sum(n)))
  


(vegplot = data_thin %>% group_by(CT_lumped_codes4, ecoregion) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x = n, y = ecoregion, fill = CT_lumped_codes4))+
  geom_bar(position = "fill", stat = "identity")+
  theme_bw(13)+
  scale_fill_manual(values = pal3, name = "forest type")+
  ylab("")+
  xlab("")+
  theme(legend.position = "bottom", legend.title=element_blank())+
  guides(fill = guide_legend(nrow = 4)))

(vegplot_all = data_thin %>% mutate(CT_lumped_codes4 = replace(CT_lumped_codes4, CT_lumped_codes4 == "Pine-Oak Woodland", "Pine-Oak"),
                                    CT_lumped_codes4 = replace(CT_lumped_codes4, CT_lumped_codes4 == "Pinyon-Juniper Woodland", "Pinyon-Juniper")) %>%
    ggplot(aes(x = CT_lumped_codes4, fill = CT_lumped_codes4))+
  geom_bar()+
  theme_bw(13)+
  scale_fill_manual(values = pal3, name = "forest type")+
  ylab("")+
  xlab("")+
  theme(legend.position = "none", legend.title=element_blank())+
  guides(fill = guide_legend(nrow = 3))+
  theme(axis.text.x = element_text(angle = 30, hjust=1))+
    ggtitle('Forest Type'))

library(PNWColors)
pal_moth = pnw_palette("Moth", 4)

(frg_all = ggplot(data_thin %>% mutate(across('FRG', str_replace, '-B|-C|-A', '')), aes(x = FRG, fill = FRG))+
  geom_bar()+
  theme_bw(13)+
  scale_fill_manual(values = pal_moth, name = "forest type")+
  ylab("")+
  xlab("")+
  theme(legend.position = "none", legend.title=element_blank())+
  guides(fill = guide_legend(nrow = 3))+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
    ggtitle('Fire Regime Group'))

(frgplot = data_thin  %>% mutate(across('FRG', str_replace, '-B|-C|-A', '')) %>% group_by(FRG, ecoregion) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x = n, y = ecoregion, fill = FRG))+
  geom_bar(position = "fill", stat = "identity")+
  theme_bw(13)+
  scale_fill_manual(values = pal_moth, name = "forest type")+
  ylab("")+
  xlab("")+
  theme(legend.position = "bottom", legend.title=element_blank())+
  guides(fill = guide_legend(nrow = 1)))

(vegplot_all + frg_all)/(vegplot + frgplot)+ plot_annotation(tag_levels = 'a')

ggsave(datadir("figures/veg_frg_newsev1k.png"), width = 9, height = 7.5)
```


```{r}
  
veg_table = data_thin %>% dplyr::select(CT_lumped_codes4, BPS_NAME) %>% distinct() %>% arrange(CT_lumped_codes4, BPS_NAME)

#save plot
clim_plot + vegplot + plot_annotation(tag_levels = 'a')

# # ggsave(datadir("figures/ecoregion_overview.png"), width = 10, height = 6)

#veg-climate space plot
(veg_clim_plot = slice(climate_bins, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = CT_lumped_codes4, x = PRISM_tmea, y = PRISM_ppt_)) +
  geom_point(alpha = .3, size = 1.8)+
  theme_bw(13)+
  ylab("Precipitation (mm)") +
  xlab(expression("Temperature " ( degree*C)))+
  scale_color_manual(values = pal3) + 
  theme(legend.position="bottom", legend.title=element_blank())+
  guides(colour = guide_legend(nrow = 2)) +
  facet_grid(~ecoregion, scales = "free")
  # stat_ellipse(aes(group = climate_bin), size = 1, type = "t") # 95% confidence level for a multivariate t-distribution
  )

climate_bins %>% group_by(CT_lumped_codes4, ecoregion, climate_bin) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x = n, y = ecoregion, fill = CT_lumped_codes4))+
  geom_bar(position = "fill", stat = "identity")+
  theme_bw(13)+
  scale_fill_manual(values = pal3, name = "forest type")+
  ylab("")+
  xlab("")+
  theme(legend.position = "bottom", legend.title=element_blank())+
  guides(fill = guide_legend(nrow = 5)) +
  facet_wrap(~climate_bin)

library(ggforce)
veg_clim_plot + geom_ellipse(aes(color = climate_bin), alpha = 0.2) +
  coord_fixed(ratio = 1)

#explore fire regime groups
xtabs(~FRG+climate_bin, data = climate_bins)
xtabs(~FRG+ecoregion, data = climate_bins)

xtabs(~CT_lumped_codes4+ecoregion, data = data_thin)
xtabs(~CT_lumped_codes4, data = data_thin)
xtabs(~CT_lumped_codes4+climate_bin+ecoregion, data = climate_bins)

#create climate quantiles for plotting
clim_p = data_thin %>% group_by(ecoregion) %>%
  summarise(warmwet = quantile(PRISM_ppt_, .7),
            hotdry = quantile(PRISM_ppt_, .2),
            coldwet = quantile(PRISM_ppt_, .8)) %>%
  pivot_longer(warmwet:coldwet, names_to = 'clim', values_to = 'PRISM_ppt_')

clim_t = data_thin %>% group_by(ecoregion) %>%
  summarise(warmwet = quantile(PRISM_tmea, .7),
            hotdry = quantile(PRISM_tmea, .85),
            coldwet = quantile(PRISM_tmea, .2)) %>%
  pivot_longer(warmwet:coldwet, names_to = 'clim', values_to = 'PRISM_tmea') %>%
  full_join(clim_p)

(veg2_plot = slice(data_thin, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = CT_lumped_codes4, x = PRISM_tmea, y = PRISM_ppt_)) +
  geom_point(alpha = .4, size = 1.8)+
  theme_bw(13)+
  ylab("Precipitation (mm)") +
  xlab(expression("Temperature " ( degree*C)))+
  scale_color_manual(values = pal3) + 
  theme(legend.position="bottom", legend.title=element_blank())+
  guides(color = guide_legend(nrow = 3, override.aes = list(size =3, alpha = .9)))+
  facet_wrap(~ecoregion, scales = "free")) 

veg2_plot + geom_point(data = clim_t, aes(x = PRISM_tmea, y = PRISM_ppt_, group = ecoregion), color = c("black"), cex = 4, pch = 10) #doesn't recognize lwd?
  # ggrepel::geom_text_repel(data = clim_t, aes(label = clim, group = ecoregion), color = "black")



  # stat_ellipse(size = 1, type = "t")) # 95% confidence level for a multivariate t-distribution
ggsave(datadir("figures/clim_veg_newsev1k.png"), width = 7, height = 6)

```


### Predict from models for plotting

Load ecoregion models

```{r}
fit_ca = readRDS(datadir("final-models/fit_ca_newsev1k.rds"))
fit_sw = readRDS(datadir("final-models/fit_sw_newsev1k.rds"))
fit_nm = readRDS(datadir("final-models/fit_nm_newsev1k.rds"))
fit_wm = readRDS(datadir("final-models/fit_wm_newsev1k.rds"))
```

Get scaled dataframes for predictions

```{r}
#scale numeric predictors for model fit - don't scale time between t or cbi or erc percentiles to maintain interpretation
preds_2scale = c("heat_load", "PRISM_tmea", "PRISM_ppt_", "TPI_1024", "fm100_02_2", "th_02_21", "vpd_02_21", "erc_02_21", "PRISM_vpdm", "AET", "fm1000_02_", "ndvi", "shrub_rebrn", "tree_reburn", "afg_rebrn", "pfg_rebrn", "cbi_initial", "time_betwe", "erc_perc", "vpd_perc", "vs_perc")

data_thin = data.frame(data_thin)

scale_df = function(df){
  data_scale = scale(df[(names(df) %in% preds_2scale)], scale = TRUE, center = TRUE)
  data_scale_all = cbind(data_scale, df[!(names(df) %in% preds_2scale)])
  return(data_scale_all)
}

data_thin_scale = data_thin %>% scale_df()

#create new dfs for each ecoregion and scale each ecoregion separately 
data_lowmod_ca = data_thin %>% filter(ecoregion == "California Coast") %>% scale_df()
data_lowmod_nm = data_thin %>% filter(ecoregion == "Northern Mountains") %>% scale_df()
data_lowmod_sw = data_thin %>% filter(ecoregion == "Southwest") %>% scale_df()
data_lowmod_wm = data_thin %>% filter(ecoregion == "Western Mountains") %>% scale_df()

unscale_df = data_thin %>%
  group_by(ecoregion) %>%
  summarise(timebetwe_sd = sd(time_betwe),
            timebetwe_mean = mean(time_betwe),
            cbi_initial_sd = sd(cbi_initial),
            cbi_initial_mean = mean(cbi_initial),
            erc_perc_sd = sd(erc_perc),
            erc_perc_mean = mean(erc_perc),
            vpd_perc_sd = sd(vpd_perc),
            vpd_perc_mean = mean(vpd_perc),
            vs_perc_sd = sd(vs_perc),
            vs_perc_mean = mean(vs_perc))
```


Just time between
```{r}

preds_func = function(gam_fit, df){
  preds = predict_gam(gam_fit, length_out = 35, values = list(cbi_initial =0, time_betwe = seq(min(df$time_betwe), max(df$time_betwe), by = .05), erc_perc = 0, PRISM_tmea = 0, PRISM_ppt_ = 0, vs_perc = 0, vpd_perc = 0, heat_load = 0, TPI_1024 = 0, count_burn = 2)) 
  return(preds)
}

ca_preds = preds_func(fit_ca, data_lowmod_ca)
nm_preds = preds_func(fit_nm, data_lowmod_nm)
sw_preds = preds_func(fit_sw, data_lowmod_sw)
wm_preds = preds_func(fit_wm, data_lowmod_wm)

ca_preds$time_betwe_unscale = (ca_preds$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])
nm_preds$time_betwe_unscale = (nm_preds$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])
sw_preds$time_betwe_unscale = (sw_preds$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])
wm_preds$time_betwe_unscale = (wm_preds$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])

ca_preds$cbi_reburn = ca_preds$fit
nm_preds$cbi_reburn = nm_preds$fit
sw_preds$cbi_reburn = sw_preds$fit
wm_preds$cbi_reburn = wm_preds$fit


plot_preds = function(preds, initial_cbi){
  
  plot = preds %>%
  ggplot(aes(time_betwe_unscale*10, fit)) +
    geom_smooth_ci(size = 1.05, legend = NA)+
  theme(legend.title = element_blank())+
  ylim(0.1, 2.5) +
    xlim(0, 36)+
    xlab("time between fires (yrs)") +
    ylab("reburn severity (CBI)") +
  geom_hline(yintercept = mean(initial_cbi), linetype = "longdash")+ 
  geom_hline(yintercept = sd(initial_cbi) + mean(initial_cbi), linetype = "dotted")+ 
  geom_hline(yintercept = mean(initial_cbi) - sd(initial_cbi), linetype = "dotted") +
  annotate("text", y= mean(initial_cbi)-.2, x=30, label="initial CBI")+
  theme_bw(13) 
  
  return(plot)
}

(pca1 = plot_preds(ca_preds, data_lowmod_ca_notscaled$cbi_initial) + ggtitle("California Coast"))

psw1 = plot_preds(sw_preds,data_lowmod_sw_notscaled$cbi_initial) + ggtitle("Southwest")+ylab("")

pnm1 = plot_preds(nm_preds, data_lowmod_nm_notscaled$cbi_initial) + ggtitle("Northern Mountains")+ylab("")+xlab("")

pwm1 = plot_preds(wm_preds, data_lowmod_wm_notscaled$cbi_initial) + ggtitle("Western Mountains")+xlab("")


pwm1 + pnm1 + pca1 + psw1 + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/gams_average_newsev1k.png'), height = 5, width = 5)

```


Weather predictions - scaled
```{r}
#prediction df with most variables set to mean = 0
preds_dfw_func = function(df, fit, vpd_mean, vpd_sd, erc_mean, erc_sd){
  df_w97 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      # AET = 0, 
                      # afg_rebrn = 0, pfg_rebrn = 0,
                      #set weather variables to 3rd quantile value by ecoregion
                      vpd_perc = as.numeric((.98-vpd_mean)/vpd_sd),
                      erc_perc = as.numeric((.98-erc_mean)/erc_sd),
                      # vs_02_21 = quantile(df$vs_02_21, 0.95),
                      vs_perc = 0,
                      #set time between fires to full range 1-36 years
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      weather = "extreme")
  
  df_w90 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = as.numeric((.95-vpd_mean)/vpd_sd),
                      erc_perc = as.numeric((.95-erc_mean)/erc_sd),
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      weather = "very high")

  df_w80 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = as.numeric((.85-vpd_mean)/vpd_sd),
                      erc_perc = as.numeric((.85-erc_mean)/erc_sd),
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      weather = "high")
  
  
  
  data_w = rbind(df_w97, df_w80, df_w90)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE)
  
  #add predictions and CIs to df for plotting
  data_w$cbi_reburn = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  #make weather a factor for plotting
  data_w$weather = factor(data_w$weather, levels = c("high", "very high", "extreme"))
  
  return(data_w)
}

#predict to weather values (calcualte scaled values of .97, .9, and .8 percentiles)
preds_w_ca <- preds_dfw_func(df = data_lowmod_ca, fit = fit_ca,
                             vpd_mean = unscale_df[1, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[1, 'vpd_perc_sd'],
                             erc_mean = unscale_df[1, 'erc_perc_mean'],
                             erc_sd = unscale_df[1, 'erc_perc_sd'])

preds_w_nm <- preds_dfw_func(df = data_lowmod_nm, fit = fit_nm,
                             vpd_mean = unscale_df[2, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[2, 'vpd_perc_sd'],
                             erc_mean = unscale_df[2, 'erc_perc_mean'],
                             erc_sd = unscale_df[2, 'erc_perc_sd'])

preds_w_sw <- preds_dfw_func(df = data_lowmod_sw, fit = fit_sw,
                             vpd_mean = unscale_df[3, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[3, 'vpd_perc_sd'],
                             erc_mean = unscale_df[3, 'erc_perc_mean'],
                             erc_sd = unscale_df[3, 'erc_perc_sd'])

preds_w_wm <- preds_dfw_func(df = data_lowmod_wm, fit = fit_wm,
                             vpd_mean = unscale_df[4, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[4, 'vpd_perc_sd'],
                             erc_mean = unscale_df[4, 'erc_perc_mean'],
                             erc_sd = unscale_df[4, 'erc_perc_sd'])

#add in unscaled time_between for plotting
preds_w_ca$time_betwe_unscale = round((preds_w_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"]),1)

preds_w_nm$time_betwe_unscale = (preds_w_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_w_sw$time_betwe_unscale = (preds_w_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_w_wm$time_betwe_unscale = (preds_w_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```

wind speed predictions - scaled
```{r}
#prediction df with most variables set to mean = 0
preds_dfvs_func = function(df, fit, vs_mean, vs_sd){
  df_w97 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      # AET = 0, 
                      # afg_rebrn = 0, pfg_rebrn = 0,
                      #set weather variables to 3rd quantile value by ecoregion
                      vs_perc = as.numeric((.75-vs_mean)/vs_sd),
                      # vs_02_21 = quantile(df$vs_02_21, 0.95),
                      vpd_perc = 0, erc_perc = 0,
                      #set time between fires to full range 1-25 years
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      windspeed = "high")
  
  df_w90 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      #set weather variables to 3rd quantile value by ecoregion
                      vs_perc = as.numeric((.5-vs_mean)/vs_sd),
                      # vs_02_21 = quantile(df$vs_02_21, 0.95),
                      vpd_perc = 0, erc_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      windspeed = "moderate")

  df_w80 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      #set weather variables to 3rd quantile value by ecoregion
                      vs_perc = as.numeric((.25-vs_mean)/vs_sd),
                      # vs_02_21 = quantile(df$vs_02_21, 0.95),
                      vpd_perc = 0, erc_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      windspeed = "low")
  
  data_w = rbind(df_w97, df_w80, df_w90)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE)
  
  #add predictions and CIs to df for plotting
  data_w$cbi_reburn = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  #make weather a factor for plotting
  data_w$windspeed = factor(data_w$windspeed, levels = c("low", "moderate", "high"))
  
  return(data_w)
}

#predict to weather values (calcualte scaled values of .97, .9, and .8 percentiles)
preds_vs_ca <- preds_dfvs_func(df = data_lowmod_ca, fit = fit_ca,
                             vs_mean = unscale_df[1, 'vs_perc_mean'],
                             vs_sd = unscale_df[1, 'vs_perc_sd'])

preds_vs_nm <- preds_dfvs_func(df = data_lowmod_nm, fit = fit_nm,
                             vs_mean = unscale_df[2, 'vs_perc_mean'],
                             vs_sd = unscale_df[2, 'vs_perc_sd'])

preds_vs_sw <- preds_dfvs_func(df = data_lowmod_sw, fit = fit_sw,
                             vs_mean = unscale_df[3, 'vs_perc_mean'],
                             vs_sd = unscale_df[3, 'vs_perc_sd'])

preds_vs_wm <- preds_dfvs_func(df = data_lowmod_wm, fit = fit_wm,
                             vs_mean = unscale_df[4, 'vs_perc_mean'],
                             vs_sd = unscale_df[4, 'vs_perc_sd'])

#add in unscaled time_between for plotting
preds_vs_ca$time_betwe_unscale = round((preds_vs_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"]),1)

preds_vs_nm$time_betwe_unscale = (preds_vs_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_vs_sw$time_betwe_unscale = (preds_vs_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_vs_wm$time_betwe_unscale = (preds_vs_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```


CBI predictions - scaled
```{r}
#prediction df with most variables set to mean = 0
preds_dfcbi_func = function(df, fit, cbi_mean, cbi_sd){
  df_lowcbi = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = as.numeric((0.5-cbi_mean)/cbi_sd), 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      # AET = 0, 
                      # afg_rebrn = 0, pfg_rebrn = 0,
                      #set weather variables to 3rd quantile value by ecoregion
                      vpd_perc = 0,
                      erc_perc = 0,
                      # vs_02_21 = quantile(df$vs_02_21, 0.95),
                      vs_perc = 0,
                      #set time between fires to full range 1-25 years
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      cbii_cat = "low")

  
  df_modcbi = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = as.numeric((2-cbi_mean)/cbi_sd), 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = 0,
                      erc_perc = 0,
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                      cbii_cat = "mod")


  
  data_w = rbind(df_lowcbi, df_modcbi)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE)
  
  #add predictions and CIs to df for plotting
  data_w$cbi_reburn = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  
  return(data_w)
}

#predict to cbi values 
preds_cbi_ca <- preds_dfcbi_func(df = data_lowmod_ca, fit = fit_ca,
                             cbi_mean = unscale_df[1, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[1, 'cbi_initial_sd'])

preds_cbi_nm <- preds_dfcbi_func(df = data_lowmod_nm, fit = fit_nm,
                             cbi_mean = unscale_df[2, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[2, 'cbi_initial_sd'])

preds_cbi_sw <- preds_dfcbi_func(df = data_lowmod_sw, fit = fit_sw,
                             cbi_mean = unscale_df[3, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[3, 'cbi_initial_sd'])

preds_cbi_wm <- preds_dfcbi_func(df = data_lowmod_wm, fit = fit_wm,
                             cbi_mean = unscale_df[4, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[4, 'cbi_initial_sd'])

#add in unscaled time_between for plotting
preds_cbi_ca$time_betwe_unscale = (preds_cbi_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])

preds_cbi_nm$time_betwe_unscale = (preds_cbi_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_cbi_sw$time_betwe_unscale = (preds_cbi_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_cbi_wm$time_betwe_unscale = (preds_cbi_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```

climate predictions
```{r}
# #find west wide climate quantiles


#prediction df with variables set to mean = 0
preds_dfc_func = function(df, fit){
  #cool wet - high productivity
  df_coldwet = data.frame(count_burn = 2, 
                        #set climate vars
                       PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.8),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.2),
                        #control for other vars - mean
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc  = 0, erc_perc = 0,
                       vs_perc = 0,
                       #set time between fires to full range 1-36 years
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                       climate = "coldwet")

  #hot dry - low productivity
  df_hotdry = data.frame(count_burn = 2, 
                        #set climate vars
                       # AET = quantile(df$AET, 0.25), 
                       PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.2),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.85),
                        #control for other vars - mean
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc  = 0, erc_perc = 0,
                       vs_perc = 0,
                       #set time between fires to full range 1-25 years
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                       climate = "hotdry")

  
    #hot dry - low productivity
  df_warmwet = data.frame(count_burn = 2, 
                        #set climate vars
                       # AET = quantile(df$AET, 0.25), 
                       PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.7),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.7),
                        #control for other vars - mean
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc  = 0, erc_perc = 0,
                       vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
                       climate = "hotwet")
  
  #     #cold dry - low productivity
  # df_colddry = data.frame(count_burn = 2, 
  #                       #set climate vars
  #                      # AET = quantile(df$AET, 0.25), 
  #                      PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.25),
  #                      PRISM_tmea = quantile(df$PRISM_tmea, 0.25),
  #                       #control for other vars - mean
  #                      cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
  #                      vpd_perc  = 0, erc_perc = 0,
  #                      vs_perc = 0,
  #                     time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 35),
  #                      climate = "colddry")
  # 
  
  
  data_c = rbind(df_coldwet, df_hotdry, df_warmwet)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_c, se.fit = TRUE)
  
  #add predictions to df for plotting
  data_c$cbi_reburn = p$fit
  data_c$upr = p$fit + (2 * p$se.fit)
  data_c$lwr = p$fit - (2 * p$se.fit)
  return(data_c)
}

preds_c_ca <- preds_dfc_func(df = data_lowmod_ca, fit = fit_ca)
preds_c_nm <- preds_dfc_func(df = data_lowmod_nm, fit = fit_nm)
preds_c_sw <- preds_dfc_func(df = data_lowmod_sw, fit = fit_sw)
preds_c_wm <- preds_dfc_func(df = data_lowmod_wm, fit = fit_wm)

#add in unscaled time_between for plotting
preds_c_ca$time_betwe_unscale = (preds_c_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])

preds_c_nm$time_betwe_unscale = (preds_c_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_c_sw$time_betwe_unscale = (preds_c_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_c_wm$time_betwe_unscale = (preds_c_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```


### Determine when slopes flatten for each prediction

CBI
```{r}
library(emmeans)
library(modelbased)
#CBI

#calculate time between slopes for 2 levels of cbi for each ecoregion
ca_cbi_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_ca$cbi_initial)"), length = 35, ci=0.95)

nm_cbi_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_nm$cbi_initial)"), length = 35, ci=0.95)

sw_cbi_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_sw$cbi_initial)"), length = 35, ci=0.95)

wm_cbi_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_wm$cbi_initial)"), length = 35, ci=0.95)

# plot(modelbased::visualisation_recipe(ca_cbi_slopes))+theme_bw()
# plot(modelbased::visualisation_recipe(nm_cbi_slopes))+theme_bw()
# plot(modelbased::visualisation_recipe(sw_cbi_slopes))+theme_bw()
# plot(modelbased::visualisation_recipe(wm_cbi_slopes))+theme_bw()

#add to predictor df
preds_cbi_ca2 = preds_cbi_ca %>%
    left_join(ca_cbi_slopes, by = c("time_betwe", "cbi_initial"))

preds_cbi_nm2 = preds_cbi_nm %>%
    left_join(nm_cbi_slopes, by = c("time_betwe", "cbi_initial"))

preds_cbi_sw2 = preds_cbi_sw %>%
    left_join(sw_cbi_slopes, by = c("time_betwe", "cbi_initial"))

preds_cbi_wm2 = preds_cbi_wm %>%
    left_join(wm_cbi_slopes, by = c("time_betwe", "cbi_initial"))

#plot ggplot
ggplot(preds_cbi_wm2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(cbii_cat))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()
```

Weather
```{r}
#calculate time between slopes for 3 levels of weather (erc & vpd) for each ecoregion
ca_w_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_ca$erc_perc)", "vpd_perc = unique(preds_w_ca$vpd_perc)"), length = 35, ci=0.95)

nm_w_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_nm$erc_perc)", "vpd_perc = unique(preds_w_nm$vpd_perc)"), length = 35, ci=0.95)

sw_w_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_sw$erc_perc)", "vpd_perc = unique(preds_w_sw$vpd_perc)"), length = 35, ci=0.95)


wm_w_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_wm$erc_perc)", "vpd_perc = unique(preds_w_wm$vpd_perc)"), length = 35, ci=0.95)

#add slope coefficients to predictor dfs
preds_w_ca2 = preds_w_ca %>%
    left_join(ca_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

preds_w_nm2 = preds_w_nm %>%
    left_join(nm_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

preds_w_sw2 = preds_w_sw %>%
    left_join(sw_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

preds_w_wm2 = preds_w_wm %>%
    left_join(wm_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

#plot
ggplot(preds_w_sw2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(weather))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()
```

wind
```{r}
#calculate time between slopes for 3 levels of windspeed for each ecoregion
ca_vs_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_ca$vs_perc)"), length = 35, ci=0.95)

nm_vs_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_nm$vs_perc)"), length = 35, ci=0.95)

sw_vs_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_sw$vs_perc)"), length = 35, ci=0.95)

wm_vs_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_wm$vs_perc)"), length = 35, ci=0.95)

plot(modelbased::visualisation_recipe(ca_vs_slopes))+theme_bw()
plot(modelbased::visualisation_recipe(nm_vs_slopes))+theme_bw()
plot(modelbased::visualisation_recipe(sw_vs_slopes))+theme_bw()
plot(modelbased::visualisation_recipe(wm_vs_slopes))+theme_bw()

#add to predictor df
preds_vs_ca2 = preds_vs_ca %>%
    left_join(ca_vs_slopes, by = c("time_betwe", "vs_perc"))

preds_vs_nm2 = preds_vs_nm %>%
    left_join(nm_vs_slopes, by = c("time_betwe", "vs_perc"))

preds_vs_sw2 = preds_vs_sw %>%
    left_join(sw_vs_slopes, by = c("time_betwe", "vs_perc"))

preds_vs_wm2 = preds_vs_wm %>%
    left_join(wm_vs_slopes, by = c("time_betwe", "vs_perc"))

#plot ggplot
ggplot(preds_vs_sw2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(windspeed))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()

```

Climate
```{r}
#Climate 
#calculate time between slopes for 3 levels of climate (temp & precip) for each ecoregion
ca_c_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_ca$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_ca$PRISM_ppt_)"), length = 35, ci=0.95)

nm_c_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_nm$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_nm$PRISM_ppt_)"), length = 35, ci=0.95)

sw_c_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_sw$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_sw$PRISM_ppt_)"), length = 35, ci=0.95)

wm_c_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_wm$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_wm$PRISM_ppt_)"), length = 35, ci=0.95)

#add slope coefficients to predictor dfs
preds_c_ca2 = preds_c_ca %>%
    left_join(ca_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

preds_c_nm2 = preds_c_nm %>%
    left_join(nm_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

preds_c_sw2 = preds_c_sw %>%
    left_join(sw_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

preds_c_wm2 = preds_c_wm %>%
    left_join(wm_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

#plot
ggplot(preds_c_ca2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(climate))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()
```



### Prep predictor DF for plotting


Truncate predictor data frames where there is no data on low and high end of time between - create bins for prediction categories
```{r}

#create new variables for each interaction category
bins = data_thin %>%
  mutate(cbi_bins = case_when(cbi_initial < 1.25 ~ "low",
                             cbi_initial >= 1.25 ~ "mod"),
         vs_bins = case_when(vs_perc <= .375 ~ "low",
                             vs_perc > .375 & vs_perc <= .625 ~ "moderate",
                             vs_perc > .625 ~ "high"),
         weather_bins = case_when(erc_perc > .97 & vpd_perc > .97 ~ "extreme",
                             erc_perc <= .9 & vpd_perc <= .9 ~ "high",
                             erc_perc > .9 & vpd_perc > .9 &
                              erc_perc <= .97 & vpd_perc <= .97 ~ "very high"),
         time_fact = as.factor(time_betwe))

#weather bins
weather_bins = bins %>% group_by(ecoregion, weather_bins) %>%
  summarise(count = n(),
            min_time = min(time_betwe), #get minimum and maximum time between for each predictor combination
            max_time = max(time_betwe)) %>%
  na.omit()

#wind speed bins
vs_bins = bins %>% group_by(ecoregion, vs_bins) %>%
  summarise(count = n(),
            min_time = min(time_betwe),
            max_time = max(time_betwe))

#cbi bins
cbi_bins = bins %>% group_by(ecoregion, cbi_bins) %>%
  summarise(count = n(),
            min_time = min(time_betwe),
            max_time = max(time_betwe))

##climate bins
climate_bins = data_thin %>% group_by(ecoregion) %>%
  mutate(climate_bin = case_when(PRISM_ppt_ <= quantile(PRISM_ppt_, .4) &
              PRISM_tmea > quantile(PRISM_tmea, .6) ~ "hotdry",
              PRISM_ppt_ > quantile(PRISM_ppt_, .6) &
              PRISM_tmea <= quantile(PRISM_tmea, .4) ~ "coldwet",
              PRISM_ppt_ > quantile(PRISM_ppt_, .5) &
              PRISM_tmea > quantile(PRISM_tmea, .5) ~ "hotwet"))

climate_bins2 = climate_bins %>% group_by(ecoregion, climate_bin) %>%
  summarise(count = n(),
            min_time = min(time_betwe),
            max_time = max(time_betwe)) %>%
 na.omit()

```

#### Truncate predictor data frames where there is no data on low and high end of time between fires

```{r}
#filter out values that fall outside the time between data range
preds_w_ca_trunc = preds_w_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= as.numeric(weather_bins[1,"min_time"]) & extreme <= as.numeric(weather_bins[1,"max_time"]) |
           high >= as.numeric(weather_bins[2,"min_time"]) & high <= as.numeric(weather_bins[2,"max_time"]) |
           `very high` >= as.numeric(weather_bins[3,"min_time"]) & `very high` <= as.numeric(weather_bins[3,"max_time"])) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "weather",
         ecoregion = "California Coast") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

preds_w_nm_trunc = preds_w_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= as.numeric(weather_bins[4,"min_time"]) & extreme <= as.numeric(weather_bins[4,"max_time"]) |
           high >= as.numeric(weather_bins[5,"min_time"]) & high <= as.numeric(weather_bins[5,"max_time"]) |
           `very high` >= as.numeric(weather_bins[6,"min_time"]) & `very high` <= as.numeric(weather_bins[6,"max_time"])) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "weather",
         ecoregion = "Northern Mountains") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

preds_w_sw_trunc = preds_w_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
    filter(extreme >= as.numeric(weather_bins[7,"min_time"]) & extreme <= as.numeric(weather_bins[7,"max_time"]) |
           high >= as.numeric(weather_bins[8,"min_time"]) & high <= as.numeric(weather_bins[8,"max_time"]) |
           `very high` >= as.numeric(weather_bins[9,"min_time"]) & `very high` <= as.numeric(weather_bins[9,"max_time"])) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "weather",
         ecoregion = "Southwest") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

preds_w_wm_trunc = preds_w_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= as.numeric(weather_bins[10,"min_time"]) & extreme <= as.numeric(weather_bins[10,"max_time"]) |
           high >= as.numeric(weather_bins[11,"min_time"]) & high <= as.numeric(weather_bins[11,"max_time"]) |
           `very high` >= as.numeric(weather_bins[12,"min_time"]) & `very high` <= as.numeric(weather_bins[12,"max_time"])) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Western Mountains",
         cat = "weather") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

```

wind speed truncated dfs
```{r}
#filter out values that fall outside the time between data range
preds_vs_ca_trunc = preds_vs_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= as.numeric(vs_bins[1,"min_time"]) & high <=  as.numeric(vs_bins[1,"max_time"]) |
           low >= as.numeric(vs_bins[2,"min_time"]) & low <= as.numeric(vs_bins[2,"max_time"]) |
           moderate >= as.numeric(vs_bins[3,"min_time"]) & moderate <= as.numeric(vs_bins[3,"max_time"])) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "windspeed",
         ecoregion = "California Coast") %>%
  na.omit()

preds_vs_nm_trunc = preds_vs_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= as.numeric(vs_bins[4,"min_time"]) & high <=  as.numeric(vs_bins[4,"max_time"]) |
           low >= as.numeric(vs_bins[5,"min_time"]) & low <= as.numeric(vs_bins[5,"max_time"]) |
           moderate >= as.numeric(vs_bins[6,"min_time"]) & moderate <= as.numeric(vs_bins[6,"max_time"])) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "windspeed",
         ecoregion = "Northern Mountains") %>%
  na.omit()

preds_vs_sw_trunc = preds_vs_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= as.numeric(vs_bins[7,"min_time"]) & high <=  as.numeric(vs_bins[7,"max_time"]) |
           low >= as.numeric(vs_bins[8,"min_time"]) & low <= as.numeric(vs_bins[8,"max_time"]) |
           moderate >= as.numeric(vs_bins[9,"min_time"]) & moderate <= as.numeric(vs_bins[9,"max_time"])) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
   mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "windspeed",
         ecoregion = "Southwest") %>%
  na.omit()

preds_vs_wm_trunc = preds_vs_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= as.numeric(vs_bins[10,"min_time"]) & high <=  as.numeric(vs_bins[10,"max_time"]) |
           low >= as.numeric(vs_bins[11,"min_time"]) & low <= as.numeric(vs_bins[11,"max_time"]) |
           moderate >= as.numeric(vs_bins[12,"min_time"]) & moderate <= as.numeric(vs_bins[12,"max_time"])) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Western Mountains",
         cat = "windspeed") %>%
  na.omit()
```

CBI truncated dfs
```{r}
preds_cbi_ca_trunc = preds_cbi_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= as.numeric(cbi_bins[1,"min_time"]) & low <= as.numeric(cbi_bins[1,"max_time"]) |
           mod >= as.numeric(cbi_bins[2,"min_time"]) & mod <= as.numeric(cbi_bins[2,"max_time"])) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "cbi",
         ecoregion = "California Coast") %>%
  na.omit()

preds_cbi_nm_trunc = preds_cbi_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= as.numeric(cbi_bins[3,"min_time"]) & low <= as.numeric(cbi_bins[3,"max_time"]) |
           mod >= as.numeric(cbi_bins[4,"min_time"]) & mod <= as.numeric(cbi_bins[4,"max_time"])) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "cbi",
         ecoregion = "Northern Mountains") %>%
  na.omit()

preds_cbi_sw_trunc = preds_cbi_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= as.numeric(cbi_bins[5,"min_time"]) & low <= as.numeric(cbi_bins[5,"max_time"]) |
           mod >= as.numeric(cbi_bins[6,"min_time"]) & mod <= as.numeric(cbi_bins[6,"max_time"])) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Southwest",
         cat = "cbi") %>%
  na.omit()

preds_cbi_wm_trunc = preds_cbi_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= as.numeric(cbi_bins[7,"min_time"]) & low <= as.numeric(cbi_bins[7,"max_time"]) |
           mod >= as.numeric(cbi_bins[8,"min_time"]) & mod <= as.numeric(cbi_bins[8,"max_time"])) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Western Mountains",
         cat = "cbi") %>%
  na.omit()

```

Truncate climate predictions
```{r}
preds_c_ca_trunc = preds_c_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
  filter(coldwet >= as.numeric(climate_bins2[1,"min_time"]) & coldwet <=  as.numeric(climate_bins2[1,"max_time"]) |
           hotdry >=  as.numeric(climate_bins2[2,"min_time"]) & hotdry <=  as.numeric(climate_bins2[2,"max_time"]) |
           hotwet >=  as.numeric(climate_bins2[3,"min_time"]) & hotwet <=  as.numeric(climate_bins2[3,"max_time"])) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "California Coast") %>%
    na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

preds_c_nm_trunc = preds_c_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
 filter(coldwet >= as.numeric(climate_bins2[4,"min_time"]) & coldwet <=  as.numeric(climate_bins2[4,"max_time"]) |
           hotdry >=  as.numeric(climate_bins2[5,"min_time"]) & hotdry <=  as.numeric(climate_bins2[5,"max_time"]) |
           hotwet >=  as.numeric(climate_bins2[6,"min_time"]) & hotwet <=  as.numeric(climate_bins2[6,"max_time"])) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "Northern Mountains") %>%
    na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

preds_c_sw_trunc = preds_c_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
 filter(coldwet >= as.numeric(climate_bins2[7,"min_time"]) & coldwet <=  as.numeric(climate_bins2[7,"max_time"]) |
           hotdry >=  as.numeric(climate_bins2[8,"min_time"]) & hotdry <=  as.numeric(climate_bins2[8,"max_time"]) |
           hotwet >=  as.numeric(climate_bins2[9,"min_time"]) & hotwet <=  as.numeric(climate_bins2[9,"max_time"])) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "Southwest") %>%
    na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

preds_c_wm_trunc = preds_c_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
 filter(coldwet >= as.numeric(climate_bins2[10,"min_time"]) & coldwet <=  as.numeric(climate_bins2[10,"max_time"]) |
           hotdry >=  as.numeric(climate_bins2[11,"min_time"]) & hotdry <=  as.numeric(climate_bins2[11,"max_time"]) |
           hotwet >=  as.numeric(climate_bins2[12,"min_time"]) & hotwet <=  as.numeric(climate_bins2[12,"max_time"])) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "Western Mountains") %>%
  na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

```


### Find points where slopes flatten (lower CI crosses 0)
```{r}
#combine all predictions
preds = bind_rows(preds_c_ca_trunc, preds_c_nm_trunc, preds_c_sw_trunc, preds_c_wm_trunc, preds_vs_ca_trunc, preds_vs_nm_trunc, preds_vs_sw_trunc, preds_vs_wm_trunc, preds_w_ca_trunc, preds_w_nm_trunc, preds_w_sw_trunc, preds_w_wm_trunc, preds_cbi_ca_trunc, preds_cbi_nm_trunc, preds_cbi_sw_trunc, preds_cbi_wm_trunc)

flat_slope_preds = preds %>% filter(CI_low <= 0) %>%
  mutate(weather = as.character(weather),
         climate = as.character(climate)) %>% #convert from factor to remove NAs
  mutate_if(is.character, ~replace_na(.,"")) %>%
  mutate(cat2 = paste(climate, cbii_cat, weather, windspeed, sep = ""),
         cat3 = paste(cat, ecoregion, sep = "-")) %>%
  group_by(cat3, cat2) %>%
  slice(which.min(time_betwe_round)) %>% #get the time between when the lower 95% CI first crosses 0
  mutate(x = time_betwe_round*10, 
         y = cbi_reburn) %>%
  dplyr::select(cat3, cat2, x, y)

flat_slope_preds$rowid = seq(1:nrow(flat_slope_preds))


flat_slope_preds2 = preds %>% filter(Coefficient <=0.05 & Coefficient >=-0.05) %>%
  mutate(weather = as.character(weather),
         climate = as.character(climate)) %>% #convert from factor to remove NAs
  mutate_if(is.character, ~replace_na(.,"")) %>%
  mutate(cat2 = paste(climate, cbii_cat, weather, windspeed, sep = ""),
         cat3 = paste(cat, ecoregion, sep = "-")) %>%
  group_by(cat3, cat2) %>%
  slice(which.min(time_betwe_round)) %>% #get the time between when the lower 95% CI first crosses 0
  mutate(x = time_betwe_round*10, 
         y = cbi_reburn) %>%
  dplyr::select(cat3, cat2, x, y)


## find point where point crosses 0 after being non-zero (hard coding these because there aren't that many). Values from preds

#climate nm hotdry
flat_slope_preds[which(flat_slope_preds$cat3 == "climate-Northern Mountains" & flat_slope_preds$cat2 == "hotdry"), "x"] = 27
flat_slope_preds[which(flat_slope_preds$cat3 == "climate-Northern Mountains" & flat_slope_preds$cat2 == "hotdry"), "y"] = 1.409510



#weather nm very high
flat_slope_preds[which(flat_slope_preds$cat3 == "weather-Northern Mountains" & flat_slope_preds$cat2 == "very high"), "x"] = 13
flat_slope_preds[which(flat_slope_preds$cat3 == "weather-Northern Mountains" & flat_slope_preds$cat2 == "very high"), "y"] = 1.673661

#weather nm very high
flat_slope_preds[which(flat_slope_preds$cat3 == "weather-Northern Mountains" & flat_slope_preds$cat2 == "extreme"), "x"] = 28

flat_slope_preds[which(flat_slope_preds$cat3 == "weather-Northern Mountains" & flat_slope_preds$cat2 == "extreme"), "y"] = 1.924664



#windspeed nm low
flat_slope_preds[which(flat_slope_preds$cat3 == "windspeed-Northern Mountains" & flat_slope_preds$cat2 == "low"), "x"] = 13
flat_slope_preds[which(flat_slope_preds$cat3 == "windspeed-Northern Mountains" & flat_slope_preds$cat2 == "low"), "y"] = 1.569492


#CBI SW mod
flat_slope_preds[which(flat_slope_preds$cat3 == "cbi-Southwest" & flat_slope_preds$cat2 == "mod"), "x"] = NA
flat_slope_preds[which(flat_slope_preds$cat3 == "cbi-Southwest" & flat_slope_preds$cat2 == "mod"), "y"] = NA




# flat_slope_preds2 = preds %>% filter(CI_low <= 0) %>%
#   mutate(weather = as.character(weather),
#          climate = as.character(climate)) %>% #convert from factor to remove NAs
#   mutate_if(is.character, ~replace_na(.,"")) %>%
#   mutate(cat2 = paste(climate, cbii_cat, weather, windspeed, sep = ""),
#          cat3 = paste(cat, ecoregion, sep = "-")) %>%
#   group_by(cat3, cat2)
#   # filter(!is.na(lag(sig))) %>%
#   # summarize(Extracted_Value = first(time_betwe_round, order_by = time_betwe_round))
# 
# 
# 
# inds = preds %>% filter(sig == "non-sig")
# df = preds 
```



#### Plot predictions


weather plots
plot interaction between weather and time between fires
```{r}
library(gridExtra)

predictor_plot_func = function(df_preds, df_raw, group_var, slope_cat){
      group_var = factor(group_var, levels = c("extreme", "very high", "high"))


  p = ggplot(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var)) +
         geom_line(size = 1.05)+
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    scale_fill_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
     ylim(0.1, 2.5) +
    # scale_x_continuous(breaks = c(5,10,15,20,25,30,35))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="weather"),
           color=guide_legend(title="weather"))
  
  p2 = p + geom_rug(data = df_raw, sides = "b",
                    aes(x = time_betwe*10), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat),
             aes(x=x,y=y, color = cat2), size = 3)
  
  return(p2)

}

(pca = predictor_plot_func(preds_w_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_w_ca_trunc$weather,
                          slope_cat = "weather-California Coast") +
  ggtitle("California Coast") +
  theme(legend.position = "none") )  

psw = predictor_plot_func(preds_w_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_w_sw_trunc$weather,
                          slope_cat = "weather-Southwest") +
  ggtitle("Southwest") + ylab("") +
  theme(legend.position = "none")    

pnm = predictor_plot_func(preds_w_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_w_nm_trunc$weather,
                          slope_cat = "weather-Northern Mountains") +
  ggtitle("Northern Mountains") + ylab("") + xlab("")

pwm = predictor_plot_func(preds_w_wm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_w_wm_trunc$weather,
                          slope_cat = "weather-Western Mountains") + 
  ggtitle("Western Mountains") + xlab("") +
  theme(legend.position = "none")    

#save plot
library(patchwork)
pwm + pnm + pca + psw + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/weather_newsev1k.png'), height = 5, width = 6.2)
```

windspeed plots
```{r}
predictor_plot_funcv = function(df_preds, df_raw, group_var, slope_cat){
    group_var = factor(group_var, levels = c("high", "moderate", "low"))

  p = ggplot() +
         geom_line(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var), size = 1.05)+
           geom_ribbon(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    scale_fill_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    ylim(0.1, 2.5) +
    # scale_x_continuous(breaks = c(5,10,15,20,25,30,35))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="wind speed"),
           color=guide_legend(title="wind speed"))
  
  p2 = p + geom_rug(data = df_raw, sides = "b",
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat),
             aes(x=x,y=y, color = cat2), size = 3)
  
  return(p2)

}

pca = predictor_plot_funcv(preds_vs_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_vs_ca_trunc$windspeed,
                          slope_cat = "windspeed-California Coast") + 
  ggtitle("California Coast") +
  theme(legend.position = "none")    

pnm = predictor_plot_funcv(preds_vs_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_vs_nm_trunc$windspeed,
                          slope_cat = "windspeed-Northern Mountains") + 
  ggtitle("Northern Mountains") + ylab("") + xlab("")
 

psw = predictor_plot_funcv(preds_vs_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_vs_sw_trunc$windspeed,
                          slope_cat = "windspeed-Southwest") + 
  ggtitle("Southwest") + ylab("") + theme(legend.position = "none")    


pwm = predictor_plot_funcv(preds_vs_wm_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_vs_wm_trunc$windspeed,
                          slope_cat = "windspeed-Western Mountains") + 
  ggtitle("Western Mountains") + xlab("") + theme(legend.position = "none")    


#plot with patchwork
pwm + pnm + pca + psw + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/windspeed_newsev1k.png'), height = 5, width = 6)
```

CBI plots - plot predictions with CBI set at 0.5 and 2
```{r}
predictor_plot_func2 = function(df_preds, df_raw, group_var, slope_cat){
    group_var = recode_factor(group_var, mod = "moderate",
                              low = "low") 
    
  p = ggplot() +
         geom_line(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var), size = 1.05)+
           geom_ribbon(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#FFBF69", "lightseagreen"))+
    scale_linetype_manual(values = c(2, 1))+
    scale_fill_manual(values = c("#FFBF69", "lightseagreen"))+
    ylim(0, 2.5) +
    # scale_x_continuous(breaks = c(5,10,15,20,25,30,35))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="initial severity"),
           color=guide_legend(title="initial severity"))+
      geom_hline(yintercept = 0.5, linetype = "longdash", color = "lightseagreen")+ 
      geom_hline(yintercept = 2, linetype = "longdash", color = "#FFBF69")
  
  p2 = p + geom_rug(data = df_raw, sides = "b",
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat) %>%
                   mutate(cat2 = recode_factor(cat2, mod = "moderate",
                                               low = "low")),
             aes(x=x,y=y, color = cat2), size = 3)
  
  return(p2)

}

pca = predictor_plot_func2(preds_cbi_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_cbi_ca_trunc$cbii_cat,
                          slope_cat = "cbi-California Coast") + 
  ggtitle("California Coast") +
  theme(legend.position = "none")    

pnm = predictor_plot_func2(preds_cbi_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_cbi_nm_trunc$cbii_cat,
                          slope_cat = "cbi-Northern Mountains") + 
  ggtitle("Northern Mountains") + ylab("") + xlab("")


psw = predictor_plot_func2(preds_cbi_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_cbi_sw_trunc$cbii_cat,
                          slope_cat = "cbi-Southwest") +
  ggtitle("Southwest") + ylab("") + theme(legend.position = "none") 


pwm = predictor_plot_func2(preds_cbi_wm_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_cbi_wm_trunc$cbii_cat,
                          slope_cat = "cbi-Western Mountains") +
    ggtitle("Western Mountains") + xlab("") + theme(legend.position = "none") 

#plot with patchwork
pwm + pnm + pca + psw + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/cbi_newsev1k.png'), height = 5, width = 6)
```

climate plots
plot interaction between climate/ productivity and time between fires
```{r}

predictor_plot_func3 = function(df_preds, df_raw, group_var, slope_cat){
    group_var = recode_factor(group_var,
                              hotdry= "hot-dry",
                              hotwet = "warm-wet",
                              coldwet = "cool-wet") 
    
  p = ggplot(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var)) +
         geom_line(size = 1.05)+
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    scale_fill_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    ylim(0.1, 2.5) +
    # scale_y_continuous(breaks = c(0.5, 1, 1.5, 2, 2.5))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="climate"),
           color=guide_legend(title="climate"))
  
  p2 = p + geom_rug(data = df_raw, sides = "b",
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat) %>%
                   mutate(cat2 = recode_factor(cat2,                            
                                  hotdry= "hot-dry",
                                  hotwet = "warm-wet",
                                  coldwet = "cool-wet")), 
             aes(x=x,y=y, color = cat2), size = 3)
  
  return(p2)

}

pca = predictor_plot_func3(preds_c_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_c_ca_trunc$climate,
                          slope_cat = "climate-California Coast") + 
  ggtitle("California Coast") + 
  theme(legend.position = "none")

pnm = predictor_plot_func3(preds_c_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_c_nm_trunc$climate,
                          slope_cat = "climate-Northern Mountains") + 
  ggtitle("Northern Mountains") + xlab("") + ylab("")


psw = predictor_plot_func3(preds_c_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_c_sw_trunc$climate,
                          slope_cat = "climate-Southwest") + 
  ggtitle("Southwest") + ylab("") +
  theme(legend.position = "none")  


pwm = predictor_plot_func3(preds_c_wm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_c_wm_trunc$climate,
                          slope_cat = "climate-Western Mountains") + 
  ggtitle("Western Mountains") + xlab("")   +
  theme(legend.position = "none")    


#plot with patchwork
pwm + pnm + pca + psw + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/climate_newsev1k.png'), height = 5, width = 6.1)
```
## Calculate strength of effect against 6-yr time between fires for 12, 18, and 24 years
```{r}
preds2 = preds %>% mutate(cat2 = str_c(str_replace_na(climate), str_replace_na(weather),str_replace_na(windspeed), str_replace_na(cbii_cat)),
                          cat2 = str_replace_all(cat2, "NA", "")) %>%
  filter(time_betwe_round == 0.6) %>%
  mutate(cat3 = paste(cat, ecoregion, sep = "-")) %>%
  full_join(flat_slope_preds2) %>%
  mutate(strength = y - cbi_reburn,
         perc_chang = strength/cbi_reburn*100,
         slope = strength/(x - time_betwe_round*10))

preds25 = preds %>% mutate(cat2 = str_c(str_replace_na(climate), str_replace_na(weather),str_replace_na(windspeed), str_replace_na(cbii_cat)),
                          cat2 = str_replace_all(cat2, "NA", "")) %>%
  filter(time_betwe_round == 2.5) %>%
  mutate(cat3 = paste(cat, ecoregion, sep = "-"),
         y25 = cbi_reburn) %>%
  dplyr::select(cat2, cat3, y25) %>%
  group_by(cat2, cat3) %>%
  slice_min(y25)

preds24 = preds %>% mutate(cat2 = str_c(str_replace_na(climate), str_replace_na(weather),str_replace_na(windspeed), str_replace_na(cbii_cat)),
                          cat2 = str_replace_all(cat2, "NA", "")) %>%
  filter(time_betwe_round == 2.4) %>%
  mutate(cat3 = paste(cat, ecoregion, sep = "-"),
         y24 = cbi_reburn) %>%
  dplyr::select(cat2, cat3, y24) %>%
  group_by(cat2, cat3) %>%
  slice_min(y24)

preds18 = preds %>% mutate(cat2 = str_c(str_replace_na(climate), str_replace_na(weather),str_replace_na(windspeed), str_replace_na(cbii_cat)),
                          cat2 = str_replace_all(cat2, "NA", "")) %>%
  filter(time_betwe_round == 1.8) %>%
  mutate(cat3 = paste(cat, ecoregion, sep = "-"),
         y18 = cbi_reburn) %>%
  dplyr::select(cat2, cat3, y18) %>%
  group_by(cat2, cat3) %>%
  slice_min(y18)

preds12 = preds %>% mutate(cat2 = str_c(str_replace_na(climate), str_replace_na(weather),str_replace_na(windspeed), str_replace_na(cbii_cat)),
                          cat2 = str_replace_all(cat2, "NA", "")) %>%
  filter(time_betwe_round == 1.2) %>%
  mutate(cat3 = paste(cat, ecoregion, sep = "-"),
         y12 = cbi_reburn) %>%
  dplyr::select(cat2, cat3, y12) %>%
  group_by(cat2, cat3) %>%
  slice_min(y12)

df_list = list(preds2, preds24, preds12, preds18)

 

preds_all = df_list %>% reduce(full_join) %>%
  mutate(strength24 = y24 - cbi_reburn,
         perc_chang24 = strength24/cbi_reburn*100,
         slope24 = strength24/(24 - time_betwe_round*10),
         
         strength18 = y18 - cbi_reburn,
         perc_chang18 = strength18/cbi_reburn*100,
         slope18 = strength18/(18 - time_betwe_round*10),
         
         strength12 = y12 - cbi_reburn,
         perc_chang12 = strength12/cbi_reburn*100,
         slope12 = strength12/(12 - time_betwe_round*10))
  # mutate(strength_all = coalesce(strength, strength25),
  #        perc_chang_all = coalesce(perc_chang, perc_chang25),
  #        slope_all = coalesce(slope, slope25))
  
(0.3405148-0.1817616)/0.1817616 *100
```
