---
title: "sample_points"
author: "CT"
date: "2022-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

data_dir = readLines(here("data_dir_D.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script generates and buffers sample points for the drivers of reburn severity project


## Generate sample points
generate sample points from fire history raster
```{r}
#read in rater with count of number of times each pixel burned
burns_r = raster(datadir("/prelim-outputs/fire_history/count_burns_int.tif"))

#generate 270m raster to snap to when resampling
r270 = aggregate(burns_r, fact=9, fun=max, na.rm = TRUE)

#resample to 270m
burns_r270 = raster::resample(burns_r, r270, method = "ngb", filename = datadir("/prelim-outputs/fire_history/count_burns_resamp270m.tif"))


#convert raster to points for only reburned areas (x > 1 time burned)
reburn_points = rasterToPoints(burns_r270, fun=function(x){x>1}, spatial = TRUE)

#write reburn points to shp
writeOGR(obj=reburn_points, dsn=datadir("/prelim-outputs/points"), layer="reburn_points_270test", driver="ESRI Shapefile")

# plot(reburn_points[1:100000,])
```

#### Buffer reburned area 
so that we don't have any sample points within 300m of edge to limit edge effects and spatial data inconsistencies

Buffer polygons inside 300m
```{r}
#get reburned area - did this in arc for faster processing: 
#> converted count_burns_int.tif to polygons %>% removed areas that only burned once

reburn_poly = st_read(datadir("/prelim-outputs/fire_history/reburn_poly.shp"))

# buffer size = 300m coordinates inward
buffer_size <- (-300)

# compute buffered polygons
perims_buffered <- st_buffer(reburn_poly, buffer_size) 

# remove degenerate polygons
perims_remaining <- perims_buffered[!st_is_empty(perims_buffered),]

#write perims_buffer
st_write(perims_remaining, datadir("prelim-outputs/buffered_perims/buffered_perims.shp"))
# perims_remaining = st_read(datadir("prelim-outputs/buffered_perims/buffered_perims.shp"))
```

Extract points that fall within the buffer to new points dataset
```{r}
reburn_points = st_read(datadir("prelim-outputs/points/reburn_points_270.shp"))

# plot(perims[2,1])
# plot(perims_remaining[2,1])

#keep only points that fall in buffered perims
points_buff = st_intersects(reburn_points, perims_remaining)
points_buff_include = lengths(points_buff) > 0 
points_buff_include = reburn_points[points_buff_include,]
  
  
#write sample points
st_write(points_buff_include, datadir("prelim-outputs/points/sample_points_buff.shp"))

```


