---
title: "Fire_severity_prep"
author: "CT"
date: "2022-10-03"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script preps severity CBI data from GEE for analysis by creating annual CBI mosiac raster layers 1985-2021

load data

```{r load data}
# import all files in a single folder as a list 
rastlist <- list.files(path = datadir("/raw-data/cbi-GEE"), pattern='.tif$', all.files=TRUE, full.names=TRUE)

included_perims = list.files(path = datadir("/raw-data/cbi-GEE"), pattern='.tif$', all.files=TRUE, full.names=FALSE) %>% 
  data.frame()

colnames(included_perims) = 'rasterID'

included_perims$Fire_ID = str_sub(included_perims$rasterID, 1, 21)
```

quality check cbi - locate missing perimeters
```{r}
#check which fire perims were not generated
perims = vect(datadir("/raw-data/fire-perims/mtbs_perimeter_data/mtbs_west_fs_nps_sub_updated21.shp"))

missing_perims = perims %>% 
  sf::st_as_sf() %>%
  st_transform(., 5070) %>%
  subset(., !(Fire_ID %in% included_perims$Fire_ID ))

#2021 fires still don't have imagery - 139 perims

missing_perims = missing_perims %>% 
  subset(Fire_Year < 2021)
#no more missing perims!

missing_perims = missing_perims %>% 
  subset(Fire_Year > 2020)
#got all 2021 fires!

```

write missing perimeter files for GEE - completed
```{r}
# #save as new file to upload to GEE and extract CBI
# missing_perims = vect(missing_perims)  
# 
# #write out 2021 fire perims for GEE
# writeVector(missing_perims, file = datadir("/raw-data/fire-perims/mtbs_perimeter_data/missing_perims.shp"), overwrite = TRUE)
```


# create raster layer for each fire year 1985-2021

```{r create folders}
# 
#   years <- as.character(1985:2021)
#   for(i in 1:length(years)){
#     folder_i <- paste(
#       here(),
#       "/data/raw-data/cbi-GEE/",
#       years[i],
#       sep = ""
#     )
#   
#     dir.create(folder_i)
#   }

```


Move rasters for each year into their respective folders
```{r move files}

#   tiff_list <- list.files(here("data/raw-data/cbi-GEE"))
# 
# # removing files using years vector from above chunk
#   tiff_list <- tiff_list[!(tiff_list %in% years)]
# 
# # get the years of the tiff list
#   yrs_tiffs <- str_sub(tiff_list, 14, 17)
# 
# # loop through years and move files
#   for(yr in years){
#   
#     tiff_ids_yr <- which(yrs_tiffs == yr)
#   
#     # from file path
#     fp_from <- paste(
#       here(),
#       "/data/raw-data/cbi-GEE/",
#       sep = ""
#     )
#   
#     # to the file
#     fp_to <- paste(
#       here(),
#       "/data/raw-data/cbi-GEE/",
#       yr, "/",
#       sep = ""
#     )
#   
#     # copy files
#     file.copy(
#       from = paste(fp_from, tiff_list[tiff_ids_yr], sep = ""),
#       to = paste(fp_to, tiff_list[tiff_ids_yr], sep = "")
#     )
#   }

```


create cbi raster for each fire year
```{r}
#get files
filelist <- list.files(path = datadir("/raw-data/cbi-GEE/2002"), pattern='.tif$', all.files=TRUE, full.names=TRUE)

#convert to raster
rast.list <- lapply(1:length(filelist), function(x) {raster(filelist[x])})

rast.list$fun <- max
tic()
rast.mosaic <- do.call(mosaic,rast.list) 
writeRaster(rast.mosaic, datadir("/raw-data/cbi-GEE/cbi_2002.tif"), overwrite = TRUE)
toc()
```

as a function
```{r}
#loop through all folders
years = as.character(1987:2021)

  for(yr in years){
    
    #get file list
    filelist <- list.files(paste(
      here(),
      "/data/raw-data/cbi-GEE/",
      yr, "/",
      sep = ""), 
      pattern='.tif$',
      all.files=TRUE, full.names=TRUE)
    
    #convert to raster
    rast.list <- lapply(1:length(filelist), function(x) {raster(filelist[x])})
    
    rast.list$fun <- max
    
    #mosiac raster
    tic()
    rast.mosaic <- do.call(mosaic,rast.list)
    
    #write raster 
    writeRaster(rast.mosaic, paste(datadir(""),"/raw-data/cbi-GEE/cbi_", yr, ".tif", sep = ""))
    toc()
    }

```


```{r}
#extrat all years cbi to all sample points
#calculate most recent fire and second most recent fire severity from each point using burn year

```



---------------------------------------------------------------------------------

Find most recently burned fires for extracting cbi to a reburn severity raster

This is does not work for extracting reburn rasters because many perimeters are involved in >2 fires

```{r Find most recently burned fires}
# perims = st_read(datadir("/raw-data/fire-perims/mtbs_perimeter_data/mtbs_west_fs_nps_sub_updated21.shp"))
# 
# 
# #extract the most recent fire from each intersection = the highest number in the list 
# perims_int = st_intersects(perims)
# 
# #check intersections - looks good.
# # plot(perims[c(4, 26,993,1169),])
# 
# perim_mostrecent = lapply(perims_int, max) %>% 
#   unlist() %>% 
#   unique() #remove duplicates
# 
# perim_mostrecent_poly = perims[(perims %in% perim_mostrecent),]
# 
# #extract second most recent fire from each intersection
# second_recent_func = function(x, na.rm) {
#   x = x[!is.na(x)]
#   if(length(x) == 0) { return(NA) }
#   x_nomax = x[x != max(x)]
#   if (length(x_nomax) == 0) { return(NA) }
#   second_recent =  max(x_nomax)
#   return(second_recent)
# }
# perim_secondrecent = lapply(perims_int, second_recent_func) %>% 
#   unlist() %>% 
#   unique()  #remove duplicates
# 
# #extract Fire_ID by index
# perim_mostrecent_polys = perims[c(perim_mostrecent),]
# perim_secondrecent_polys = perims[c(perim_secondrecent),]


```

