---
title: "Topography data prep"
author: "CT"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time
library(spatialEco)

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script preps landfire data for the reburn project


move temp files to external drive to save disk space

currently temp files saved to "C:\\Users\\tortorec\\AppData\\Local\\Temp\\RtmpQxwFZk"

```{r}
# write("TMP = 'D:/Rtemp'", file=file.path(Sys.getenv('R_USER'), '.Renviron'))
# tempdir()

```


### Calculate heat load 
Creates raster class object of McCune & Keon (2002) Heat Load Index - https://rdrr.io/cran/spatialEco/man/hli.html


read in cropped elevation raster
```{r}
#read in raster
elev = raster(datadir("/raw-data/Landfire/LC20_Elev_220_crop.tif"))

#set values <0 to NA to reduce file size for faster processing
elevNA = reclassify(elev, cbind(-Inf, 0, NA), right=FALSE)
dataType(elevNA) <- 'INT2S'  

writeRaster(elevNA, datadir("/raw-data/Landfire/LC20_Elev_220_crop_NA.tif"), overwrite = TRUE)
```

calcualte heat load
```{r}

heat.load <- hli(elev)

writeRaster(heat.load, datadir("/raw-data/Landfire/heat_load.tif"), overwrite = TRUE)
```


Remove temp files to clean hard drive

https://stackoverflow.com/questions/25426405/raster-package-taking-all-hard-drive
```{r}
dirname(rasterTmpFile()) 

#remove temp files older than 30 mins
removeTmpFiles(0.5)
```

### Calculate topographic position index - 
following Cansler et al 2022
https://rdrr.io/cran/spatialEco/man/tpi.html

-this isn't working properly, trying tiling the raster first. -problem was with the DEM. using the original cropepd DEM

Decided to go with 500m TPI to reduce processing time and because Cansler et al 2022 found them to be comparable for predciting reburn severity
```{r}
#spatialEco
# testDEM = raster(datadir("/raw-data/Landfire/test_dem_clip.tif"))

# library(meteo)
elevNA = raster(datadir("/raw-data/Landfire/LC20_Elev_220_crop.tif"))
#generate 30km tiles with 600m buffer for DEM
t = tiling(elevNA, tilesize = 30000, overlapping = 600, asfiles = T, tilename = "DEM_tile_", tiles_folder = datadir("raw-data/Landfire/DEM_tiles/"))
```


```{r}
# tic()
# tpi500 = tpi(elevNA, scale = 500, win = "circle")
# 
# writeRaster(tpi500, datadir("prelim-outputs/topography/TPI_1024.tif"), overwrite = TRUE)
# toc()

tic()
tpi2000 = tpi(elevNA, scale = 2000, win = "circle")

writeRaster(tpi2000, datadir("prelim-outputs/topography/TPI2000_1024.tif"), overwrite = TRUE)
toc()
```
