---
title: "Topography data prep"
author: "CT"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time
library(spatialEco)

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script preps landfire data for the reburn project


move temp files to external drive to save disk space

currently temp files saved to "C:\\Users\\tortorec\\AppData\\Local\\Temp\\RtmpQxwFZk"

```{r}
# write("TMP = 'D:/Rtemp'", file=file.path(Sys.getenv('R_USER'), '.Renviron'))
# tempdir()

```


### Calculate heat load 
Creates raster class object of McCune & Keon (2002) Heat Load Index - https://rdrr.io/cran/spatialEco/man/hli.html


read in cropped elevation raster
```{r}
#read in raster
elev = raster(datadir("/raw-data/Landfire/LC20_Elev_220_crop.tif"))

#set values <0 to NA to reduce file size for faster processing
elevNA = reclassify(elev, cbind(-Inf, 0, NA), right=FALSE)
dataType(elevNA) <- 'INT2S'  

writeRaster(elevNA, datadir("/raw-data/Landfire/LC20_Elev_220_crop_NA.tif"), overwrite = TRUE)
```

calcualte heat load
```{r}

heat.load <- hli(elev)

writeRaster(heat.load, datadir("/raw-data/Landfire/heat_load.tif"), overwrite = TRUE)
```


Remove temp files to clean hard drive

https://stackoverflow.com/questions/25426405/raster-package-taking-all-hard-drive
```{r}
dirname(rasterTmpFile()) 

#remove temp files older than 30 mins
removeTmpFiles(0.5)
```

Calculate topographic position index for 2000m scale - following Cansler et al 2022
https://rdrr.io/cran/spatialEco/man/tpi.html
```{r}
#spatialEco
tpi = tpi(elevNA, scale = 2000, win = "circle")

writeRaster(tpi, datadir("/raw-data/prelim-outputs/topography/tpi.tif"), overwrite = TRUE)
```
