---
title: "reburn-mtbs-data-prep"
output: html_document
date: '2022-06-03'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(here)
library(rgdal)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script preps MTBS data for the CA reburn analysis

# Organize MTBS data

load mtbs data

```{r}
#load mtbs data
mtbs_perims = vect(datadir("/raw-data/fire-perims/mtbs_perimeter_data/mtbs_perims_DD.shp")) #Mercator (EPSG: 4269)

#add fire year
mtbs_perims$YEAR = as.numeric(substr(as.character(mtbs_perims$Ig_Date), 1,4)) 
#DY: This line isn't necessary since the correct CRS is stored in the shapefile: crs(mtbs_perims) = 'epsg:4269'
```

### Subset analysis for faster processing - CA

```{r}

#extract states of interest
#load state boundaries
states = vect(datadir("/raw-data/geography/cb_2018_us_state_20m/cb_2018_us_state_20m.shp")) 

states = project(states, mtbs_perims)

#intersect states and mtbs perims - this is kind of slow
mtbs_states = intersect(mtbs_perims, states)

#extract just CA for initial analysis
# states = states[, "NAME"] #select name column
mtbs_ca = mtbs_states[mtbs_states$NAME == "California",]


#reproject to UTMs Zone 10 for units = m
newcrs = "epsg:26910"
mtbs_ca = project(mtbs_ca, newcrs)

```

Subset dataset further for processing
```{r}
#grid out CA and subset
#DY: This tiling approach looks great. We should keep an eye out for possible cases where edge effects might be an issue, and if they exist, add a buffer to the grid.

z <- rast(mtbs_ca)
dim(z) <- c(5,5)
values(z) <- 1:25
names(z) <- 'Zone'
# coerce SpatRaster to SpatVector polygons
z <- as.polygons(z)

# add 300m buffer around zones
bz <- buffer(z, width=300) 

# plot(bz)
# plot(mtbs_ca, add = TRUE)


#extract mtbs data from one zone for faster processing
mtbs_ca1 = intersect(mtbs_ca, z[1,])


```


----
## Mapping reburned areas


Rasterize mtbs for easy extraction
```{r}
# Rasterize mtbs for easy extraction.  - need to use project coordinate system - like Albers equal area conical 
# Equal Area Conus Albers. - epsg:5070


r = rast(mtbs_ca1, resolution = 30, crs = "epsg:26910")

```

Extract data on the most recent and second most recent fires
```{r}

tic()
mostrecent_fire = raster::rasterize(mtbs_ca1 %>% as("Spatial"), r %>% raster(), field="YEAR", fun=max)
toc()

#function to extract second most recent fire 
second_recent_func = function(x, na.rm) {
  x = x[!is.na(x)]
  if(length(x) == 0) { return(NA) }
  x_nomax = x[x != max(x)]
  if (length(x_nomax) == 0) { return(NA) }
  second_recent =  max(x_nomax)
  return(second_recent)
}

tic()
secondrecent_fire = raster::rasterize(mtbs_ca1 %>% as("Spatial"), r %>% raster, field="YEAR", fun=second_recent_func)
toc()
#~2.3 mins at 30m res

```

Create grid of points for extracting data to 
*this is really slow at 30m resolution!!
```{r}
## create a grid of points
tic()
grdpts = sf::st_make_grid(mtbs_ca1, cellsize = 30, what = "centers")
toc()

## convert it to an `sf` object, as opposed to an `sfc`
grdpts = sf::st_sf(grdpts)
grdpts = vect(grdpts)

# plot(grdpts)
# plot(mtbs_ca1, add = TRUE)

#add identifier
grdpts$ID = 1:length(grdpts)
```

extract mtbs data to points
```{r}
mtbs_pnts = intersect(mtbs_ca1, grdpts)

```