---
title: "reburn-mtbs-data-prep"
output: html_document
date: '2022-06-03'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script preps MTBS data for the CA reburn analysis

# Organize MTBS data

load mtbs data

```{r load data}
#load mtbs data
mtbs_perims = vect(datadir("/raw-data/fire-perims/mtbs_perimeter_data/mtbs_perims_DD.shp")) #Mercator (EPSG: 4269)

#add fire year
mtbs_perims$YEAR = as.numeric(substr(as.character(mtbs_perims$Ig_Date), 1,4)) 
#DY: This line isn't necessary since the correct CRS is stored in the shapefile: crs(mtbs_perims) = 'epsg:4269'
```

### Subset data for faster processing - CA

Subset by state
```{r subset data  by state}

#extract states of interest
#load state boundaries
states = vect(datadir("/raw-data/geography/cb_2018_us_state_20m/cb_2018_us_state_20m.shp")) 

states = project(states, mtbs_perims)

#intersect states and mtbs perims - this takes a few seconds
mtbs_states = terra::intersect(mtbs_perims, states)

#extract just CA for initial analysis
# states = states[, "NAME"] #select name column
mtbs_ca = mtbs_states[mtbs_states$NAME == "California",]


#reproject to UTMs Zone 10 for units = m
newcrs = "epsg:26910"
mtbs_ca = project(mtbs_ca, newcrs)

```



Prep perims for GEE cbi extraction and subset by year
```{r gee prep}
#grid out CA and subset
#DY: This tiling approach looks great. We should keep an eye out for possible cases where edge effects might be an issue, and if they exist, add a buffer to the grid.

#rename mtbs cols for GEE - following Parks et al 2019 GEE code for calculating CBI
# //       Fire_ID        unique identifier for each fire
# //       Fire_Year      year of fire
# //       Start_Day      start day of fire season in julian days, e.g. June 15 = 166
# //       End_Day        end day of fire season in julian days

mtbs_ca$Fire_ID = mtbs_ca$Event_ID
mtbs_ca$Fire_Year = as.integer(mtbs_ca$YEAR)

mtbs_ca$Start_Day = yday(mtbs_ca$Ig_Date) #Q: not sure if this should be actual start date or average for CA?
mtbs_ca$End_Day = 273

mtbs_ca = mtbs_ca[,c("Fire_ID", "Fire_Year", "Start_Day", "End_Day")]

#remove fires from 1984 because we don't have pre-fire imagery
mtbs_ca_85 = mtbs_ca[which(mtbs_ca$Fire_Year > 1985),]
```

tile mtbs data for faster processing
```{r tiling}
#larger grid for GEE
z <- rast(mtbs_ca_85)

dim(z) <- c(5,5)
values(z) <- 1:25
names(z) <- 'Zone'
z <- as.polygons(z) # coerce SpatRaster to SpatVector polygons
bz <- buffer(z, width=300) # add 300m buffer around zones

#create a list of zone names
zones <- paste("mtbs_ca_z",1:25, sep = "")
zone_list <- vector(mode="list", length = 25)
names(zone_list) = zones

#create mtbs variable for each zone with 300m buffer
for (i in 1:length(bz)){
  zone_list[[i]] <- terra::intersect(mtbs_ca_85, bz[i,])
}

#remove mtbs zones with no fires leaving 17 zones to run
zone_list <- zone_list[sapply(zone_list, function(x) length(x) > 0)]

```

save shapefiles
```{r save shps, eval = FALSE}
#save mtbs zones as shapefiles
lapply(1:length(zone_list), function(i) writeVector(zone_list[[i]], 
                                      file = datadir( paste0("/raw-data/fire-perims/mtbs_perimeter_data/mtbs_ca_zones/", 
                                                             names(zone_list[i]), ".shp")),
                                      overwrite = TRUE))
plot(zone_list[[1])
```

----
## Mapping reburned areas


Rasterize mtbs for easy extraction
```{r}
# Rasterize mtbs for easy extraction.  - need to use project coordinate system - like Albers equal area conical 
# Equal Area Conus Albers. - epsg:5070


r = rast(zone_list[[1]], resolution = 30, crs = "epsg:26910")

```

Extract data on the most recent and second most recent fires
```{r extract most recent fires}

tic()
mostrecent_fire = raster::rasterize(zone_list[[1]] %>% as("Spatial"), r %>% raster(), field="Fire_Year", fun=max)
toc()

#function to extract second most recent fire 
second_recent_func = function(x, na.rm) {
  x = x[!is.na(x)]
  if(length(x) == 0) { return(NA) }
  x_nomax = x[x != max(x)]
  if (length(x_nomax) == 0) { return(NA) }
  second_recent =  max(x_nomax)
  return(second_recent)
}

tic()
secondrecent_fire = raster::rasterize(zone_list[[1]] %>% as("Spatial"), r %>% raster, field="Fire_Year", fun=second_recent_func)
toc()
#~2.3 mins at 30m res

```

---------------------------------------------------------------------------------
# Subset by FS land

remove fires that did not burn on FS land (keep perims with >=50 ha overlap)
```{r FS land, eval = FALSE}
#load ca fs land spatial data
fs = vect(datadir("/raw-data/ownership/fsland_ca.gpkg")) 
# plot(fs)
fs = project(fs, newcrs) #project to UTMs

fs_sf = sf::st_as_sf(fs)
mtbs_ca_sf = sf::st_as_sf(mtbs_ca)


# intersect - note that sf is intelligent with attribute data!
fs_mtbs_int <- st_intersection(mtbs_ca_sf, fs_sf)
# plot(mtbs_ca_sf$geometry, axes = TRUE)
# plot(fs_mtbs_int$geometry, add = TRUE, col = 'red')

# add in areas in m2
fs_area <- fs_mtbs_int %>% 
  mutate(area = st_area(.) %>% as.numeric())

# for each perim, get area overlap with forest service land, then remove areas with <50 ha overlap
fs_area = fs_area %>% 
  as_tibble() %>% 
  group_by(Event_ID, OWNERCLASSIFICATION) %>% 
  summarize(area = sum(area)) %>%
  filter(area >= 500000)

#select fire perims that >= 50ha identified in fs_area
fs_mtbs_ca = mtbs_ca[mtbs_ca$Event_ID %in% fs_area$Event_ID,]

#some summary stats/data vis
# fs_mtbs_ca %>% data.frame() %>% 
#   group_by(YEAR) %>% 
#   summarise(acres_yr = sum(BurnBndAc)) %>% 
#   plot(acres_yr ~ YEAR)
```

------------------------------------------------------------------------------------
# point method 

Create grid of points for extracting data to 
*this is really slow at 30m resolution!!
```{r point grid, eval = FALSE}
## create a grid of points
tic()
grdpts = sf::st_make_grid(mtbs_ca1, cellsize = 30, what = "centers")
toc()

## convert it to an `sf` object, as opposed to an `sfc`
grdpts = sf::st_sf(grdpts)
grdpts = vect(grdpts)

# plot(grdpts)
# plot(mtbs_ca1, add = TRUE)

#add identifier
grdpts$ID = 1:length(grdpts)
```

extract mtbs data to points
```{r}
mtbs_pnts = intersect(mtbs_ca1, grdpts)

```