---
title: "RAP-data-prep"
author: "CT"
date: '2022-07-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(here)
library(rgdal)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

# remotes::install_github("brownag/rapr")
library(rapr)

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))
```

obtain RAP grids for the corresponding extent
```{r}
#load mtbs zones for extracting rap grids


#get yearly rap for CA
res <- get_rap(
  ca, 
  version = "v3",
  year = 2020,
  progress = FALSE
)


```

To cite rapr in publications use:
#>
#>   Andrew Brown (2022). rapr: Interface to Rangeland Analysis Platform
#>   (RAP) Vegetation Biomass and Cover Products. R package version 0.1.0.
#>   http://github.com/brownag/rapr
#>
#>   Jones, M.O., N.P. Robinson, D.E. Naugle, J.D. Maestas, M.C. Reeves,
#>   R.W. Lankston, and B.W. Allred. 2021. Annual and 16-Day Rangeland
#>   Production Estimates for the Western United States. Rangeland Ecology
#>   & Management 77:112-117. http://dx.doi.org/10.1016/j.rama.2021.04.003
#>
#>   Robinson, N. P., M. O. Jones, A. Moreno, T. A. Erickson, D. E.
#>   Naugle, and B. W. Allred. 2019. Rangeland productivity partitioned to
#>   sub-pixel plant functional types. Remote Sensing 11:1427.
#>   http://dx.doi.org/10.3390/rs11121427
#>
#>   Allred, B. W., B. T. Bestelmeyer, C. S. Boyd, C. Brown, K. W. Davies,
#>   L. M. Ellsworth, T. A. Erickson, S. D. Fuhlendorf, T. V. Griffiths,
#>   V. Jansen, M. O. Jones, J. Karl, J. D. Maestas, J. J. Maynard, S. E.
#>   McCord, D. E. Naugle, H. D. Starns, D. Twidwell, and D. R. Uden.
#>   2021. Improving Landsat predictions of rangeland fractional cover
#>   with multitask learning and uncertainty. Methods in Ecology and
#>   Evolution. http://dx.doi.org/10.1111/2041-210x.13564
#>
#> To see these entries in BibTeX format, use 'print(<citation>,
#> bibtex=TRUE)', 'toBibtex(.)', or set
#> 'options(citation.bibtex.max=999)'.