---
title: "Reburn analysis"
author: "CT"
date: "2022-11-30"
output: html_document
---


This script cleans and analyzes data for reburn analyses.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)





library(randomForest)
library(ranger) #a faster version of the RF algorithm
library(pdp)# for partial dependence plots (PDPs)

library(tidyverse)
library(here)
library(sp)
library(sf)
library(terra)
library(raster)

library(mgcv)
library(tidymv)

data_dir = readLines(here("data_dir_D.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

Load reburn points for analysis

```{r}
data = st_read(datadir("prelim-outputs/points/points_for_analysis_1-13.3-23.gpkg"))  

#add lat long coordinates to data frame
data_wgs = st_transform(data, 4326)
coords = data.frame(st_coordinates(data_wgs))

data$lat = coords$Y
data$long = coords$X
```

### Prep data for spatial partioning into test and training datasets 
> backwards selection
> mumin
> RF for important predictors, spatial zone splitting? 

Assign each point to spatial sampling zone 12km
```{r}
#determine data extent
extent(data)
# plot(extent)
e = ext(c(-2280909, -626619.3, 1007070, 3117660)) #xmin, xmax, ymin, ymax extent of data
p = as.polygons(e) #covert to polygons
crs(p) = crs(data) #set crs

r = rast(p) #convert raster

zones = rast(p, resolution = c(12000, 12000))
values(zones) = 1:ncell(zones)

# plot(zones)

#extract zone value to points
data$sample_zones = extract(zones, data, ID=FALSE)

#get rid of ID var in sample zones 
data = data %>% mutate(sample_zones = sample_zones$lyr.1) 

```

# Select points for analysis based on following criteria:

- Not on FS land (focus analysis on forested FS land, and for management)
- Less than 25% tree cover at time of reburn (focus analysis on forested areas)
- second fire occurred before 2002 (limited by MODIS for day of fire weather)
- burned at <0.5 CBI in initial or reburn fires (remove unburned areas to limit residual error where fire did not make it to a point within a fire perimeter)

```{r}

#remove points that... 
data_sub = data %>% 
  filter(tree_reburn >= 40 & #have greater than 25% tree cover at time of reburn
           mostrecent > 2001 & #second fire occurred after 2001 (for MODIS weather data)
           cbi_initial > 0.5 & #burned at >0.5 CBI in initial or reburn fires (remove unburned areas to limit residual error
           cni85_21 > 0.5) %>%
  filter(!is.na(ecoregion)) #remove points that fall outside of our 4 ecoregions


# st_write(data_sub, datadir("/prelim-outputs/points/data_sub_1-12-23.gpkg"))
```

add in erc percentiles
```{r}
#add missing data 
#add weather percentiles for erc, vs, and vpd 
perc_w = read_sf(datadir("prelim-outputs/points/dob_percentiles_1-12-23.gpkg"))

data_sub1 = st_join(data_sub, perc_w[,c("erc_perc","vpd_perc","vs_perc")])
```

clean data
```{r}
#clean data
data_sub2 = data_sub1 %>% 
  rename(cbi_reburn = cni85_21) %>% data.frame() %>%
  #remove rows where ndvi = NA (29 rows) 
  drop_na(ndvi) %>%
 
  #lump wildfire codes
  mutate(reburn_fire_type = recode(reburn_fire_type, 'Unknown' = 'Wildfire', 'Wildfire Daily Fire Perimeter' = 'Wildfire', 'Wildfire Final Fire Perimeter' = 'Wildfire', 'Wildland Fire Use' = 'Wildfire'),
  initial_fire_type = recode(initial_fire_type, 'Unknown' = 'Wildfire', 'Wildfire Daily Fire Perimeter' = 'Wildfire', 'Wildfire Final Fire Perimeter' = 'Wildfire', 'Wildland Fire Use' = 'Wildfire'),
  ecoregion = as.factor(ecoregion)) %>%
  
  #remove 8 points with no weather data 
  filter(fm100_02_2 >0 & th_02_21 > 0 & vpd_02_21 > 0 &  fm1000_02_ > 0 & vs_02_21 > 0 &  erc_02_21 > 0) %>%
  
  #remove 122 points without AET data
  filter(AET > 0) %>%
  
  #set point id to row names 
  column_to_rownames(., 'pointid')
  

#clean up facts management data
data_facts = data_sub2 %>%
    #lump facts management into one variable (time since management)
  mutate(time2manage1 = pmin(pfire_yrs2reburn, mechanical_yrs2reburn, na.rm = TRUE),
  #set NAs to time since last fire
        time2manage = coalesce(time2manage1, time_betwe),
  #add logical variable for if a point was managed within 10 years of the reburn
        postmanage10 = time2manage1 <= 10,
        postmanage10 = coalesce(postmanage10, FALSE),
    #add logical variable for if a point was managed within 10 years of the initial fire
        time2manage_preburn = pmin(pfire_yrs2initial, mechanical_yrs2initial, na.rm = TRUE),
        premanage10 = time2manage_preburn <= 10,
        premanage10 = coalesce(premanage10, FALSE)) #set NBAs to FALSE


  
data_clean = data_facts %>%
  #rescale time between to decades
  mutate(time_betwe = time_betwe/10) %>%
  #remove unnecessary columns
  dplyr::select(-c(LC20_Elev_, geom, grid_code, secondrece, max_yr_mechanical, max_yr_preburn_mechanical, max_yr_pfire, max_yr_preburn_pfire, pfire_yrs2reburn, pfire_yrs2initial, count_trt, count_trt_pfire, count_trt_mechanical, mechanical_yrs2reburn, mechanical_yrs2initial, count_trt_preburn, time2manage1, time2manage_preburn)) 

#write out cleaned_data
# st_write(data_clean,datadir("prelim-outputs/points/data_clean_for_analaysis.gpkg"))

#remove rows that were managed between initial and reburn (save these for a separate analyis)
data_nomanage = data_clean %>% filter(postmanage10 == FALSE)

#create binary category for high severity reburn
data_binom = data_nomanage %>% 
  mutate(burn_cat = case_when(cbi_reburn >= 2.5 ~ 1, 
                             TRUE ~ 0))


# #subsample large fires (max points per fire = 1000)
# data_subsample1 = data_nomanage %>%
#   group_by(reburn_fireID) %>%
#   mutate(count_points_reburn = n()) %>%
#   filter(count_points_reburn > 1128) %>% #median count
#   slice_sample(n = 1128)
# 
# # summary(data_subsample1$count_points_reburn)
# 
# data_subsample2 = data_nomanage %>%
#   group_by(reburn_fireID) %>%
#   mutate(count_points_reburn = n()) %>%
#   filter(count_points_reburn <= 1128)
# 
# data_subsample = bind_rows(data_subsample1, data_subsample2)
  

#break into low, moderate, and high initial fire severity categories
data_low = data_binom %>%
  filter(cbi_initial < 1.5)

data_mod = data_binom %>%
  filter(cbi_initial >= 1.5 & cbi_initial < 2.5)

data_high = data_binom %>% 
  filter(cbi_initial >= 2.5)

data_lowmod = data_binom %>%
  filter(cbi_initial < 2.5)
```


### Fit separate models for each ecoregion and scale predictors

```{r}
#scale numeric predictors for model fit - don't scale time between t or cbi or erc percentiles to maintain interpretation
allpreds = c("heat_load", "cbi_reburn", "count_burn", "time_betwe", "PRISM_tmea", "PRISM_ppt_", "TPI_1024",
                "LC16_BPS_2", "fm100_02_2", "th_02_21", "vpd_02_21", "vs_02_21",
                "erc_02_21", "PRISM_vpdm",  "AET", "fm1000_02_", "ndvi", "shrub_rebrn",
                "tree_reburn", "afg_rebrn", "pfg_rebrn", "cbi_initial", "ecoregion",
                "count_trt_preburn_mechanical", "count_trt_preburn_pfire", "reburn_fire_type",
                "reburn_fireID", "initial_fire_type", "initial_fireID", "lat",
                "long",  "sample_zones", "erc_perc", "vpd_perc",  "vs_perc", "time2manage", 
                "postmanage10", "premanage10")

preds_not2scale = c( "burn_cat", "cbi_reburn", "time_betwe", "ecoregion", 
                     "erc_perc", "vpd_perc",  "vs_perc",
                "count_trt_preburn_mechanical", "count_trt_preburn_pfire", "reburn_fire_type",
                "reburn_fireID", "initial_fire_type", "initial_fireID", "lat",
                "long",  "sample_zones", "time2manage", 
                "postmanage10", "premanage10")

scale_df = function(df){
  data_scale = scale(df[!(names(df) %in% preds_not2scale)], scale = TRUE, center = TRUE)
  data_scale_all = cbind(data_scale, df[names(df) %in% preds_not2scale])
  return(data_scale_all)
}

#create new dfs for each ecoregion and scale each ecoregion separately 
data_lowmod_ca = data_lowmod %>% filter(ecoregion == "California Coast") %>%
  scale_df()

data_lowmod_nm = data_lowmod %>% filter(ecoregion == "Northern Mountains") %>%
  scale_df()

data_lowmod_sw = data_lowmod %>% filter(ecoregion == "Southwest") %>%
  scale_df()

data_lowmod_wm = data_lowmod %>% filter(ecoregion == "Western Mountains") %>%
  scale_df()

```

Test for spatial autocorrelation
```{r}
library(ape)
inv_dist = with(data_lowmod_sw, 1/dist(cbind(lat, long), diag = TRUE, upper = TRUE))
inv_dist = as.matrix(inv_dist)
ape::Moran.I(data_lowmod_sw$cbi_reburn, weight = inv_dist, scaled = TRUE)

#observed = 0.16  - CA; 0.14 - NM; 0.23 for SW
# While statistically significant, there actually isn’t too much going on, though it may be enough to warrant dealing with in some fashion.

```


Model fit/selection with mgcv 
>https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/gam.selection.html 
>https://stats.stackexchange.com/questions/519433/gam-and-multiple-continuous-continuous-interactions-tensor-smooths #helpful for interpretting and fitting interactions (ti vs te vs s)

>all about gams: https://m-clark.github.io/generalized-additive-models/application.html#gam-1

```{r}

gm_binom = function(df, k){
  
  #fit gam - removed fm100 & fm1000, reburn shrub cover, ndvi, and management 
  gamfit = bam(burn_cat ~ count_burn + 
                  # s(AET, k = k, bs = "cs") +
                  # s(count_points_reburn, k = k, bs = "cs")+
                  s(PRISM_ppt_, k = k, bs = "cs") +
                  s(PRISM_tmea, k = k, bs = "cs") +
                  # s(PRISM_vpdm, k = k, bs = "cs") +
                  s(cbi_initial, k = k, bs = "cs") +
                  s(time_betwe, k = k, bs = "cs") +
                  s(erc_perc, k = k, bs = "cs") +
                  s(vs_perc, k = k, bs = "cs") +
                  s(vpd_perc, k = k, bs = "cs") +
                  # s(afg_rebrn, k = 5, bs = "cs") +
                  # s(pfg_rebrn, k = 5, bs = "cs") +
                  s(heat_load, k = k, bs = "cs") +
                  s(TPI_1024, k = k, bs = "cs") +
                  # s(ndvi, k = k, bs = "cs") +
                  # s(tree_reburn, k = k, bs = "cs") +
                  # ti(time_betwe, by = AET, k = k, bs = "cs") +
                  ti(time_betwe, by = PRISM_ppt_, k = k, bs = "cs") +
                  ti(time_betwe, by = PRISM_tmea, k = k, bs = "cs") +
                  ti(time_betwe, by = vpd_perc, k = k, bs = "cs") +
                  ti(time_betwe, by = vs_perc, k = k, bs = "cs") +
                  ti(time_betwe, by = erc_perc, k = k, bs = "cs") +
                 # ti(time_betwe, by = count_points_reburn, k = k, bs = "cs") +
                 #account for spatial autocorrelation
                 s(long, lat, bs = 'gp', k = 100, m = 2), #Using the Gaussian process smooth produces a result that is akin to kriging. There are many other features to play with, as well as other bases that would be applicable
                 # correlation = corSpatial(form = ~ lon + lat, type = 'gaussian'),
                data = df,
             select = T,
             method = "fREML", family = binomial("logit")) 

  return(gamfit)

}
```

Examine model fits
```{r}

#not aet, includes prism temp/precip
fit3_ca = gm_binom(data_lowmod_ca, k = 4)
fit3_nm = gm_binom(data_lowmod_nm, k = 4)
fit3_sw = gm_binom(data_lowmod_sw, k = 4)
fit3_wm = gm_binom(data_lowmod_wm, k = 4)

```




### Predict


```{r}
#prediction df with most variables set to mean = 0
preds_dfw_func = function(df, fit){
  df_w97 = data.frame(count_burn = 0, 
                       # AET = 0, 
                       PRISM_ppt_ = 0, PRISM_tmea = 0,
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                      # afg_rebrn = 0, pfg_rebrn = 0,
                       #set weather variables to 3rd quantile value by ecoregion
                       vpd_perc = 0.97,
                       erc_perc = 0.97,
                       # vs_02_21 = quantile(df$vs_02_21, 0.95),
                       vs_perc = .50,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1),
                       weather = "97perc",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  df_w90 = data.frame(count_burn = 0, 
                       # AET = 0, 
                       PRISM_ppt_ = 0, PRISM_tmea = 0,
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                      # afg_rebrn = 0, pfg_rebrn = 0,
                       #set weather variables to mean 
                       vpd_perc = .90, erc_perc = .90, 
                       vs_perc = .50,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1),
                      weather = "90perc",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  df_w80 = data.frame(count_burn = 0, 
                       # AET = 0, 
                       PRISM_ppt_ = 0, PRISM_tmea = 0,
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                      # afg_rebrn = 0, pfg_rebrn = 0,
                       #set weather variables to 1st quantile value  
                       vpd_perc = .8,
                       erc_perc = .8,
                       # vs_perc = quantile(df$vs_perc, 0.25),
                       vs_perc = 0.5,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1), 
                      weather = "80perc",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  
  data_w = rbind(df_w97, df_w80, df_w90)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE, type="response")
  
  #add predictions to df for plotting
  data_w$high_sev_prob = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  return(data_w)
}

preds_w_ca <- preds_dfw_func(df = data_lowmod_ca, fit = fit3_ca)
preds_w_nm <- preds_dfw_func(df = data_lowmod_nm, fit = fit3_nm)
preds_w_sw <- preds_dfw_func(df = data_lowmod_sw, fit = fit3_sw)
preds_w_wm <- preds_dfw_func(df = data_lowmod_wm, fit = fit3_wm)
```

climate predictions
```{r}
# #find west wide climate quantiles
# precip75 = quantile(data_lowmod$PRISM_ppt_, 0.75)


#prediction df with variables set to mean = 0
preds_dfc_func = function(df, fit){
  #cool wet - high productivity
  df_coolwet = data.frame(count_burn = 0, 
                        #set climate vars
                       # AET = quantile(df$AET, 0.75), 
                       PRISM_ppt_ = quantile(df$PRISM_ppt, 0.75),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.25),
                        #control for other vars - mean
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc = .9, erc_perc = .9,vs_perc = .5,
                       # afg_rebrn = 0, pfg_rebrn = 0,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1),
                       climate = "cool/wet",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  #all vars at mean
  df_mod = data.frame(count_burn = 0, 
                       # AET = 0, 
                       PRISM_ppt_ = 0, PRISM_tmea = 0,
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       #set weather variables to mean 
                       vpd_perc = .9, erc_perc = .9,vs_perc = .5,
                      # afg_rebrn = 0, pfg_rebrn = 0,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1),
                      climate = "mean",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  
  #hot dry - low productivity
  df_hotdry = data.frame(count_burn = 0, 
                        #set climate vars
                       # AET = quantile(df$AET, 0.25), 
                       PRISM_ppt_ = quantile(df$PRISM_ppt, 0.25),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.75),
                        #control for other vars - mean
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc = .9, erc_perc = .9,vs_perc = .5,
                       # afg_rebrn = 0, pfg_rebrn = 0,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1),
                       climate = "hot/dry",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  
    #hot dry - low productivity
  df_warmwet = data.frame(count_burn = 0, 
                        #set climate vars
                       # AET = quantile(df$AET, 0.25), 
                       PRISM_ppt_ = quantile(df$PRISM_ppt, 0.75),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.50),
                        #control for other vars - mean
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc = .9, erc_perc = .9,vs_perc = .5,
                       # afg_rebrn = 0, pfg_rebrn = 0,
                       #set time between fires to full range 1-25 years
                       time_betwe = seq(0.1, 3.6, by = .1),
                       climate = "warm/wet",
                       #control for lat long
                       lat = mean(df$lat),
                       long = mean(df$long))
  
  
  data_c = rbind(df_coolwet, df_hotdry, df_warmwet, df_mod)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_c, se.fit = TRUE, type="response")
  
  #add predictions to df for plotting
  data_c$high_sev_prob = p$fit
  data_c$upr = p$fit + (2 * p$se.fit)
  data_c$lwr = p$fit - (2 * p$se.fit)
  return(data_c)
}

preds_c_ca <- preds_dfc_func(df = data_lowmod_ca, fit = fit3_ca)
preds_c_nm <- preds_dfc_func(df = data_lowmod_nm, fit = fit3_nm)
preds_c_sw <- preds_dfc_func(df = data_lowmod_sw, fit = fit3_sw)
preds_c_wm <- preds_dfc_func(df = data_lowmod_wm, fit = fit3_wm)
```

#### Plot predictions

plot interaction between weather and time between fires
```{r}
library(gridExtra)

pca = ggplot(data = preds_w_ca, aes(x = time_betwe, y = high_sev_prob, color = weather)) +
         geom_line(size = 1.05) +
         geom_rug(alpha = .3, position = "jitter", length = unit(0.05, "npc")) +
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = weather), alpha = 0.1, colour = NA)+
         theme_bw() +
        scale_color_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
    scale_fill_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
  # ylim(0, 1)+
  ggtitle("California coast")

psw = ggplot(data = preds_w_sw, aes(x = time_betwe, y = high_sev_prob, color = weather)) +
         geom_line(size = 1.05) +
           geom_rug(alpha = .3, position = "jitter", length = unit(0.05, "npc")) +
         geom_ribbon(aes(ymin = lwr, ymax = upr, fill = weather), alpha = 0.1, colour = NA)+
         theme_bw()+
        scale_color_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
    scale_fill_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
    # ylim(0, 1)+
  ggtitle("Southwest")

pnm = ggplot(data = preds_w_nm, aes(x = time_betwe, y = high_sev_prob, color = weather)) +
         geom_line(size = 1.05) +
           geom_rug(alpha = .3, position = "jitter", length = unit(0.05, "npc")) +
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = weather), alpha = 0.1, colour = NA)+
         theme_bw()+
        scale_color_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
    scale_fill_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
  # ylim(0, 1)+
  ggtitle("Northern mountains")

pwm = ggplot(data = preds_w_wm, aes(x = time_betwe, y = high_sev_prob, color = weather)) +
         geom_line(size = 1.05) +
           geom_rug(alpha = .3, position = "jitter", length = unit(0.05, "npc")) +
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = weather), alpha = 0.1, colour = NA)+
         theme_bw()+
        scale_color_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
    scale_fill_manual(values = c("lightseagreen", "sienna1", "tomato3"))+
  # ylim(0, 1)+
  ggtitle("Western mountains")

grid.arrange(pca, psw, pnm, pwm, nrow= 2, ncol=2)
```

plot interaction between climate/ productivity and time between fires
```{r}
pca = ggplot(data = preds_c_ca, aes(x = time_betwe, y = high_sev_prob, color = climate)) +

    # ylim(0.5, 3)+
            geom_line(size = 1.01) +
             geom_ribbon(aes(ymin = lwr, ymax = upr, fill = climate), alpha = 0.1, colour = NA)+
         theme_bw() +
      scale_color_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
    scale_fill_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
  ggtitle("California coast")

psw = ggplot(data = preds_c_sw, aes(x = time_betwe, y = high_sev_prob, color = climate)) +

    # ylim(0.5, 3)+
         geom_line(size = 1.01) +
             geom_ribbon(aes(ymin = lwr, ymax = upr, fill = climate), alpha = 0.1, colour = NA)+
         theme_bw()+
      scale_color_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
    scale_fill_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
  ggtitle("Southwest")

pnm = ggplot(data = preds_c_nm, aes(x = time_betwe, y = high_sev_prob, color = climate)) +

    # ylim(0.5, 3)+
         geom_line(size = 1.01) +
             geom_ribbon(aes(ymin = lwr, ymax = upr, fill = climate), alpha = 0.1, colour = NA)+
         theme_bw()+
      scale_color_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
    scale_fill_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
  ggtitle("Northern mountains")

pwm = ggplot() +
    # ylim(0.5, 3)+
         geom_line(data = preds_c_wm, aes(x = time_betwe, y = high_sev_prob, color = climate), size = 1.01) +
             geom_ribbon(data = preds_c_wm, aes(x = time_betwe, ymin = lwr, ymax = upr, fill = climate), alpha = 0.1, colour = NA)+
         theme_bw()+
      scale_color_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
    scale_fill_manual(values = c("#247ba0", "sienna1", "#21272A", "#76B041"))+
  ggtitle("Western mountains")


grid.arrange(pca, psw, pnm, pwm, nrow= 2, ncol=2)



```
