---
title: "Reburn analysis"
author: "CT"
date: "2022-11-30"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)



library(randomForest)
library(ranger) #for faster random forest
library(pdp)# for partial dependence plots (PDPs)

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

Load reburn points for analysis
```{r}
data = read_csv(datadir("prelim-outputs/points/points_for_analysis_11-30-22.csv"))

#clean data
data = data %>%
  rename(cbi_reburn = cni85_21) %>%
  select(-c(LC20_Elev_, geom, grid_code)) %>% #remove column with no data
  mutate_all(~replace(., is.na(.), time_betwe)) %>% #set NAs to 0
  mutate(reburn_fire_type = recode(reburn_fire_type, 'Unknown' = 'Wildfire', 'Wildfire Daily Fire Perimeter' = 'Wildfire', 'Wildfire Final Fire Perimeter' = 'Wildfire', 'Wildland Fire Use' = 'Wildfire'),
  initial_fire_type = recode(initial_fire_type, 'Unknown' = 'Wildfire', 'Wildfire Daily Fire Perimeter' = 'Wildfire', 'Wildfire Final Fire Perimeter' = 'Wildfire', 'Wildland Fire Use' = 'Wildfire'),
  ecoregion = as.factor(ecoregion),
  )


```


# Random forests analysis

Partition data to test and train sets
```{r}
datarf = data %>%
    #assign new response var for random forsets - high vs. low/mod reburn severity
    mutate(cbi_reburn_class = cbi_reburn,
      cbi_reburn_class = case_when(cbi_reburn <2.5 ~ 'low-mod',
                           TRUE ~ 'high'),
    cbi_reburn_class = as.factor(cbi_reburn_class)) %>%
  select(-c(cbi_reburn, max_yr_mechanical, max_yr_pfire))
  
set.seed(222)
ind <- sample(2, nrow(data), replace = TRUE, prob = c(0.7, 0.3))
train <- data[ind==1,]
test <- data[ind==2,]


```

RF settings
>important to tune the parameters - #trees and max.depth
>look into importance settings
> RF with raw cbi instead of groups

### Random forest in R
```{r}
rf <- ranger(cbi_reburn_class~., 
             data=train, 
             num.trees = 100,
             max.depth = 8,
             probability=TRUE,
             importance = 'impurity_corrected') #look into these importance options

rf
```

```{r}
#relative importance
sort(importance(rf))

# Prediction wrapper that returns average prediction for each class
pfun <- function(object, newdata) {
  colMeans(predict(object, data = newdata)$predictions)
}

# Partial dependence of probability for each class on petal width
p <- partial(rf, pred.var = "time_betwe", pred.fun = pfun)

ggplot(p, aes(time_betwe, yhat, color = yhat.id)) +
  geom_line() +
  theme(legend.title = element_blank())
```

GAMs

Variable selection for GAMs - 
> based on prediction rather than sig.
> could use RF importance score - how to best generate? (RF: rm pointid, fire years, combine facts, change facts 0s)
> 

Testing model fit:
>tileing > random selection
>change 

```{r}
library(mgcv)

g1 <- bam(cbi_reburn ~ s(AET) + s(time_betwe, by = ecoregion) + s(time_betwe, AET) +
    s(vpd_02_21) + s(th_02_21) + s(fm_02_21) + s(fm100_02_2) + s(vs_02_21) + s(TPI_1024) + s(cbi_initial) + ecoregion, 
    data = train, 
    method = "REML")
g1
summary(g1)
k.check(g1)
gam.check(g1)

g2 <- bam(cbi_reburn ~ s(time_betwe, by = ecoregion),
    data = train, 
    method = "REML")

g2 <- bam(cbi_reburn ~ s(time_betwe, by = ecoregion),
    data = train, 
    method = "REML")

g3 <- bam(cbi_reburn ~ s(time_betwe, AET),
    data = train, 
    method = "REML")


plot(g1)
vis.gam(g1, view = c("time_betwe", "AET"),
    theta = 50, n.grid = 50, lwd = 0.4)
    
vis.gam(g2, view = c("time_betwe", "ecoregion"),
    theta = 50, n.grid = 50, lwd = 0.4)
    
plot(g2)

summary.gam(g1)

newdata = data.frame(cbi_reburn = seq(0.5,3, length.out = 1000))
p1 = predict(g1, newdata = newdata, type = "response" )
```

Other options for variable selection:
> Lasso prior

>group variable list: (explore correlations )
-weather
-topography (TPI)
-history (time since fire, number of burns, intial cbi, management)
-fuels? (ndvi, tree, shrub)
-climate (AET, PRISM)

use GAMs to assess predictive performance with and without predictors




