---
title: "sample_points"
author: "CT"
date: "2022-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

This script generates sample points for the drivers of reburn severity project




## Generate sample points
generate sample points from fire history raster
```{r}
#read in rater with count of number of times each pixel burned
burns_r = raster(datadir("/prelim-outputs/fire_history/count_burns_int.tif"))

#generate 270m raster to snap to when resampling
r270 = aggregate(burns_r, fact=9, fun=max, na.rm = TRUE)

#resample to 270m
burns_r270 = raster::resample(burns_r, r270, method = "ngb", filename = datadir("/prelim-outputs/fire_history/count_burns_resamp270m.tif"))


#convert raster to points for only reburned areas (x > 1 time burned)
reburn_points = rasterToPoints(burns_r270, fun=function(x){x>1}, spatial = TRUE)

#write reburn points to shp
writeOGR(obj=reburn_points, dsn=datadir("/prelim-outputs/points"), layer="reburn_points_270test", driver="ESRI Shapefile")

# plot(reburn_points[1:100000,])
```

#### Buffer reburned area 
so that we don't have any sample points within 300m of edge to limit edge effects and spatial data inconsistencies

Buffer polygons inside 300m
```{r}
#get reburned area - did this in arc for faster processing: 
#> converted count_burns_int.tif to polygons %>% removed areas that only burned once

reburn_poly = st_read(datadir("/prelim-outputs/fire_history/reburn_poly.shp"))

# buffer size = 300m coordinates inward
buffer_size <- (-300)

# compute buffered polygons
perims_buffered <- st_buffer(reburn_poly, buffer_size) 

# remove degenerate polygons
perims_remaining <- perims_buffered[!st_is_empty(perims_buffered),]

#write perims_buffer
st_write(perims_remaining, datadir("prelim-outputs/buffered_perims/buffered_perims.shp"))
# perims_remaining = st_read(datadir("prelim-outputs/buffered_perims/buffered_perims.shp"))
```

Extract points that fall within the buffer to new points dataset
```{r}
reburn_points = st_read(datadir("prelim-outputs/points/reburn_points_270.shp"))

# plot(perims[2,1])
# plot(perims_remaining[2,1])

#keep only points that fall in buffered perims
points_buff = st_intersects(reburn_points, perims_remaining)
points_buff_include = lengths(points_buff) > 0 
points_buff_include = reburn_points[points_buff_include,]
  
  
#write sample points
st_write(points_buff_include, datadir("prelim-outputs/points/sample_points_buff.shp"))

```

# Extract data to points

### Extract raster values to sample points

I did this step with the extract multi values to points tool in Arc. This was failing in R because the files were too big.

Extracted:
> CBI 1985-2021
> reburn cbi ('cni85_21')
> prism climate norms - precip ('PRISM_ppt_30yr_normal_800mM3_annual_bil.bil')
> prism norms - temp ('PRISM_tmean_30yr_normal_800mM3_annual_bil.bil')
> TPI 500 ('TPI_1024.tif')
> Elevation ('LC20_Elev_220_crop_Clip')
> heat load
> burn count ('count_burns.tif')
> time between burns (time_between_burns_all.tif)
> year of second most recent fire (secondrecent_fire_all.tif)
> year of most recent fire (mostrecent_fire_all.tif)
>6 weather variables: fm100, fm1000, erc, th, vpd, ts

```{r}
#get cbi data
# rastlist <- list.files(path = datadir("/raw-data/cbi-GEE"), pattern='.tif$', all.files=TRUE, full.names=TRUE)
# cbi_stack <- stack(rastlist, quick=TRUE) #stack annual cbi
# 
# #extract values to points
# points_cbi = extract(cbi_stack, points_buff_include) #failing because files are too big - going to extract raster data in ArcGIS instead.
```


## Combine point data extracted from GEE with other point data
```{r GEE ndvi and veg data}
points = st_read(datadir("prelim-outputs/points/sample_points_buff.shp"))

## read in GEE ndvi data
ndvi1 = read_csv(datadir("prelim-outputs/GEE_points_extract/ndvi/ndvi_extract1.csv"))
ndvi2 = read_csv(datadir("prelim-outputs/GEE_points_extract/ndvi/ndvi_extract2.csv"))
ndvi3 = read_csv(datadir("prelim-outputs/GEE_points_extract/ndvi/ndvi_extract3.csv"))
ndvi4 = read_csv(datadir("prelim-outputs/GEE_points_extract/ndvi/ndvi_extract4.csv"))
ndvi5 = read_csv(datadir("prelim-outputs/GEE_points_extract/ndvi/ndvi_extract5.csv"))

#combine data subsets
ndvi = bind_rows(ndvi1, ndvi2, ndvi3, ndvi4, ndvi5)

## read in GEE RAP data (these are in multiple parts to allow for GEE extract within memory limits)
shrub1 = read_csv(datadir("prelim-outputs/GEE_points_extract/rap-veg/ShrubExtractPoints.csv"))
shrub2 = read_csv(datadir("prelim-outputs/GEE_points_extract/rap-veg/ShrubExtractPoints2.csv"))
shrub3 = read_csv(datadir("prelim-outputs/GEE_points_extract/rap-veg/ShrubExtractPoints3.csv"))

tree1 = read_csv(datadir("prelim-outputs/GEE_points_extract/rap-veg/TreeExtractPoints.csv"))
tree2 = read_csv(datadir("prelim-outputs/GEE_points_extract/rap-veg/TreeExtractPoints2.csv"))
tree3 = read_csv(datadir("prelim-outputs/GEE_points_extract/rap-veg/TreeExtractPoints3.csv"))

#combine data subsets
shrub = bind_rows(shrub1, shrub2, shrub3)
tree = bind_rows(tree1, tree2, tree3)

points_reburnyr = points %>% select(pointid, mostrecent)

#keep veg data for year prior to reburn
shrubl = shrub %>% 
  pivot_longer(!pointid, names_to = "shrub_yr", values_to = "shrub_rebrn") %>%
  merge(points_reburnyr, ., all = TRUE) %>% 
  mutate(shrub_yr = substr(shrub_yr, 1, 4),
         subset_yr = mostrecent - 1) %>% 
  subset(shrub_yr == subset_yr) %>%
  select(pointid, shrub_rebrn)

treel = tree %>% 
  pivot_longer(!pointid, names_to = "tree_yr", values_to = "tree_reburn") %>%
  merge(points_reburnyr, ., all = TRUE) %>% 
  mutate(tree_yr = substr(tree_yr, 1, 4),
         subset_yr = mostrecent - 1) %>% 
  subset(tree_yr == subset_yr) %>%
  select(pointid, tree_reburn)

#select only NDVI from just before reburn 

#rename columns to align with correct years
colnames(ndvi) <- c(2001, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2002, 2021, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, "pointid")

ndvil = ndvi %>% pivot_longer(!pointid, names_to = "ndvi_yr", values_to = "ndvi") %>%
  full_join(points_reburnyr, .) %>%
  mutate(subset_yr = mostrecent - 1) %>%
  subset(ndvi_yr == subset_yr) %>%
  select(pointid, ndvi)



#merge GEE data to points
points_gee = full_join(points, data.frame(ndvil))
points_gee = full_join(points_gee, data.frame(shrubl))
points_gee = full_join(points_gee, data.frame(treel))
```


#### Create new variable for intial burn CBI
```{r initial burn cbi}
cbi = points_gee %>% 
  data.frame() %>%
  select(pointid, secondrece, cbi_2021:cbi_1995) %>%
  pivot_longer(!c(pointid, secondrece), names_to = "cbi_yr", values_to = "cbi_initial") %>%
  mutate(cbi_yr = as.numeric(substr(cbi_yr, 5, 8))) %>%
  subset(cbi_yr == secondrece) %>%
  select(pointid, cbi_initial)

points_gee = full_join(points_gee, data.frame(cbi))

#remove uneccesary cols
points_cleaned = points_gee %>%
  select(-c(cbi_2021:cbi_1985))
```

### Add in ecoregion data
```{r}
eco = st_read(datadir("raw-data/ecoregion/ecoregions_grouped.shp"))
# plot(eco)

points_cleaned = st_intersection(points_cleaned, eco)

```


#### Extract cleaned FACTS management data to points & keep only relevant management
```{r extract FACTS to points}
#read in facts
facts <- st_read(datadir("raw-data/facts/compiled/facts_compiled_for_points_extract_west.gpkg"))

facts = facts %>% st_transform(st_crs(points_cleaned)) %>%
  mutate(activiy_yr = paste(ACTIVITY, year, sep = "_")) %>%
  select(activity_yr, geom)
# plot(facts[1:5000,1])

#select relevant cols for facts intersection 
points_for_facts_int = points_cleaned %>% select(pointid, mostrecent, secondrece)  

# Extract FACTS history at each plot 
facts_int = st_intersection(points_for_facts_int, facts)
# st_write(facts_int, datadir("prelim-outputs/points/points_buff_facts_int.gpkg"))
# facts_int = st_read(datadir("prelim-outputs/points/points_buff_facts_int.gpkg"))

#Keep only management that occurs between burn years
facts_keep = facts_int %>%
  mutate(facts_yr = as.numeric(str_sub(activity_yr,-4,-1)),
         facts_activ = as.factor(str_sub(activity_yr, 1, -6))) %>%
  filter(facts_yr > secondrece & facts_yr < mostrecent)

#group so that each point has only one row for management 
facts_group = facts_keep %>% data.frame() %>%  
  select(-activity_yr) %>%
  group_by(pointid, facts_activ) %>%
  summarise(max_yr = max(facts_yr),
            count_trt = n()) %>%
  left_join(., data.frame(points_for_facts_int)) %>% #add fire year cols
  pivot_wider(-geometry, names_from = facts_activ, values_from = c(max_yr,count_trt)) %>%
  #add columns for years between treatment and reburn
  mutate(mechanical_yrs2reburn = mostrecent - max_yr_mechanical, 
         pfire_yrs2reburn = mostrecent - max_yr_pfire,
         count_trt = sum(count_trt_mechanical, count_trt_pfire, na.rm = TRUE))

#check for duplicates
# summary(duplicated(facts_group$pointid))

#join management info to points_cleaned
points_cleaned_facts = left_join(points_cleaned, facts_group)

# write.csv(points_cleaned_facts, datadir("prelim-outputs/points/points_cleaned_facts.csv"))
```


### Extract Fire perimeter data - MTBS & 2021 perimeter information

Determine if fires were wildfires or perscribed 
```{r}
#perimeter data
perims = st_read(datadir("raw-data/fire-perims/mtbs_perimeter_data/mtbs_west_fs_nps_sub_updated21.shp"))

points_perims_int = st_intersection(points, perims)

# st_write(points_perims_int, datadir("prelim-outputs/points/points_perims_int.shp"))
# points_perims_int = st_read(datadir("prelim-outputs/points/points_perims_int.shp"))

#get reburn fire id
reburn_fire_id = points_perims_int %>% data.frame() %>%
    full_join(points_for_facts_int, .) %>%
    filter(Fire_Year == mostrecent,
           Fire_Year > 2001,) %>% #filtering out reburns that have no weather data and will be excluded from analysis
  select(pointid, Fire_ID) %>% data.frame()

#get fire ID for second most recent burn
initial_fire_id = points_perims_int %>% data.frame() %>%
    full_join(points_for_facts_int, .) %>%
    filter(Fire_Year == secondrece) %>% 
  select(pointid, Fire_ID, secondrece) %>% data.frame()

### join information about fire id to point

#read in mtbs and wfigs fire_id infmation and select relevant columns
mtbs_ids = read.csv(datadir("raw-data/fire_id_tables/mtbs_id_table.csv")) %>%
  mutate(Fire_ID = Event_ID, 
         Fire_Type = as.factor(Incid_Type)) %>%
  select(Fire_ID, Fire_Type) 

wfigs_ids = read.csv(datadir("raw-data/fire_id_tables/wfigs21_id_table.csv")) %>% 
    mutate(Fire_ID = irwin_UniqueFireIdentifier, 
         Fire_Type = as.factor(poly_FeatureCategory)) %>%
  select(Fire_ID, Fire_Type) 

#combine into one fire id table
fire_id_table = rbind(mtbs_ids, wfigs_ids)

#join to fire peimeter points interesct dfs above
reburn_fire_id = left_join(reburn_fire_id, fire_id_table) %>%
  mutate(reburn_fire_type = Fire_Type) %>%
  select(pointid, reburn_fire_type)

initial_fire_id = left_join(initial_fire_id, fire_id_table) %>%
    mutate(initial_fire_type = Fire_Type) %>%
  select(pointid, initial_fire_type)

fire_id = left_join(reburn_fire_id, initial_fire_id)

# check if there are multiple fires in the same year for the same point id
# fire_id[duplicated(fire_id$pointid)==TRUE,] #22 duplicate ids for initial fires before 2000 and all wildfire, so not a problem

#join to points data
points_cleaned_facts_perims = left_join(points_cleaned_facts, fire_id)


#write intersected sample points
st_write(points_cleaned_facts_perims, datadir("prelim-outputs/points/points_cleaned_facts_perims.gpkg"))
write.csv(points_cleaned_facts_perims, datadir("prelim-outputs/points/points_cleaned_facts_perims.csv"))
```

# Select points for analysis based on following criteria:

- Not on FS land (focus analysis on forested FS land, and for management)
- Less than 25% tree cover at time of reburn (focus analysis on forested areas)
- second fire occurred before 2002 (limited by MODIS for day of fire weather)
- burned at <0.5 CBI in initial or reburn fires (remove unburned areas to limit residual error where fire did not make it to a point within a fire perimeter)


```{r remove points based on criteria}
#load ca fs land spatial data
fs = read_sf(datadir("/raw-data/ownership/S_USA.BasicOwnership.gdb"), "BasicOwnership") %>% subset(OWNERCLASSIFICATION == "USDA FOREST SERVICE")

fs = st_transform(fs, st_crs(points_cleaned_facts_perims))  #set crs equal to points

#keep only points that fall within FS land
points_fs = st_intersects(points_cleaned_facts_perims, fs)
points_fs_include = lengths(points_fs) > 0 
points_fs_include = points_cleaned_facts_perims[points_fs_include,]

#remove points that... 
points_crit = points_fs_include %>% 
  filter(tree_reburn >= 25 & #have greater than 25% tree cover at time of reburn
           mostrecent > 2001 & #second fire occured after 2001 (for MODIS weather data)
           cbi_initial > 0.5 & #burned at >0.5 CBI in initial or reburn fires (remove unburned areas to limit residual error
           cni85_21 > 0.5) %>%
  filter(!is.na(ecoregion)) #remove points that fall outside of our 4 ecoregions

#write data for analysis
st_write(points_crit, datadir("prelim-outputs/points/points_for_analysis_11-30-22.gpkg"))  
write.csv(points_crit, datadir("prelim-outputs/points/points_for_analysis_11-30-22.csv"))  
```

