---
title: "Reburn analysis"
author: "CT"
date: "2022-11-30"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


This script analyzes data for reburn analyses.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(randomForest)
library(ranger) #a faster version of the RF algorithm
# library(pdp)# for partial dependence plots (PDPs)

library(tidyverse)
library(here)
library(sp)
library(sf)
library(terra)
library(raster)

library(mgcv)
library(tidymv)

library(gridExtra)
library(BiodiversityR)


data_dir = readLines(here("data_dir_D.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

Load reburn points with low/moderate initial CBI for analysis

```{r}
data_thin = read_csv(datadir("final-points/thinned_data_final.csv"))

```

## Explore data


summary stats
```{r}

count_unique_reburns = data_thin %>%
  group_by(initial_fireID, reburn_fireID, ecoregion) %>%
  summarise(n = n()) %>%
  group_by(ecoregion) %>%
  summarise(n = n())
sum(count_unique_reburns$n)

#count sample points
data_thin %>%
  group_by(ecoregion) %>%
  summarise(n = n()) 
nrow(data_thin)


#summary stats
sum_stat_tbl = data_thin %>% gather(var, value, c(time_betwe, cbi_initial, cbi_reburn)) %>%
  group_by(var, ecoregion) %>%
  summarise(mean = round(mean(value),2),
            SD = round(sd(value),2),
            median = round(quantile(value, probs = 0.5),2),
            lower.x = round(quantile(value, probs = 0.25),2),
            upper.x = round(quantile(value, probs = 0.75),2))

summary(as.factor(data_thin$time_betwe))
```

### Explore correlations between variables
```{r}
library(purrr)

#keep only numeric variables
data_num = data_thin %>%
  keep(is.numeric)

# library(PerformanceAnalytics)
# cor(data_num, method = "pearson", use = "complete.obs")
# chart.Correlation(cor(data_num))

weather = data_thin %>% data.frame() %>%
  dplyr::select(erc_perc, vs_perc, vpd_perc, erc_02_21, vs_02_21, vpd_02_21, th_02_21, fm1000_02_, fm100_02_2, cbi_reburn, lat, long)

clim_topo = data_thin %>% data.frame() %>%
  dplyr::select(heat_load, PRISM_tmea, PRISM_ppt_, PRISM_vpdm, AET, ndvi, shrub_rebrn, tree_reburn, afg_rebrn, pfg_rebrn, TPI_1024, cbi_reburn, time_betwe, lat, long)

firehist = data_thin %>% data.frame() %>%
  dplyr::select(time_betwe, count_burn, cbi_reburn, cbi_initial, lat, long)

corplot = data_thin %>% data.frame() %>%
  dplyr::select(erc_perc, vs_perc, vpd_perc, heat_load, PRISM_tmea, PRISM_ppt_, PRISM_vpdm, AET, ndvi, shrub_rebrn, tree_reburn, afg_rebrn, pfg_rebrn, TPI_1024, cbi_reburn, time_betwe, count_burn, cbi_initial, cbi_reburn)
```

Plot correlations
```{r}
library(GGally)
library(corrplot)

corrplot(cor(weather),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

corrplot(cor(clim_topo),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

corrplot(cor(firehist),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

#save corrplot
png(height=800, width=800, file=datadir("figures/corrplot.png"))

corrplot(cor(corplot),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)
dev.off()

#strongest correlations between weather variables, fuels variables, and lat/long
#> selected precip and temp over AET and PRISM_vpd, and erc over fm100/1000
```



### Fit separate models for each ecoregion

```{r}
#scale numeric predictors for model fit - don't scale time between t or cbi or erc percentiles to maintain interpretation
preds_2scale = c("heat_load", "PRISM_tmea", "PRISM_ppt_", "TPI_1024", "fm100_02_2", "th_02_21", "vpd_02_21", "erc_02_21", "PRISM_vpdm", "AET", "fm1000_02_", "ndvi", "shrub_rebrn", "tree_reburn", "afg_rebrn", "pfg_rebrn", "cbi_initial", "time_betwe", "erc_perc", "vpd_perc", "vs_perc")

data_thin = data.frame(data_thin)

scale_df = function(df){
  data_scale = scale(df[(names(df) %in% preds_2scale)], scale = TRUE, center = TRUE)
  data_scale_all = cbind(data_scale, df[!(names(df) %in% preds_2scale)])
  return(data_scale_all)
}

data_thin_scale = data_thin %>% scale_df()

#create new dfs for each ecoregion and scale each ecoregion separately 
data_lowmod_ca = data_thin %>% filter(ecoregion == "California Coast") %>% scale_df()
data_lowmod_nm = data_thin %>% filter(ecoregion == "Northern Mountains") %>% scale_df()
data_lowmod_sw = data_thin %>% filter(ecoregion == "Southwest") %>% scale_df()
data_lowmod_wm = data_thin %>% filter(ecoregion == "Western Mountains") %>% scale_df()

unscale_df = data_thin %>%
  group_by(ecoregion) %>%
  summarise(timebetwe_sd = sd(time_betwe),
            timebetwe_mean = mean(time_betwe),
            cbi_initial_sd = sd(cbi_initial),
            cbi_initial_mean = mean(cbi_initial),
            erc_perc_sd = sd(erc_perc),
            erc_perc_mean = mean(erc_perc),
            vpd_perc_sd = sd(vpd_perc),
            vpd_perc_mean = mean(vpd_perc),
            vs_perc_sd = sd(vs_perc),
            vs_perc_mean = mean(vs_perc))
```



Test for spatial autocorrelation
```{r}
library(ape)
inv_dist = with(data_lowmod_ca, 1/dist(cbind(lat, long), diag = TRUE, upper = TRUE))
inv_dist = as.matrix(inv_dist)
ape::Moran.I(data_lowmod_ca$cbi_reburn, weight = inv_dist, scaled = TRUE)

#observed = 0.15  - CA; 0.11 - NM; 0.17 for SW, 0.07 for wm
# While statistically significant, there actually isn’t too much going on, though it may be enough to warrant dealing with in some fashion.

```


## Model fit with mgcv 
>https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/gam.selection.html 
>https://stats.stackexchange.com/questions/519433/gam-and-multiple-continuous-continuous-interactions-tensor-smooths #helpful for interpretting and fitting interactions (ti vs te vs s)

>all about gams: https://m-clark.github.io/generalized-additive-models/application.html#gam-1


```{r}

fit_ca = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe,PRISM_ppt_, k = 3, bs = "cs") +
                  ti(time_betwe,PRISM_tmea, k = 3, bs = "cs") +
                  ti(time_betwe,vpd_perc, k = 3, bs = "cs") +
                  ti(time_betwe,vs_perc, k = 3, bs = "cs") +
                  ti(time_betwe,erc_perc, k = 3, bs = "cs") +
                  ti(time_betwe,cbi_initial, k = 3, bs = "cs"),
             data = data_lowmod_ca,
             select = T,
             gamma = 1.4,
             method = "REML") 

fit_nm = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe,PRISM_ppt_, k = 3, bs= "cs") +
                  ti(time_betwe,PRISM_tmea, k = 3, bs = "cs") +
                  ti(time_betwe,vpd_perc, k = 3, bs = "cs") +
                  ti(time_betwe,vs_perc, k = 3, bs = "cs") +
                  ti(time_betwe,erc_perc, k = 3, bs = "cs") +
                  ti(time_betwe,cbi_initial, k = 3, bs = "cs"),
             data = data_lowmod_nm,
             select = T,
             gamma = 1.4,
             method = "REML") 

fit_sw = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe,PRISM_ppt_, k = 3, bs = "cs") +
                  ti(time_betwe,PRISM_tmea, k = 3, bs = "cs") +
                  ti(time_betwe,vpd_perc, k = 3, bs = "cs") +
                  ti(time_betwe,vs_perc, k = 3, bs = "cs") +
                  ti(time_betwe,erc_perc, k = 3, bs = "cs") +
                  ti(time_betwe,cbi_initial, k = 3, bs = "cs"),
             data = data_lowmod_sw,
             select = T,
             gamma = 1.4,
             method = "REML") 

fit_wm = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe,PRISM_ppt_, k = 3, bs = "cs") +
                  ti(time_betwe,PRISM_tmea, k = 3, bs = "cs") +
                  ti(time_betwe,vpd_perc, k = 3, bs = "cs") +
                  ti(time_betwe,vs_perc, k = 3, bs = "cs") +
                  ti(time_betwe,erc_perc, k = 3, bs = "cs") +
                  ti(time_betwe,cbi_initial, k = 3, bs = "cs"),
             data = data_lowmod_wm,
             select = T,
             gamma = 1.4,
             method = "REML") 
```


Check residuals
```{r}
library(mgcViz)
#k = 5, bs = "cs"
gam.check(fit_ca) 
gam.check(fit_nm) 
gam.check(fit_sw) 
gam.check(fit_wm)
#- still some significance in the residuals tests, but edf is below 6 so there are likely other issues with the models? thinning helped this a bit

check(getViz(fit_ca))
check(getViz(fit_nm)) 
check(getViz(fit_sw))
check(getViz(fit_wm))

#including lat long did not improve the residuals and it is highly correlated with other predictors, so leaving out


#check for concurvity - non-linearly related variables; Ignore interactions? 
concurv_ca = concurvity(fit_ca, full = FALSE)  
concurv_ca = concurv_ca$observed

resid_ca = residuals(fit_ca) 
resid_nm = residuals(fit_nm)
resid_sw = residuals(fit_sw)
resid_wm = residuals(fit_wm)

#add residuals to dfs
data_lowmod_ca$resids = resid_ca
data_lowmod_nm$resids = resid_nm
data_lowmod_sw$resids = resid_sw
data_lowmod_wm$resids = resid_wm

#plot residuals against covariates
plot_resids_cov = function(data){
  resid_plot = data %>%
  dplyr::select(PRISM_ppt_, TPI_1024,
                PRISM_tmea, cbi_initial, time_betwe,
                erc_perc, vs_perc, vpd_perc, heat_load, resids) %>%
  pivot_longer(-resids, names_to = "key", values_to = "value") %>% #convert to long format for fast plotting
  ggplot(aes(x = scale(value), y = resids)) +
  facet_wrap(~ key, scales = "free") +
  geom_smooth(formula = y ~ s(x, k = 6)) +
  theme_bw(13)
  return(resid_plot)
}

plot_resids_cov(data_lowmod_ca)
plot_resids_cov(data_lowmod_nm)
plot_resids_cov(data_lowmod_sw)
plot_resids_cov(data_lowmod_wm)
#covaraite/residual relationships look good! 
```

Test residual autocorrelation
```{r}
#investigate spatial autocorrelation
library(gstat) 

#project data
project_func = function(df){
  xy <- df[,c("long","lat")]
  spdf <- SpatialPointsDataFrame(coords = xy, data = df,
                             proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84
                                               +towgs84=0,0,0"))
  return(spdf)
}

ca_proj = project_func(data_lowmod_ca)
sw_proj = project_func(data_lowmod_sw)
nm_proj = project_func(data_lowmod_nm)
wm_proj = project_func(data_lowmod_wm)

vca = plot(variogram(resid_ca~1, data=ca_proj, cutoff = 5), main = "cali coast")
vnm = plot(variogram(resid_nm~1, data=nm_proj, cutoff = 5), main = "northern mnts")
vsw = plot(variogram(resid_sw~1, data=sw_proj, cutoff = 5), main = "southwest")
vwm = plot(variogram(resid_wm~1, data=wm_proj, cutoff = 5), main = "Western mnts")

grid.arrange(vca, vsw, vnm, vwm, nrow= 2, ncol=2)

#check moran's I
pixel.dists <- as.matrix(dist(cbind(ca_proj$long, ca_proj$lat)))
pixel.dists.inv <- 1/pixel.dists
diag(pixel.dists.inv) <- 0

pixel.dists <- as.matrix(dist(cbind(wm_proj$long, ca_proj$lat)))
pixel.dists.inv <- 1/pixel.dists
diag(pixel.dists.inv) <- 0


print(Moran.I(resid_wm,pixel.dists.inv))
```

Summarise models
```{r}
library(texreg) #for creating easy tables from gam output
extract(fit_ca)
extract(fit_wm)
extract(fit_sw)
extract(fit_nm)

summary(fit_ca)

texreg(list(fit_ca,fit_nm,fit_sw, fit_wm),
       caption="GAM output by ecoregion",
       dcolumn=FALSE,
       model.names=c("California Coast","Northern Mountains","Southwest", "Western Mountains"))

```

Save models for plotting
```{r}
saveRDS(fit_ca, datadir("final-models/fit_ca.rds"))
saveRDS(fit_sw, datadir("final-models/fit_sw.rds"))
saveRDS(fit_nm, datadir("final-models/fit_nm.rds"))
saveRDS(fit_wm, datadir("final-models/fit_wm.rds"))

```

