---
title: "extract-reburn-area"
author: "CT"
date: '2022-07-20'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(lubridate)
library(tidyverse)
library(here)
library(rgdal)
library(rgeos)
library(geos)
library(sp)
library(sf)
library(readxl)
library(terra)
library(raster) # for the rasterize function in this package
library(tictoc) # for simple tracking of computation time

data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```



This script extracts reburned area from fire perimeters for analysis

```{r load data}
#load mtbs data
mtbs_perims = vect(datadir("/raw-data/fire-perims/mtbs_perimeter_data/mtbs_ca_zones/mtbs_west.shp")) 

```



## Mapping reburned areas

Map areas that have burned at least 2x

> Q I think that is what this code is doing. I was having a lot of issues with providing only one polygon to st_intersection. See next chunk.

```{r}
#convert spatvector to sf 
mtbs_sf = mtbs_perims %>% st_as_sf()


mtbs_int = mtbs_sf[2] %>% st_buffer(., dist = 0) %>% st_intersection(mtbs_sf[2], mtbs_sf[2])
reburn2 = mtbs_int[which(mtbs_int$Fire_Year != mtbs_int$Fire_Year.1),]

reburn_polys = st_collection_extract(reburn2, "POLYGON")
# st_write(reburn_polys, datadir("/prelim-outputs/reburns/reburn85_21.shp")) 

# plot(mtbs_sf["Fire_Year"])
# plot(reburn["Fire_Year.1"])

# mtbs_intersect = st_intersects(mtbs_sf)
```

union polygons within reburn for extracting information from predictor varibales
```{r}
#combine polygons within reburn
reburn_union = reburn_polys %>% group_by(Fire_Year.1) %>%
    summarise(geometry = sf::st_union(geometry)) %>%
    ungroup()

plot(reburn_union)
```

--------------------------------------------------------------------------------------------



intersection with one vector to get n.overlaps (below) is not working :( 

Works with smaller set of polygons (~100) - but not even with 1/36t the size of CA <200 polygons
CBR: result (after common-bits addition) is INVALID: Self-intersection at or near point 475670.37177563249 4608529.2648357973 (475670.37177563249134 4608529.2648357972503)
CPL_nary_intersection(x) : GEOS exception
```{r}

# 
# mtbs_int_all = mtbs_sf %>% st_buffer(dist = 0) %>% st_intersection()
# 
# mtbs_reburn = st_sf(mtbs_int)
# mtbs_int[which(mtbs_int$n.overlaps > 1),]
# 
# plot(mtbs_reburn)
# 
# crs(mtbs_sf)
# 
# mtbs_int = mtbs_sf_test[1] %>% st_buffer(dist = 0) %>% st_intersection() #works with small dataset but not with even 1/36th of california fire perims #Error in CPL_nary_intersection(x) : GEOS exception


# plot(mtbs_int["n.overlaps"]) 
# 
#   
# plot(mtbs_int[2])  
#   
#  %>% st_intersection(x = mtbs_sf) #test intersection
# 
# test = z1 %>% st_sf() %>% st_buffer(0) %>% st_intersection()
# 
# reburn = zone_list_sf[[1]][1:2] %>%
#   st_intersection(st_buffer(., 0)) %>% 


#get areas with two different Fire_IDs (I think these are reburned areas?)
# reburn$reburned = reburn$Fire_ID == reburn$Fire_ID.1

```

intersect with terra and SpatVectors - I don't think this is doing what we want..

```{r}
# #intersect fire perims
# mtbs_intersect = intersect(mtbs_perims, mtbs_perims)
# #select relevant columns
# mtbs_intersect = mtbs_intersect[,c(1:2,14:15)]
# #rename columns
# names(mtbs_intersect) = c("Fire_ID", "Fire_Year", "Fire_ID2", "Fire_Year2")
# #extract only areas that reburned
# reburn = mtbs_intersect[which(mtbs_intersect$Fire_ID != mtbs_intersect$Fire_Year2),]

# plot(reburn[1:500])

```


-------------------------------------------------
Raster approach

Rasterize mtbs for easy extraction
```{r}
# Rasterize mtbs for easy extraction.  - need to use project coordinate system - like Albers equal area conical 
# Equal Area Conus Albers. - epsg:5070


r = rast(zone_list[[1]], resolution = 30, crs = "epsg:26910")

```

Extract data on the most recent and second most recent fires
```{r extract most recent fires}

tic()
mostrecent_fire = raster::rasterize(zone_list[[1]] %>% as("Spatial"), r %>% raster(), field="Fire_Year", fun=max)
toc()

#function to extract second most recent fire 
second_recent_func = function(x, na.rm) {
  x = x[!is.na(x)]
  if(length(x) == 0) { return(NA) }
  x_nomax = x[x != max(x)]
  if (length(x_nomax) == 0) { return(NA) }
  second_recent =  max(x_nomax)
  return(second_recent)
}

tic()
secondrecent_fire = raster::rasterize(zone_list[[1]] %>% as("Spatial"), r %>% raster, field="Fire_Year", fun=second_recent_func)
toc()
#~2.3 mins at 30m res

```

---------------------------------------------------------------------------------

------------------------------------------------------------------------------------
# point method 

Create grid of points for extracting data to 
*this is really slow at 30m resolution!!
```{r point grid, eval = FALSE}
## create a grid of points
tic()
grdpts = sf::st_make_grid(mtbs_ca1, cellsize = 30, what = "centers")
toc()

## convert it to an `sf` object, as opposed to an `sfc`
grdpts = sf::st_sf(grdpts)
grdpts = vect(grdpts)

# plot(grdpts)
# plot(mtbs_ca1, add = TRUE)

#add identifier
grdpts$ID = 1:length(grdpts)
```

extract mtbs data to points
```{r}
mtbs_pnts = intersect(mtbs_ca1, grdpts)

```