---
title: "Reburn analysis"
author: "CT"
date: "2022-11-30"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


This script analyzes data for reburn analyses.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(randomForest)
library(ranger) #a faster version of the RF algorithm
# library(pdp)# for partial dependence plots (PDPs)

library(tidyverse)
library(here)
library(sp)
library(sf)
library(terra)
library(raster)

library(mgcv)
library(tidymv)

library(gridExtra)
library(BiodiversityR)
library(patchwork)


data_dir = readLines(here("data_dir_D.txt"), n=1)
source(here("scripts/convenience_functions.R"))

```

Load reburn poitns with low/moderate initial CBI for analysis

```{r}
data_thin = read_csv(datadir("prelim-outputs/points/thinned_data_3-22-23.csv"))


ggplot(data = data_thin, aes(x = ecoregion, y = AET)) +
  geom_boxplot() + theme_bw() + ylab("Actual evapotranspiration (AET)") + xlab("")

ggplot(data = data_thin, aes(x = ecoregion, y = PRISM_ppt_)) +
  geom_boxplot()

ggplot(data = data_thin, aes(x = ecoregion, y = AET)) +
  geom_boxplot()
```

## Explore data

summary stats
```{r}

count_unique_reburns = data_thin %>%
  group_by(initial_fireID, reburn_fireID, ecoregion) %>%
  summarise(n = n()) %>%
  group_by(ecoregion) %>%
  summarise(n = n())
sum(count_unique_reburns$n)

#count sample points
data_thin %>%
  group_by(ecoregion) %>%
  summarise(n = n()) 
nrow(data_thin)


#summary stats
sum_stat_tbl = data_thin %>% gather(var, value, c(time_betwe, cbi_initial, cbi_reburn)) %>%
  group_by(var, ecoregion) %>%
  summarise(mean = round(mean(value),2),
            SD = round(sd(value),2),
            median = round(quantile(value, probs = 0.5),2),
            lower.x = round(quantile(value, probs = 0.25),2),
            upper.x = round(quantile(value, probs = 0.75),2))

summary(as.factor(data_thin$time_betwe))

#summarise high severity
high_sev = data_thin %>% filter(cbi_reburn >2.5) %>% group_by(ecoregion) %>% summarise(highsev_count = n())

ecoregion_sum = data_thin %>% group_by(ecoregion) %>%
  summarise(n = n())

high_sev$highsev_perc = high_sev$highsev_count/ecoregion_sum$n *100

#percent of prescribed fire in each ecoregion
data_thin %>%
    group_by(ecoregion, reburn_fire_type) %>%
    summarise(n = n())

data_thin %>% filter(reburn_fire_type == "Prescribed Fire") %>% nrow() / nrow(data_thin) *100
```

```{r}
#ecoregion pallets 
pal1 = c("#264653","#2A9D8F","#dbd053","#6a0136")
pal2 = c("#264653","#2A9D8F","#883c0c", "#dbd053")


library(scales)
#climate space plot
clim_plot = slice(data_thin, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = ecoregion, x = PRISM_tmea, y = PRISM_ppt_)) +
  geom_point(alpha = .3, size = 1.8)+
  stat_ellipse(size = 1, type = "t")+ # 95% confidence level for a multivariate t-distribution
  theme_bw(13)+
  ylab("Precipitation (mm)") +
  xlab(expression("Temperature " ( degree*C)))+
  scale_color_manual(values = pal2) + 
  theme(legend.position="bottom", legend.title=element_blank())+
  guides(colour = guide_legend(nrow = 2))

#weather space plot
weather_plot = slice(data_thin, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = ecoregion, x = erc_perc, y = vpd_perc)) +
  geom_point(alpha = .3, size = 1.8)+
  # stat_ellipse(size = 1, type = "t")+ # 95% confidence level for a multivariate t-distribution
  theme_bw(13)+
  ylab("ERC percentile") +
  xlab("VPD percentile")+
  scale_color_manual(values = pal2) + 
  theme(legend.position = "none")

weather_plot  + clim_plot + plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect") & theme(legend.position = 'bottom') & theme(legend.title=element_blank())
# ggsave('scatter_plots.png', height = 4.5, width = 7)
```

Histograms
```{r}
## Histograms 
library(ggridges)
#time between
time_hist = ggplot(data = data_thin, aes(fill = ecoregion, x = time_betwe*10)) +
  geom_histogram(bins = 20)+
  theme_bw(13)+
  xlab("time between initial fire and reburn") +
  scale_fill_manual(values = pal2)+
  facet_wrap(~ ecoregion)+
      theme(legend.position = "none")


#reburn cbi
reburn_hist = ggplot(data = data_thin, aes(fill = ecoregion, x = cbi_reburn)) +
  geom_histogram()+
  theme_bw(13)+
  xlab("Reburn CBI") +
    ylab("")+
  scale_fill_manual(values = pal2)+
  scale_x_continuous(breaks = c(0.51, 1, 1.5, 2, 2.5))+
  facet_wrap(~ecoregion)+
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())


#initial cbi
initial_hist = ggplot(data = data_thin, aes(fill = ecoregion, x = cbi_initial)) +
  geom_histogram()+
  theme_bw(13)+
  xlab("Initial CBI") +
      ylab("")+
  scale_fill_manual(values = pal2)+
  scale_x_continuous(breaks = c(0.51, 1, 1.5, 2, 2.5))+
  facet_wrap(~ecoregion)+
        theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")


time_hist + initial_hist + reburn_hist

#raincloud plots
library(ggforce)
library(ggdist)
#time between but only one point per unique fire combo instead of by sample point
time_plot = data_thin %>%
    group_by(initial_fireID, reburn_fireID, ecoregion) %>%
    summarise(time_betwe = mean(time_betwe)) %>%
    ggplot(aes(x = time_betwe*10, y = ecoregion, fill = ecoregion, color = ecoregion)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA) + 
  geom_point(size = 1, alpha = .05,
    position = position_jitter(seed = 1, width = .2, height = .2))  +
    geom_boxplot(width = .15, outlier.shape = NA, fill = NA, color = "black", size = 0.5) +
  xlab("Years between initial fire and reburn") +
  ylab("")+
  scale_fill_manual(values = pal2)+
  scale_color_manual(values = pal2)+
      theme_bw(13)+
      theme(legend.position = "none")


```

Variable raincloud plots
```{r}
##Raincloud plots
rain_plot_func = function(var){
  p = ggplot(data_thin, aes(x = var, y = ecoregion, 
                                   fill = ecoregion, color =
                          ecoregion)) + 
    ggdist::stat_halfeye(
      adjust = .5, width = .6, .width = 0, justification = -.2,
      point_colour = NA) + 
    geom_point(size = 1, alpha = .05,
               position = position_jitter(seed = 1, height = .15))+
    geom_boxplot(width = .15, outlier.shape = NA, 
                 fill = NA, color = "black", size = 0.5) +
    ylab("")+
    scale_fill_manual(values = pal2)+
    scale_color_manual(values = pal2)+
    theme_bw(13)+
    ylab("")
  
  return(p)
}

#initial cbi
(icbi_plot = rain_plot_func(data_thin$cbi_initial) +
        xlab("Initial fire severity (CBI)") +
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none") +
        scale_x_continuous(breaks = c(0.51, 1, 1.5, 2, 2.5)))

#reburn cbi
(rcbi_plot = rain_plot_func(data_thin$cbi_reburn) +
        xlab("Reburn fire severity (CBI)") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")+
  scale_x_continuous(breaks = c(0.51, 1, 1.5, 2, 2.5,3)))

#reburn cbi
(tbf_plot = rain_plot_func(data_thin$time_betwe*10) +
        xlab("Time between fires (years)") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))

#wind speed
(vs_plot = rain_plot_func(data_thin$vs_perc) +
        xlab("Wind speed percentile") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))

(erc_plot = rain_plot_func(data_thin$erc_perc) +
        xlab("ERC percentile") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))

(vpd_plot = rain_plot_func(data_thin$vpd_perc) +
        xlab("VPD percentile") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none"))


rcbi_plot + tbf_plot + erc_plot +
  icbi_plot + vs_plot + vpd_plot  + plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect") & theme(legend.position = 'bottom') & theme(legend.title=element_blank())

ggsave(datadir("figures/predictor_plots.png"), width = 9, height = 6)
``` 

Explore vegetation
```{r}
#fix Klamath - serpetine categories
data_thin = data_thin %>% 
  mutate(CT_lumped_codes4 = ifelse(BPS_NAME == 'Klamath-Siskiyou Upper Montane Serpentine Mixed Conifer Woodland' | 
                                     BPS_NAME == 'Klamath-Siskiyou Lower Montane Serpentine Mixed Conifer Woodland', 'Moist-Mixed Conifer', CT_lumped_codes4))


data_thin %>% group_by(CT_lumped_codes4) %>% summarise(n = n())

pal3 = c("#FFD662", "#244630", "#cb8c04","#46895D", "#B4C692", "#A06C18", "#4F3511", "#2A9D8F", "#264653")

library(formattable)

veg1 = data_thin %>% group_by(CT_lumped_codes4) %>%
  summarise(n = n()) %>% 
  mutate(freq = formattable::percent(n / sum(n)))

veg = data_thin %>% group_by(ecoregion, CT_lumped_codes4) %>%
  summarise(n = n()) %>% 
  mutate(freq = formattable::percent(n / sum(n)))
  
pal4 = c("#FFD662", "#244630", "#cb8c04","#46895D", "#B4C692", "#AA4731", "#4E0012", "#2A9D8F", "#264653")

vegplot = data_thin %>% group_by(CT_lumped_codes4, ecoregion) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x = n, y = ecoregion, fill = CT_lumped_codes4))+
  geom_bar(position = "fill", stat = "identity")+
  theme_bw(13)+
  scale_fill_manual(values = pal4, name = "forest type")+
  ylab("")+
  xlab("")+
  theme(legend.position = "bottom", legend.title=element_blank())+
  guides(fill = guide_legend(nrow = 3))
  
veg_table = data_thin %>% dplyr::select(CT_lumped_codes4, BPS_NAME) %>% distinct() %>% arrange(CT_lumped_codes4, BPS_NAME)
```

```{r}

#, "#264653","#2A9D8F","#883c0c", "#dbd053")
#climate space plot
# library(ggforce)
(vegclim_plot = slice(data_thin, sample(1:n())) %>% #shuffle rows for better visualization
  ggplot(aes(color = CT_lumped_codes4, x = PRISM_tmea, y = PRISM_ppt_)) +
  geom_point(alpha = .5, size = 1.5)+
  # geom_smooth()+
# 95% confidence level for a multivariate t-distribution
  theme_bw(13)+
  ylab("Precipitation (mm)") +
  xlab(expression("Temperature " ( degree*C)))+
  stat_ellipse(aes(linetype = ecoregion, x = PRISM_tmea, y = PRISM_ppt_), size = 1, type = "t", color = "black", show.legend = FALSE) +

  scale_linetype_manual(values = c(1,1,1,1))+
  scale_color_manual(values = pal4) + 
  theme(legend.position="NULL", legend.title=element_blank())+
  guides(colour = guide_legend(nrow = 2))+
  guides(color = guide_legend(override.aes = list(size = 3)))
  # annotate("text", x = 2, y = 300, label = "Northern Mountains", size = 4)+
  # annotate("text", x = 17, y = 350, label = "California Coast", size = 4)+
  # annotate("text", x = 15.5, y = 2700, label = "Western Mountains", size = 4)+
  # annotate("text", x = 8.8, y = 250, label = "Southwest", size = 4)
)



#save plot
(vegclim_plot + theme(legend.position = "none")) + vegplot + 
  plot_annotation(tag_levels = 'a') + plot_layout(guides = "auto")

ggsave(datadir("figures/ecoregion_overview3.png"), width = 11, height = 6)
```

### Explore correlations between variables
```{r}
library(purrr)

#keep only numeric variables
data_num = data_thin %>%
  keep(is.numeric)

# library(PerformanceAnalytics)
# cor(data_num, method = "pearson", use = "complete.obs")
# chart.Correlation(cor(data_num))

weather = data_thin %>% data.frame() %>%
  dplyr::select(erc_perc, vs_perc, vpd_perc, erc_02_21, vs_02_21, vpd_02_21, th_02_21, fm1000_02_, fm100_02_2, cbi_reburn, lat, long)

clim_topo = data_thin %>% data.frame() %>%
  dplyr::select(heat_load, PRISM_tmea, PRISM_ppt_, PRISM_vpdm, AET, ndvi, shrub_rebrn, tree_reburn, afg_rebrn, pfg_rebrn, TPI_1024, cbi_reburn, time_betwe, lat, long)

firehist = data_thin %>% data.frame() %>%
  dplyr::select(time_betwe, count_burn, cbi_reburn, cbi_initial, lat, long)

corplot = data_thin %>% data.frame() %>%
  dplyr::select(erc_perc, vs_perc, vpd_perc, heat_load, PRISM_tmea, PRISM_ppt_, PRISM_vpdm, AET, ndvi, shrub_rebrn, tree_reburn, afg_rebrn, pfg_rebrn, TPI_1024, cbi_reburn, time_betwe, count_burn, cbi_initial, cbi_reburn)
```

Plot correlations
```{r}
library(GGally)
library(corrplot)

corrplot(cor(weather),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

corrplot(cor(clim_topo),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

corrplot(cor(firehist),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

#save corrplot
png(height=800, width=800, file=datadir("figures/corrplot.png"))

corrplot(cor(corplot),  method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)
dev.off()

#strongest correlations between weather variables, fuels variables, and lat/long
#> selected precip and temp over AET and PRISM_vpd, and erc over fm100/1000
```





### Fit separate models for each ecoregion

```{r}
#scale numeric predictors for model fit - don't scale time between t or cbi or erc percentiles to maintain interpretation
preds_2scale = c("heat_load", "PRISM_tmea", "PRISM_ppt_", "TPI_1024", "fm100_02_2", "th_02_21", "vpd_02_21", "erc_02_21", "PRISM_vpdm", "AET", "fm1000_02_", "ndvi", "shrub_rebrn", "tree_reburn", "afg_rebrn", "pfg_rebrn", "cbi_initial", "time_betwe", "erc_perc", "vpd_perc", "vs_perc")

data_thin = data.frame(data_thin)

scale_df = function(df){
  data_scale = scale(df[(names(df) %in% preds_2scale)], scale = TRUE, center = TRUE)
  data_scale_all = cbind(data_scale, df[!(names(df) %in% preds_2scale)])
  return(data_scale_all)
}

data_thin_scale = data_thin %>% scale_df()

#create new dfs for each ecoregion and scale each ecoregion separately 
data_lowmod_ca = data_thin %>% filter(ecoregion == "California Coast") %>% scale_df()
data_lowmod_nm = data_thin %>% filter(ecoregion == "Northern Mountains") %>% scale_df()
data_lowmod_sw = data_thin %>% filter(ecoregion == "Southwest") %>% scale_df()
data_lowmod_wm = data_thin %>% filter(ecoregion == "Western Mountains") %>% scale_df()

unscale_df = data_thin %>%
  group_by(ecoregion) %>%
  summarise(timebetwe_sd = sd(time_betwe),
            timebetwe_mean = mean(time_betwe),
            cbi_initial_sd = sd(cbi_initial),
            cbi_initial_mean = mean(cbi_initial),
            erc_perc_sd = sd(erc_perc),
            erc_perc_mean = mean(erc_perc),
            vpd_perc_sd = sd(vpd_perc),
            vpd_perc_mean = mean(vpd_perc),
            vs_perc_sd = sd(vs_perc),
            vs_perc_mean = mean(vs_perc))
```



Test for spatial autocorrelation
```{r}
library(ape)
inv_dist = with(data_lowmod_ca, 1/dist(cbind(lat, long), diag = TRUE, upper = TRUE))
inv_dist = as.matrix(inv_dist)
ape::Moran.I(data_lowmod_ca$cbi_reburn, weight = inv_dist, scaled = TRUE)

#observed = 0.15  - CA; 0.11 - NM; 0.17 for SW, 0.07 for wm
# While statistically significant, there actually isn’t too much going on, though it may be enough to warrant dealing with in some fashion.

```


## Model fit/selection with mgcv 
>https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/gam.selection.html 
>https://stats.stackexchange.com/questions/519433/gam-and-multiple-continuous-continuous-interactions-tensor-smooths #helpful for interpretting and fitting interactions (ti vs te vs s)

>all about gams: https://m-clark.github.io/generalized-additive-models/application.html#gam-1


```{r}

fit_ca = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_ppt_, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_tmea, k = 5, bs = "cs") +
                  ti(time_betwe, by = vpd_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = vs_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = erc_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = cbi_initial, k = 5, bs = "cs"),
             data = data_lowmod_ca,
             select = T,
             gamma = 1.4,
             method = "REML") 

fit_nm = gam(cbi_reburn ~ count_burn +
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_ppt_, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_tmea, k = 5, bs = "cs") +
                  ti(time_betwe, by = vpd_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = vs_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = erc_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = erc_perc, k = 5, bs = "cs"),
             data = data_lowmod_nm,
             select = T,
             gamma = 1.4,
             method = "REML") 

fit_sw = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_ppt_, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_tmea, k = 5, bs = "cs") +
                  ti(time_betwe, by = vpd_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = vs_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = erc_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = cbi_initial, k = 5, bs = "cs"),
             data = data_lowmod_sw,
             select = T,
             gamma = 1.4,
             method = "REML") 

fit_wm = gam(cbi_reburn ~ count_burn + 
                  s(PRISM_ppt_, k = 5, bs = "cs") +
                  s(PRISM_tmea, k = 5, bs = "cs") +
                  s(cbi_initial, k = 5, bs = "cs") +
                  s(time_betwe, k = 5, bs = "cs") +
                  s(erc_perc, k = 5, bs = "cs") +
                  s(vs_perc, k = 5, bs = "cs") +
                  s(vpd_perc, k = 5, bs = "cs") +
                  s(heat_load, k = 5, bs = "cs") +
                  s(TPI_1024, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_ppt_, k = 5, bs = "cs") +
                  ti(time_betwe, by = PRISM_tmea, k = 5, bs = "cs") +
                  ti(time_betwe, by = vpd_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = vs_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = erc_perc, k = 5, bs = "cs") +
                  ti(time_betwe, by = cbi_initial, k = 5, bs = "cs"),
             data = data_lowmod_wm,
             select = T,
             gamma = 1.4,
             method = "REML") 
```



Check residuals
```{r}
library(mgcViz)
#k = 5
gam.check(fit_ca) 
gam.check(fit_nm) 
gam.check(fit_sw) 
gam.check(fit_wm)
#- still some significance in the residuals tests, but edf is below 6 so there are likely other issues with the models? thinning helped this a bit

check(getViz(fit_ca))
check(getViz(fit_nm)) 
check(getViz(fit_sw))
check(getViz(fit_wm))

#including lat long did not improve the residuals and it is highly correlated with other predictors, so leaving out


#check for concurvity - non-linearly related variables; Ignore interactions? 
concurv_ca = concurvity(fit_ca, full = FALSE)  
concurv_ca = concurv_ca$observed

resid_ca = residuals(fit_ca) 
resid_nm = residuals(fit_nm)
resid_sw = residuals(fit_sw)
resid_wm = residuals(fit_wm)

#add residuals to dfs
data_lowmod_ca$resids = resid_ca
data_lowmod_nm$resids = resid_nm
data_lowmod_sw$resids = resid_sw
data_lowmod_wm$resids = resid_wm

#plot residuals against covariates
plot_resids_cov = function(data){
  resid_plot = data %>%
  dplyr::select(PRISM_ppt_, TPI_1024,
                PRISM_tmea, cbi_initial, time_betwe,
                erc_perc, vs_perc, vpd_perc, heat_load, resids) %>%
  pivot_longer(-resids, names_to = "key", values_to = "value") %>% #convert to long format for fast plotting
  ggplot(aes(x = scale(value), y = resids)) +
  facet_wrap(~ key, scales = "free") +
  geom_smooth(formula = y ~ s(x, bs = "cs", k = 6)) +
  theme_bw(13)
  return(resid_plot)
}

plot_resids_cov(data_lowmod_ca)
plot_resids_cov(data_lowmod_nm)
plot_resids_cov(data_lowmod_sw)
plot_resids_cov(data_lowmod_wm)
#covaraite/residual relationships look good! 



#investigate spatial autocorrelation - must be decimal degrees? 
library(gstat) 

#project data
project_func = function(df){
  xy <- df[,c("long","lat")]
  spdf <- SpatialPointsDataFrame(coords = xy, data = df,
                             proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84
                                               +towgs84=0,0,0"))
  return(spdf)
}

ca_proj = project_func(data_lowmod_ca)
sw_proj = project_func(data_lowmod_sw)
nm_proj = project_func(data_lowmod_nm)
wm_proj = project_func(data_lowmod_wm)

vca = plot(variogram(resid_ca~1, data=ca_proj, cutoff = 5), main = "cali coast")
vnm = plot(variogram(resid_nm~1, data=nm_proj, cutoff = 5), main = "northern mnts")
vsw = plot(variogram(resid_sw~1, data=sw_proj, cutoff = 5), main = "southwest")
vwm = plot(variogram(resid_wm~1, data=wm_proj, cutoff = 5), main = "Western mnts")

grid.arrange(vca, vsw, vnm, vwm, nrow= 2, ncol=2)

#check moran's I
pixel.dists <- as.matrix(dist(cbind(ca_proj$long, ca_proj$lat)))
pixel.dists.inv <- 1/pixel.dists
diag(pixel.dists.inv) <- 0

pixel.dists <- as.matrix(dist(cbind(wm_proj$long, ca_proj$lat)))
pixel.dists.inv <- 1/pixel.dists
diag(pixel.dists.inv) <- 0


print(Moran.I(resid_wm,pixel.dists.inv))
```

Summarise models
```{r}
library(texreg) #for creating easy tables from gam output
extract(fit_ca)
extract(fit_wm)
extract(fit_sw)
extract(fit_nm)

summary(fit_ca)

texreg(list(fit_ca,fit_nm,fit_sw, fit_wm),
       caption="GAM output by ecoregion",
       dcolumn=FALSE,
       model.names=c("California Coast","Northern Mountains","Southwest", "Western Mountains"))


```

visualize interactions
```{r}
library(mgcViz)
library(rgl)

plot(fit_ca)
plot(fit_nm)
plot(fit_sw)
plot(fit_wm)
```


### Predict

Just time between
```{r}

preds_func = function(gam_fit, df){
  preds = predict_gam(gam_fit, length_out = 30, values = list(PRISM_ppt_ = 0, PRISM_tmea = 0, cbi_initial =0, time_betwe = seq(min(df$time_betwe), max(df$time_betwe), by = .05), erc_perc = 0, vs_perc = 0, vpd_perc = 0, heat_load = 0, TPI_1024 = 0, count_burn = 2)) 
  return(preds)
}

ca_preds = preds_func(fit_ca, data_lowmod_ca)
nm_preds = preds_func(fit_nm, data_lowmod_nm)
sw_preds = preds_func(fit_sw, data_lowmod_sw)
wm_preds = preds_func(fit_wm, data_lowmod_wm)

ca_preds$time_betwe_unscale = (ca_preds$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])
nm_preds$time_betwe_unscale = (nm_preds$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])
sw_preds$time_betwe_unscale = (sw_preds$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])
wm_preds$time_betwe_unscale = (wm_preds$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])

ca_preds$cbi_reburn = ca_preds$fit
nm_preds$cbi_reburn = nm_preds$fit
sw_preds$cbi_reburn = sw_preds$fit
wm_preds$cbi_reburn = wm_preds$fit


plot_preds = function(preds){
  
  plot = preds %>%
  ggplot(aes(time_betwe_unscale, fit)) +
    geom_smooth_ci(size = 1.05, legend = NA)+
  theme_bw(13) +
  theme(legend.title = element_blank())+
  ylim(0.5, 2.5) +
    xlab("time between fires (yrs)") +
    ylab("rebrun severity (CBI)")
  
  return(plot)
}

pca1 = plot_preds(ca_preds) + ggtitle("California Coast")
psw1 = plot_preds(sw_preds) + ggtitle("Southwest")
pnm1 = plot_preds(nm_preds) + ggtitle("Northern Mountains")
pwm1 = plot_preds(wm_preds) + ggtitle("Western Mountains")

pwm1 + pnm1 + pca1 + psw1 + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/gams_average.png'), height = 5, width = 5)
```


Weather predictions - scaled
```{r}
#prediction df with most variables set to mean = 0
preds_dfw_func = function(df, fit, vpd_mean, vpd_sd, erc_mean, erc_sd){
  df_w97 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      #set weather variables (vpd/erc) to 98th percentile
                      vpd_perc = as.numeric((.98-vpd_mean)/vpd_sd),
                      erc_perc = as.numeric((.98-erc_mean)/erc_sd),
                      vs_perc = 0,
                      #set time between fires to full range of time between
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      weather = "extreme")
  
  df_w90 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = as.numeric((.95-vpd_mean)/vpd_sd),
                      erc_perc = as.numeric((.95-erc_mean)/erc_sd),
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      weather = "very high")

  df_w80 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = as.numeric((.85-vpd_mean)/vpd_sd),
                      erc_perc = as.numeric((.85-erc_mean)/erc_sd),
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      weather = "high")
  
  
  
  data_w = rbind(df_w97, df_w80, df_w90)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE)
  
  #add predictions and CIs to df for plotting
  data_w$cbi_reburn = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  #make weather a factor for plotting
  data_w$weather = factor(data_w$weather, levels = c("high", "very high", "extreme"))
  
  return(data_w)
}

#predict to weather values (calcualte scaled values of .97, .9, and .8 percentiles)
preds_w_ca <- preds_dfw_func(df = data_lowmod_ca, fit = fit_ca,
                             vpd_mean = unscale_df[1, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[1, 'vpd_perc_sd'],
                             erc_mean = unscale_df[1, 'erc_perc_mean'],
                             erc_sd = unscale_df[1, 'erc_perc_sd'])

preds_w_nm <- preds_dfw_func(df = data_lowmod_nm, fit = fit_nm,
                             vpd_mean = unscale_df[2, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[2, 'vpd_perc_sd'],
                             erc_mean = unscale_df[2, 'erc_perc_mean'],
                             erc_sd = unscale_df[2, 'erc_perc_sd'])

preds_w_sw <- preds_dfw_func(df = data_lowmod_sw, fit = fit_sw,
                             vpd_mean = unscale_df[3, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[3, 'vpd_perc_sd'],
                             erc_mean = unscale_df[3, 'erc_perc_mean'],
                             erc_sd = unscale_df[3, 'erc_perc_sd'])

preds_w_wm <- preds_dfw_func(df = data_lowmod_wm, fit = fit_wm,
                             vpd_mean = unscale_df[4, 'vpd_perc_mean'],
                             vpd_sd = unscale_df[4, 'vpd_perc_sd'],
                             erc_mean = unscale_df[4, 'erc_perc_mean'],
                             erc_sd = unscale_df[4, 'erc_perc_sd'])

#add in unscaled time_between for plotting
preds_w_ca$time_betwe_unscale = (preds_w_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])

preds_w_nm$time_betwe_unscale = (preds_w_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_w_sw$time_betwe_unscale = (preds_w_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_w_wm$time_betwe_unscale = (preds_w_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```

wind speed predictions - scaled
```{r}
#prediction df with most variables set to mean = 0
preds_dfvs_func = function(df, fit, vs_mean, vs_sd){
  df_w97 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      #set wind speed to 75th percentile
                      vs_perc = as.numeric((.75-vs_mean)/vs_sd),
                      vpd_perc = 0, erc_perc = 0,
                      #set time between fires to full range 1-25 years
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      windspeed = "high")
  
  df_w90 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      #set windspeed to median
                      vs_perc = as.numeric((.5-vs_mean)/vs_sd),
                      vpd_perc = 0, erc_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      windspeed = "moderate")

  df_w80 = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      cbi_initial = 0, 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      #set wind speed to 25th percentile
                      vs_perc = as.numeric((.25-vs_mean)/vs_sd),
                      vpd_perc = 0, erc_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      windspeed = "low")
  
  data_w = rbind(df_w97, df_w80, df_w90)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE)
  
  #add predictions and CIs to df for plotting
  data_w$cbi_reburn = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  #make weather a factor for plotting
  data_w$windspeed = factor(data_w$windspeed, levels = c("low", "moderate", "high"))
  
  return(data_w)
}

#predict reburn severity for 3 levels of wind speed
preds_vs_ca <- preds_dfvs_func(df = data_lowmod_ca, fit = fit_ca,
                             vs_mean = unscale_df[1, 'vs_perc_mean'],
                             vs_sd = unscale_df[1, 'vs_perc_sd'])

preds_vs_nm <- preds_dfvs_func(df = data_lowmod_nm, fit = fit_nm,
                             vs_mean = unscale_df[2, 'vs_perc_mean'],
                             vs_sd = unscale_df[2, 'vs_perc_sd'])

preds_vs_sw <- preds_dfvs_func(df = data_lowmod_sw, fit = fit_sw,
                             vs_mean = unscale_df[3, 'vs_perc_mean'],
                             vs_sd = unscale_df[3, 'vs_perc_sd'])

preds_vs_wm <- preds_dfvs_func(df = data_lowmod_wm, fit = fit_wm,
                             vs_mean = unscale_df[4, 'vs_perc_mean'],
                             vs_sd = unscale_df[4, 'vs_perc_sd'])

#add in unscaled time_between for plotting
preds_vs_ca$time_betwe_unscale = (preds_vs_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])

preds_vs_nm$time_betwe_unscale = (preds_vs_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_vs_sw$time_betwe_unscale = (preds_vs_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_vs_wm$time_betwe_unscale = (preds_vs_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```


CBI predictions - scaled
```{r}
#prediction df with most variables set to mean = 0
preds_dfcbi_func = function(df, fit, cbi_mean, cbi_sd){
  df_lowcbi = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      #set cbi initial to 1
                      cbi_initial = as.numeric((1.0-cbi_mean)/cbi_sd), 
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = 0,
                      erc_perc = 0,
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      cbii_cat = "low")

  
  df_modcbi = data.frame(count_burn = 2, #unscaled
                      PRISM_ppt_ = 0, 
                      PRISM_tmea = 0,
                      #set cbi initial to 2.25
                      cbi_initial = as.numeric((2.25-cbi_mean)/cbi_sd),
                      heat_load = 0, 
                      TPI_1024 = 0,
                      vpd_perc = 0,
                      erc_perc = 0,
                      vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                      cbii_cat = "mod")


  
  data_w = rbind(df_lowcbi, df_modcbi)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_w, se.fit = TRUE)
  
  #add predictions and CIs to df for plotting
  data_w$cbi_reburn = p$fit
  data_w$upr = p$fit + (2 * p$se.fit)
  data_w$lwr = p$fit - (2 * p$se.fit)
  
  
  return(data_w)
}

#predict to weather values (calculate scaled values of .97, .9, and .8 percentiles)
preds_cbi_ca <- preds_dfcbi_func(df = data_lowmod_ca, fit = fit_ca,
                             cbi_mean = unscale_df[1, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[1, 'cbi_initial_sd'])

preds_cbi_nm <- preds_dfcbi_func(df = data_lowmod_nm, fit = fit_nm,
                             cbi_mean = unscale_df[2, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[2, 'cbi_initial_sd'])

preds_cbi_sw <- preds_dfcbi_func(df = data_lowmod_sw, fit = fit_sw,
                             cbi_mean = unscale_df[3, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[3, 'cbi_initial_sd'])

preds_cbi_wm <- preds_dfcbi_func(df = data_lowmod_wm, fit = fit_wm,
                             cbi_mean = unscale_df[4, 'cbi_initial_mean'],
                             cbi_sd = unscale_df[4, 'cbi_initial_sd'])

#add in unscaled time_between for plotting
preds_cbi_ca$time_betwe_unscale = (preds_cbi_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])

preds_cbi_nm$time_betwe_unscale = (preds_cbi_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_cbi_sw$time_betwe_unscale = (preds_cbi_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_cbi_wm$time_betwe_unscale = (preds_cbi_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```


climate predictions
```{r}
# #find west wide climate quantiles


#prediction df with variables set to mean = 0
preds_dfc_func = function(df, fit){
  #cool wet - high productivity
  df_coolwet = data.frame(count_burn = 2, 
                        #set climate vars
                       PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.75),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.25),
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc  = 0, erc_perc = 0,
                       vs_perc = 0,
                      time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                       climate = "coldwet")

  #hot dry - low productivity
  df_hotdry = data.frame(count_burn = 2, 
                        #set climate vars
                       PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.25),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.75),
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc  = 0, erc_perc = 0,
                       vs_perc = 0,
                       time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                       climate = "hotdry")

  
    #hot dry - low productivity
  df_warmwet = data.frame(count_burn = 2, 
                        #set climate vars
                       PRISM_ppt_ = quantile(df$PRISM_ppt_, 0.75),
                       PRISM_tmea = quantile(df$PRISM_tmea, 0.75),
                       cbi_initial = 0, heat_load = 0, TPI_1024 = 0,
                       vpd_perc  = 0, erc_perc = 0,
                       vs_perc = 0,
                       time_betwe = seq(min(df$time_betwe), max(df$time_betwe), length.out = 36),
                       climate = "hotwet")
  
  
  
  data_c = rbind(df_coolwet, df_hotdry, df_warmwet)
  
  #predict to new data
  p = predict.gam(fit, newdata = data_c, se.fit = TRUE)
  
  #add predictions to df for plotting
  data_c$cbi_reburn = p$fit
  data_c$upr = p$fit + (2 * p$se.fit)
  data_c$lwr = p$fit - (2 * p$se.fit)
  return(data_c)
}

preds_c_ca <- preds_dfc_func(df = data_lowmod_ca, fit = fit_ca)
preds_c_nm <- preds_dfc_func(df = data_lowmod_nm, fit = fit_nm)
preds_c_sw <- preds_dfc_func(df = data_lowmod_sw, fit = fit_sw)
preds_c_wm <- preds_dfc_func(df = data_lowmod_wm, fit = fit_wm)

#add in unscaled time_between for plotting
preds_c_ca$time_betwe_unscale = (preds_c_ca$time_betwe * as.numeric(unscale_df[1, "timebetwe_sd"])) + as.numeric(unscale_df[1, "timebetwe_mean"])

preds_c_nm$time_betwe_unscale = (preds_c_nm$time_betwe * as.numeric(unscale_df[2, "timebetwe_sd"])) + as.numeric(unscale_df[2, "timebetwe_mean"])

preds_c_sw$time_betwe_unscale = (preds_c_sw$time_betwe * as.numeric(unscale_df[3, "timebetwe_sd"])) + as.numeric(unscale_df[3, "timebetwe_mean"])

preds_c_wm$time_betwe_unscale = (preds_c_wm$time_betwe * as.numeric(unscale_df[4, "timebetwe_sd"])) + as.numeric(unscale_df[4, "timebetwe_mean"])
```

Exploring weird trend with cool wet climate in the NM
```{r}
ggplot(data = data_lowmod_nm, aes(x = PRISM_tmea, y = PRISM_ppt_)) +
  geom_point()+
  geom_vline(xintercept = quantile(data_lowmod_nm$PRISM_tmea, .25))+
  geom_hline(yintercept = quantile(data_lowmod_nm$PRISM_ppt_, .75))

ggplot(data = data_lowmod_nm, aes(x = PRISM_tmea, y = cbi_reburn)) +
  geom_point()+
  geom_smooth()+
  geom_vline(xintercept = quantile(data_lowmod_nm$PRISM_tmea, .25))

ggplot(data = data_lowmod_nm, aes(x = PRISM_ppt_, y = cbi_reburn)) +
  geom_point()+
  geom_smooth()+
  geom_vline(xintercept = quantile(data_lowmod_nm$PRISM_ppt_, .75))
```


### Determine when slopes flatten for each prediction

CBI
```{r}
library(emmeans)
library(modelbased)
#CBI

#calculate time between slopes for 2 levels of cbi
ca_cbi_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_ca$cbi_initial)"), length = 36, ci=0.95)
nm_cbi_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_nm$cbi_initial)"), length = 36, ci=0.95)
sw_cbi_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_sw$cbi_initial)"), length = 36, ci=0.95)
wm_cbi_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "cbi_initial = unique(preds_cbi_wm$cbi_initial)"), length = 36, ci=0.95)


#add to predictor df
preds_cbi_ca2 = preds_cbi_ca %>%
    left_join(ca_cbi_slopes, by = c("time_betwe", "cbi_initial"))

preds_cbi_nm2 = preds_cbi_nm %>%
    left_join(nm_cbi_slopes, by = c("time_betwe", "cbi_initial"))

preds_cbi_sw2 = preds_cbi_sw %>%
    left_join(sw_cbi_slopes, by = c("time_betwe", "cbi_initial"))

preds_cbi_wm2 = preds_cbi_wm %>%
    left_join(wm_cbi_slopes, by = c("time_betwe", "cbi_initial"))

#plot ggplot
ggplot(preds_cbi_wm2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(cbii_cat))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()
```

Weather
```{r}
#calculate time between slopes for 3 levels of weather (erc & vpd)
ca_w_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_ca$erc_perc)", "vpd_perc = unique(preds_w_ca$vpd_perc)"), length = 36, ci=0.95)

#calculate time between slopes for 3 levels of weather (erc & vpd)
nm_w_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_nm$erc_perc)", "vpd_perc = unique(preds_w_nm$vpd_perc)"), length = 36, ci=0.95)

#calculate time between slopes for 3 levels of weather (erc & vpd)
sw_w_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_sw$erc_perc)", "vpd_perc = unique(preds_w_sw$vpd_perc)"), length = 36, ci=0.95)

#calculate time between slopes for 3 levels of weather (erc & vpd)
wm_w_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "erc_perc = unique(preds_w_wm$erc_perc)", "vpd_perc = unique(preds_w_wm$vpd_perc)"), length = 36, ci=0.95)

#add slope coefficients to predictor dfs
preds_w_ca2 = preds_w_ca %>%
    left_join(ca_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

preds_w_nm2 = preds_w_nm %>%
    left_join(nm_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

preds_w_sw2 = preds_w_sw %>%
    left_join(sw_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

preds_w_wm2 = preds_w_wm %>%
    left_join(wm_w_slopes, by = c("time_betwe", "erc_perc", "vpd_perc"))

#plot
ggplot(preds_w_sw2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(weather))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()
```

wind
```{r}
#calculate time between slopes for 3 levels of windspeed
ca_vs_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_ca$vs_perc)"), length = 36, ci=0.95)
nm_vs_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_nm$vs_perc)"), length = 36, ci=0.95)
sw_vs_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_sw$vs_perc)"), length = 36, ci=0.95)
wm_vs_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "vs_perc = unique(preds_vs_wm$vs_perc)"), length = 36, ci=0.95)

plot(modelbased::visualisation_recipe(ca_vs_slopes))+theme_bw()
plot(modelbased::visualisation_recipe(nm_vs_slopes))+theme_bw()
plot(modelbased::visualisation_recipe(sw_vs_slopes))+theme_bw()
plot(modelbased::visualisation_recipe(wm_vs_slopes))+theme_bw()

#add to predictor df
preds_vs_ca2 = preds_vs_ca %>%
    left_join(ca_vs_slopes, by = c("time_betwe", "vs_perc"))

preds_vs_nm2 = preds_vs_nm %>%
    left_join(nm_vs_slopes, by = c("time_betwe", "vs_perc"))

preds_vs_sw2 = preds_vs_sw %>%
    left_join(sw_vs_slopes, by = c("time_betwe", "vs_perc"))

preds_vs_wm2 = preds_vs_wm %>%
    left_join(wm_vs_slopes, by = c("time_betwe", "vs_perc"))

#plot ggplot
ggplot(preds_vs_sw2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(windspeed))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()

```

Climate
```{r}
#Climate 
#calculate time between slopes for 3 levels of climate (temp & precip) and mean
ca_c_slopes <- modelbased::estimate_slopes(fit_ca, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_ca$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_ca$PRISM_ppt_)"), length = 36, ci=0.95)

nm_c_slopes <- modelbased::estimate_slopes(fit_nm, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_nm$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_nm$PRISM_ppt_)"), length = 36, ci=0.95)

sw_c_slopes <- modelbased::estimate_slopes(fit_sw, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_sw$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_sw$PRISM_ppt_)"), length = 36, ci=0.95)

wm_c_slopes <- modelbased::estimate_slopes(fit_wm, trend = "time_betwe", at = c("time_betwe", "PRISM_tmea = unique(preds_c_wm$PRISM_tmea)", "PRISM_ppt_ = unique(preds_c_wm$PRISM_ppt_)"), length = 36, ci=0.95)

#add slope coefficients to predictor dfs
preds_c_ca2 = preds_c_ca %>%
    left_join(ca_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

preds_c_nm2 = preds_c_nm %>%
    left_join(nm_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

preds_c_sw2 = preds_c_sw %>%
    left_join(sw_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

preds_c_wm2 = preds_c_wm %>%
    left_join(wm_c_slopes, by = c("time_betwe", "PRISM_tmea", "PRISM_ppt_"))

#plot
ggplot(preds_c_wm2, aes(x = time_betwe_unscale*10, y = Coefficient, color = as.factor(climate))) +
  geom_line(se = FALSE)+
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high), alpha = 0.2)+
  theme_bw()
```



### Prep predictor DF for plotting


Truncate predictor data frames where there is no data on low and high end of time between - create bins for prediciton categories
```{r}

#create new variables for each interaction category
bins = data_thin %>%
  mutate(cbi_bins = case_when(cbi_initial <= 1.5 ~ "low",
                             cbi_initial > 1.5 ~ "mod"),
         vs_bins = case_when(vs_perc <= .375 ~ "low",
                             vs_perc > .375 & vs_perc <= .625 ~ "moderate",
                             vs_perc > .625 ~ "high"),
         weather_bins = case_when(erc_perc > .97 & vpd_perc > .97 ~ "extreme",
                             erc_perc <= .9 & vpd_perc <= .9 ~ "high",
                             erc_perc > .9 & vpd_perc > .9 &
                              erc_perc <= .97 & vpd_perc <= .97 ~ "very high"),
         time_fact = as.factor(time_betwe))

#weather bins
weather_bins = bins %>% group_by(ecoregion, weather_bins) %>%
  summarise(count = n(),
            min_time = min(time_betwe), #get minimum and maximum time between for each predictor combination
            max_time = max(time_betwe)) %>%
  na.omit()

#wind speed bins
vs_bins = bins %>% group_by(ecoregion, vs_bins) %>%
  summarise(count = n(),
            min_time = min(time_betwe),
            max_time = max(time_betwe))

#cbi bins
cbi_bins = bins %>% group_by(ecoregion, cbi_bins) %>%
  summarise(count = n(),
            min_time = min(time_betwe),
            max_time = max(time_betwe))

##climate bins
climate_bins = data_thin %>% group_by(ecoregion) %>%
  mutate(climate_bin = case_when(PRISM_ppt_ <= quantile(PRISM_ppt_, .5) &
              PRISM_tmea > quantile(PRISM_tmea, .5) ~ "hotdry",
              PRISM_ppt_ > quantile(PRISM_ppt_, .5) &
              PRISM_tmea <= quantile(PRISM_tmea, .5) ~ "coldwet",
              PRISM_ppt_ > quantile(PRISM_ppt_, .5) &
              PRISM_tmea > quantile(PRISM_tmea, .5) ~ "hotwet"))

climate_bins2 = climate_bins %>% group_by(ecoregion, climate_bin) %>%
  summarise(count = n(),
            min_time = min(time_betwe),
            max_time = max(time_betwe)) %>%
 na.omit()

```

#### Truncate predictor data frames where there is no data on low and high end of time between
```{r}
#filter out values that fall outside the time between data range
preds_w_ca_trunc = preds_w_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= .9 & extreme <= 3.2 |
           high >= .1 & high <= 3.4 |
           `very high` >= .5 & `very high` <= 3.5) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "weather",
         ecoregion = "California Coast") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

preds_w_nm_trunc = preds_w_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= .1 & extreme <= 3.6 |
           high >= .1 & high <= 3.4 |
           `very high` >= .1 & `very high` <= 3.6) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "weather",
         ecoregion = "Northern Mountains") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

preds_w_sw_trunc = preds_w_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= .1 & extreme <= 2.7 |
           high >= .1 & high <= 2.9 |
           `very high` >= .3 & `very high` <= 2.6) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "weather",
         ecoregion = "Southwest") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

preds_w_wm_trunc = preds_w_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = weather, values_from = time_betwe_round) %>%
  filter(extreme >= .2 & extreme <= 3.6 |
           high >= .1 & high <= 3.4 |
           `very high` >= .3 & `very high` <= 3.4) %>%
  pivot_longer(extreme:`very high`, names_to = "weather", values_to = "time_betwe_round") %>%
  na.omit() %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Western Mountains",
         cat = "weather") %>%
  mutate(weather = factor(weather, levels = c("high", "very high", "extreme")))

```

wind speed truncated dfs
```{r}
#filter out values that fall outside the time between data range
preds_vs_ca_trunc = preds_vs_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= .1 & high <= 3.5 |
           low >= .1 & low <= 3.3 |
           moderate >= .1 & moderate <= 3.4) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "windspeed",
         ecoregion = "California Coast") %>%
  na.omit()

preds_vs_nm_trunc = preds_vs_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= .1 & high <= 3.3 |
           low >= .1 & low <= 3.6 |
           moderate >= .1 & moderate <= 3.4) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "windspeed",
         ecoregion = "Northern Mountains") %>%
  na.omit()

preds_vs_sw_trunc = preds_vs_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= .1 & high <= 3.3 |
           low >= .1 & low <= 2.7 |
           moderate >= .1 & moderate <= 2.9) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
   mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "windspeed",
         ecoregion = "Southwest") %>%
  na.omit()

preds_vs_wm_trunc = preds_vs_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = windspeed, values_from = time_betwe_round) %>%
  filter(high >= .2 & high <= 3.4 |
           low >= .1 & low <= 3.6 |
           moderate >= .1 & moderate <= 3.6) %>%
  pivot_longer(high:moderate, names_to = "windspeed", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Western Mountains",
         cat = "windspeed") %>%
  na.omit()
```

CBI truncated dfs
```{r}
preds_cbi_ca_trunc = preds_cbi_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= .1 & low <= 3.5 |
           mod >= .1 & mod <= 3.5) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "cbi",
         ecoregion = "California Coast") %>%
  na.omit()

preds_cbi_nm_trunc = preds_cbi_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= .1 & low <= 3.6 |
           mod >= .1 & mod <= 3.6) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "cbi",
         ecoregion = "Northern Mountains") %>%
  na.omit()

preds_cbi_sw_trunc = preds_cbi_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= .1 & low <= 3.3 |
           mod >= .1 & mod <= 2.9) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Southwest",
         cat = "cbi") %>%
  na.omit()

preds_cbi_wm_trunc = preds_cbi_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = cbii_cat, values_from = time_betwe_round) %>%
  filter(low >= .1 & low <= 3.6 |
           mod >= .1 & mod <= 3.6) %>%
  pivot_longer(low:mod, names_to = "cbii_cat", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         ecoregion = "Western Mountains",
         cat = "cbi") %>%
  na.omit()

```

Truncate climate predictions
```{r}
preds_c_ca_trunc = preds_c_ca2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
  filter(coldwet >= .3 & coldwet <= 3.3 |
           hotdry >= .6 & hotdry <= 3.5 |
           hotwet >= .8 & hotwet <= 3.5) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "California Coast") %>%
    na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

preds_c_nm_trunc = preds_c_nm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
  filter(coldwet >= .1 & coldwet <= 3.6 |
           hotdry >= .1 & hotdry <= 3.3 |
           hotwet >= .2 & hotwet <= 3.4) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "Northern Mountains") %>%
    na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

preds_c_sw_trunc = preds_c_sw2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
  filter(coldwet >= .1 & coldwet <= 2.5 |
           hotdry >= .1 & hotdry <= 2.7 |
           hotwet >= .1 & hotwet <= 2.7) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "Southwest") %>%
    na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

preds_c_wm_trunc = preds_c_wm2 %>% 
  mutate(time_betwe_round = round(time_betwe_unscale, 1)) %>%
  pivot_wider(names_from = climate, values_from = time_betwe_round) %>%
  filter(coldwet >= .1 & coldwet <= 3.6 |
           hotdry >= .3 & hotdry <= 3.6 |
           hotwet >= .2 & hotwet <= 3.6) %>%
  pivot_longer(coldwet:hotwet, names_to = "climate", values_to = "time_betwe_round") %>%
  mutate(sig = case_when(p < 0.05 ~ 'sig',
                         p >= 0.05 ~ 'non-sig'),
         cat = "climate",
         ecoregion = "Western Mountains") %>%
  na.omit() %>%
  mutate(climate = factor(climate, levels = c("coldwet", "hotdry", "hotwet")))

```


### Find points where slopes flatten (lower CI crosses 0)
```{r}
#combine all predictions
preds = bind_rows(preds_c_ca_trunc, preds_c_nm_trunc, preds_c_sw_trunc, preds_c_wm_trunc, preds_vs_ca_trunc, preds_vs_nm_trunc, preds_vs_sw_trunc, preds_vs_wm_trunc, preds_w_ca_trunc, preds_w_nm_trunc, preds_w_sw_trunc, preds_w_wm_trunc, preds_cbi_ca_trunc, preds_cbi_nm_trunc, preds_cbi_sw_trunc, preds_cbi_wm_trunc)

flat_slope_preds = preds %>% filter(CI_low <= 0) %>%
  mutate(weather = as.character(weather),
         climate = as.character(climate)) %>% #convert from factor to remove NAs
  mutate_if(is.character, ~replace_na(.,"")) %>%
  mutate(cat2 = paste(climate, cbii_cat, weather, windspeed, sep = ""),
         cat3 = paste(cat, ecoregion, sep = "-")) %>%
  group_by(cat3, cat2) %>% 
  #extract the time between fires when the lower 95% CI first crosses 0
  slice(which.min(time_betwe_round)) %>% 
  mutate(x = time_betwe_round*10, 
         y = cbi_reburn) %>%
  dplyr::select(cat3, cat2, x, y)

#update the flat slope CIs so that it's the first time the CI crosses 0 after being above 0, so it can't be at TBF = 1. I did this manually by finding the first time the lower CI was at or below 0 after the slope was significant in the preds dataframe
flat_slope_preds2 = flat_slope_preds
  
  
flat_slope_preds2[6, 'x'] = 24 #cbi-sw moderate
flat_slope_preds2[6, 'y'] = 1.6224019 #cbi-sw moderate

flat_slope_preds2[13, 'x'] = 26 #cilmate - nm hotdry
flat_slope_preds2[13, 'y'] = 1.7824215 #cilmate - nm hotdry

flat_slope_preds2[37, 'x'] = 25 #windspeed - nm low
flat_slope_preds2[37, 'y'] = 1.8479243 #windspeed - nm low

flat_slope_preds2[40, 'x'] = 16 #windspeed - sw low
flat_slope_preds2[40, 'y'] = 1.1684846 #windspeed - sw low

flat_slope_preds2[41, 'x'] = 20 #windspeed - sw mod
flat_slope_preds2[41, 'y'] = 1.2556077 #windspeed - sw mod

#extract the time between fires when the slope first crosses 0
flat_slope_fit = preds %>% filter(Coefficient <= 0) %>%
  mutate(weather = as.character(weather),
         climate = as.character(climate)) %>% #convert from factor to remove NAs
  mutate_if(is.character, ~replace_na(.,"")) %>%
  mutate(cat2 = paste(climate, cbii_cat, weather, windspeed, sep = ""),
         cat3 = paste(cat, ecoregion, sep = "-")) %>%
  group_by(cat3, cat2) %>% 
  slice(which.min(time_betwe_round)) %>% 
  mutate(x = time_betwe_round*10, 
         y = cbi_reburn) %>%
  dplyr::select(cat3, cat2, x, y)
```



#### Plot predictions


weather plots
plot interaction between weather and time between fires
```{r}
library(gridExtra)

predictor_plot_func = function(df_preds, df_raw, group_var, slope_cat){
      group_var = factor(group_var, levels = c("extreme", "very high", "high"))


  p = ggplot(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var)) +
         geom_line(size = 1.05)+
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    scale_fill_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
  ylim(0.5, 2.5) +
    # scale_x_continuous(breaks = c(5,10,15,20,25,30,35))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="weather"),
           color=guide_legend(title="weather"))
  
  #add in rug plot and points
  p2 = p + geom_rug(data = df_raw, 
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
    #add point when lower CI crosses 0
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat),
             aes(x=x,y=y, color = cat2), size = 2)
    #add point when slope crosses 0
      # geom_point(data = flat_slope_fit %>% filter(cat3 == slope_cat),
      #        aes(x=x,y=y, color = cat2), size = 2, shape = 21, fill = "white")
    
  
  return(p2)

}

pca = predictor_plot_func(preds_w_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_w_ca_trunc$weather,
                          slope_cat = "weather-California Coast") +
  ggtitle("California Coast") + xlab("") +
  theme(legend.position = "none")   

psw = predictor_plot_func(preds_w_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_w_sw_trunc$weather,
                          slope_cat = "weather-Southwest") +
  ggtitle("Southwest") + xlab("") + ylab("")

pnm = predictor_plot_func(preds_w_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_w_nm_trunc$weather,
                          slope_cat = "weather-Northern Mountains") +
  ggtitle("Northern Mountains") +
  theme(legend.position = "none")    

pwm = predictor_plot_func(preds_w_wm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_w_wm_trunc$weather,
                          slope_cat = "weather-Western Mountains") + 
  ggtitle("Western Mountains") + ylab("") +
  theme(legend.position = "none")    

#save plot

pca + psw + pnm + pwm + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/weather3.png'), height = 5, width = 6)
```

windspeed plots
```{r}
predictor_plot_funcv = function(df_preds, df_raw, group_var, slope_cat){
    group_var = factor(group_var, levels = c("high", "moderate", "low"))

  p = ggplot() +
         geom_line(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var), size = 1.05)+
           geom_ribbon(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    scale_fill_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    ylim(0.5, 2.5) +
    # scale_x_continuous(breaks = c(5,10,15,20,25,30,35))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="wind speed"),
           color=guide_legend(title="wind speed"))
  
  p2 = p + geom_rug(data = df_raw, 
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
    #add point when lower CI crosses 0
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat),
             aes(x=x,y=y, color = cat2), size = 2)
    #add point when slope crosses 0
      # geom_point(data = flat_slope_fit %>% filter(cat3 == slope_cat),
      #        aes(x=x,y=y, color = cat2), size = 2, shape = 21, fill = "white")
  
  return(p2)

}

pca = predictor_plot_funcv(preds_vs_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_vs_ca_trunc$windspeed,
                          slope_cat = "windspeed-California Coast") + 
  ggtitle("California Coast") + xlab("") +
  theme(legend.position = "none")    

pnm = predictor_plot_funcv(preds_vs_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_vs_nm_trunc$windspeed,
                          slope_cat = "windspeed-Northern Mountains") + 
  ggtitle("Northern Mountains") +
  theme(legend.position = "none")  

psw = predictor_plot_funcv(preds_vs_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_vs_sw_trunc$windspeed,
                          slope_cat = "windspeed-Southwest") + 
  ggtitle("Southwest") + xlab("") + ylab("")


pwm = predictor_plot_funcv(preds_vs_wm_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_vs_wm_trunc$windspeed,
                          slope_cat = "windspeed-Western Mountains") + 
  ggtitle("Western Mountains") + ylab("") + theme(legend.position = "none")    


#plot with patchwork
pca + psw + pnm + pwm + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/windspeed3.png'), height = 5, width = 6)
```

CBI plots
```{r}
predictor_plot_func2 = function(df_preds, df_raw, group_var, slope_cat){
    group_var = recode_factor(group_var, mod = "moderate",
                              low = "low") 
    
  p = ggplot() +
         geom_line(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var), size = 1.05)+
           geom_ribbon(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#FFBF69", "lightseagreen"))+
    scale_linetype_manual(values = c(2, 1))+
    scale_fill_manual(values = c("#FFBF69", "lightseagreen"))+
    ylim(0.5, 2.5) +
    # scale_x_continuous(breaks = c(5,10,15,20,25,30,35))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="initial severity"),
           color=guide_legend(title="initial severity"))
  
  p2 = p + geom_rug(data = df_raw, 
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
        #add point when lower CI crosses 0
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat) %>%
                   mutate(cat2 = recode_factor(cat2, mod = "moderate",
                                               low = "low")),
             aes(x=x,y=y, color = cat2), size = 2) 
    #add point when slope crosses 0
      # geom_point(data = flat_slope_fit %>% filter(cat3 == slope_cat) %>%
      #              mutate(cat2 = recode_factor(cat2, mod = "moderate severity (CBI = 2.25)",
      #                                          low = "low severity (CBI = 1.0)")),
      #        aes(x=x,y=y, color = cat2), size = 2, shape = 21, fill = "white")

  
  return(p2)

}

pca = predictor_plot_func2(preds_cbi_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_cbi_ca_trunc$cbii_cat,
                          slope_cat = "cbi-California Coast") + 
  ggtitle("California Coast") + xlab("") +
  theme(legend.position = "none")    

pnm = predictor_plot_func2(preds_cbi_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_cbi_nm_trunc$cbii_cat,
                          slope_cat = "cbi-Northern Mountains") + 
  ggtitle("Northern Mountains") +
  theme(legend.position = "none")  

psw = predictor_plot_func2(preds_cbi_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_cbi_sw_trunc$cbii_cat,
                          slope_cat = "cbi-Southwest") +
  ggtitle("Southwest") + xlab("") + ylab("")


pwm = predictor_plot_func2(preds_cbi_wm_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_cbi_wm_trunc$cbii_cat,
                          slope_cat = "cbi-Western Mountains") +
    ggtitle("Western Mountains") + ylab("") + theme(legend.position = "none") 

#plot with patchwork
pca + psw + pnm + pwm + 
  plot_layout(ncol = 2)


ggsave(datadir('figures/cbi3.png'), height = 5, width = 6)
```

climate plots
plot interaction between climate/ productivity and time between fires
```{r}

predictor_plot_func3 = function(df_preds, df_raw, group_var, slope_cat){
    group_var = recode_factor(group_var,
                              hotdry= "hot-dry",
                              hotwet = "hot-wet",
                              coldwet = "cool-wet") 
    
  p = ggplot(data = df_preds, aes(x = time_betwe_unscale*10, y = cbi_reburn, color = group_var)) +
         geom_line(size = 1.05)+
           geom_ribbon(aes(ymin = lwr, ymax = upr, fill = group_var), alpha = 0.2, colour = NA)+
         theme_bw() +
    scale_color_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
    scale_fill_manual(values = c("#852750", "#FFBF69",  "lightseagreen"))+
  ylim(0.5, 2.8) +
    scale_y_continuous(breaks = c(0.5, 1, 1.5, 2, 2.5))+
    xlab("time between fires") +
    ylab("reburn severity (CBI)")+
   theme(panel.grid.minor = element_blank())+
    guides(fill=guide_legend(title="climate"),
           color=guide_legend(title="climate"))
  
  p2 = p + geom_rug(data = df_raw, 
                    aes(x = time_betwe*10, y = cbi_reburn), 
                    color = "grey", alpha = .3, position = "jitter", length = unit(0.05, "npc"))+
      #add point when lower CI crosses 0
      geom_point(data = flat_slope_preds2 %>% filter(cat3 == slope_cat) %>%
                   mutate(cat2 = recode_factor(cat2,                            
                              hotdry= "hot-dry",
                              hotwet = "hot-wet",
                              coldwet = "cool-wet")),
             aes(x=x,y=y, color = cat2), size = 2)
    #add point when slope crosses 0
      # geom_point(data = flat_slope_fit %>% filter(cat3 == slope_cat) %>%
      #              mutate(cat2 = recode_factor(cat2,                            
      #                                          hotdry= "hot dry",
      #                                          hotwet = "hot wet",
      #                                          coldwet = "cool wet")), 
      #        aes(x=x,y=y, color = cat2), size = 2, shape = 21, fill = "white")

  
  return(p2)

}

pca = predictor_plot_func3(preds_c_ca_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'California Coast'),
                          preds_c_ca_trunc$climate,
                          slope_cat = "climate-California Coast") + 
  ggtitle("California Coast") + xlab("") +
  theme(legend.position = "none")

pnm = predictor_plot_func3(preds_c_nm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Northern Mountains'),
                          preds_c_nm_trunc$climate,
                          slope_cat = "climate-Northern Mountains") + 
  ggtitle("Northern Mountains") +
  theme(legend.position = "none")  

psw = predictor_plot_func3(preds_c_sw_trunc,
                          df_raw = data_thin %>% filter(ecoregion == 'Southwest'),
                          preds_c_sw_trunc$climate,
                          slope_cat = "climate-Southwest") + 
  ggtitle("Southwest") + xlab("") + ylab("")


pwm = predictor_plot_func3(preds_c_wm_trunc, 
                          df_raw = data_thin %>% filter(ecoregion == 'Western Mountains'), 
                          preds_c_wm_trunc$climate,
                          slope_cat = "climate-Western Mountains") + 
  ggtitle("Western Mountains") + ylab("") +
  theme(legend.position = "none")    

library(patchwork)
#plot with patchwork
pca + psw + pnm + pwm + 
  plot_layout(ncol = 2)

ggsave(datadir('figures/climate3.png'), height = 5, width = 6.1)
```
